# Project: Disparities in Sedation Practice by Patient Preferred Language
# Author: Alex Ortiz

# Setup Step - Setup libraries
```{r setup}
# Install Packages

packages <- c("zoo", "data.table", "dplyr", "lubridate", "tidyr", "fst", "arrow", "table1", "lme4", "metafor", "yaml", "pscl", "MASS", "htmltools", "car", "WeightIt", "MatchIt", "purrr", "survival", "mediation", "cmprsk", "gfoRmula", "nnet", "medflex", "bruceR", "tidyverse", "survey", "ipw", "geepack", "readr", "collapse", "stringr", "rprojroot", "jsonlite", "tableone", "lmtest")
install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)

# Load necessary libraries
library(data.table)
library(dplyr)
library(zoo)  # for na.locf function
library(lubridate)
library(tidyr)
library(fst)
library(arrow)
library(table1)
library(metafor)
library(yaml)
library(htmltools)
library(MASS)
library(pscl)
library(car)
library(stringr)
library(rprojroot)
library(jsonlite)
library(tableone)
library(lmtest)


# Read the YAML from the config folder
project_root <- file.path( find_root(rprojroot::has_dir("deep_sedation_language_final")),
  "deep_sedation_language_final")
knitr::opts_knit$set(root.dir = file.path(project_root, "config"))
config <- yaml::read_yaml(file.path(project_root, "config", "deep_sedation_language_config.yaml"))

# Define directories based on YAML configuration
data_dir <- config$data_dir
output_dir <- config$output_dir
intermediate_dir <- file.path(output_dir, "intermediate_data_DO_NOT_UPLOAD/")  # Define output_dir based on data_dir
final_data_dir <- file.path(output_dir, "UPLOAD_ME/")
cohort_characteristics <- file.path(final_data_dir, "cohort_characteristics")
sedation_project <- file.path(final_data_dir, "sedation_project")
sedation_tables_dir <- file.path(sedation_project, "final_tables/")  
sedation_final_data <- file.path(sedation_project, "final_data/")
paralytics_project <- file.path(final_data_dir, "paralytics_project")
paralytics_tables_dir <- file.path(sedation_project, "final_tables/")  
paralytics_final_data <- file.path(sedation_project, "final_data/")
#second_dir <- file.path(output_dir, "secondary_project/")
file_type <- config$file_type
institution <- config$institution
time_zone <- config$time_zone

dirs_to_create <- c(intermediate_dir, final_data_dir, sedation_project, sedation_tables_dir, sedation_final_data, paralytics_project, paralytics_tables_dir, paralytics_final_data, cohort_characteristics)
for (dir in dirs_to_create) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)
  }
}
```

#Upload Data
```{r}
read_data <- function(file_path, select = NULL) {
  if (grepl("\\.csv$|\\.csv\\.gz$", file_path)) {
    # Use fread with the select argument for CSV and CSV.GZ files
    return(fread(file_path, select = all_of(select)))
  } else if (grepl("\\.parquet$", file_path)) {
    # Use read_parquet with col_select for Parquet files
    return(arrow::read_parquet(file_path, col_select = all_of(select)))
  } else if (grepl("\\.fst$", file_path)) {
    # Use read_fst with columns for FST files
    return(fst::read_fst(file_path, columns = all_of(select)))
  } else {
    stop("Unsupported file format. Only CSV, Parquet, and FST are supported.")
  }
}

# Upload Analytic File
total_cohort <- read_parquet(paste0(intermediate_dir, "analytic_file", institution, ".parquet"))

total_cohort$icu_imv <- factor(
  total_cohort$icu_imv,
  levels = c("cardiac_icu", "general_icu", "medical_icu", 
             "mixed_cardiothoracic_icu", "mixed_neuro_icu", "neuro_icu", "neurosurgical_icu", "surgical_icu", "trauma_icu", "burn_icu", "cardiothoracic_surgical_icu"),
  labels = c("Cardiac ICU", "General ICU", "Medical ICU", 
             "Mixed Cardiothoracic ICU", "Mixed Neuro ICU", "Neuro ICU", "Neurosurgical ICU", "Surgical ICU", "Trauma ICU", "Burn ICU", "Cardiothoracic Surgical ICU")
)

catVars <- c("sex_category", "ethnicity_category", "race_category", "hospital_id", "icu_imv", "location_of_intubation")
```

#IPW Calculation
```{r}
ipw_outcome <- total_cohort[, .(
  reached_48h = as.integer(imv_duration_hours_short >= 48)
), by = hospitalization_id]

# Merge the reached_48h flag back to your dataset
ipw_sensitivity_analysis <- merge(total_cohort, ipw_outcome, by = "hospitalization_id")

# Use patient-level covariates to predict censoring
ipw_model <- glm(reached_48h ~ age_category + sex_category + race_category + ethnicity_category + laps2 + factor(hospital_icu_id) + elixhauser_score,
                 data = unique(ipw_sensitivity_analysis, by = "hospitalization_id"),
                 family = binomial())

total_cohort[, prob_reach_48h := predict(ipw_model, newdata = total_cohort, type = "response")]
total_cohort[, ipw := 1 / prob_reach_48h]
```

# STEP ONE: Define Cohorts
# Total Cohort for Non-English Analysis
```{r}
# Identify reference points for model
total_cohort$language_category <- relevel(as.factor(total_cohort$language_category), ref = "english")
total_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(total_cohort$top_five_non_eng_non_spanish), ref = "english")
total_cohort$non_eng <- relevel(as.factor(total_cohort$non_eng), ref = "english")
total_cohort$three_language_cat <- relevel(as.factor(total_cohort$three_language_cat), ref = "english")
total_cohort$sex_category <- relevel(as.factor(total_cohort$sex_category), ref = "male")
total_cohort$race_category <- relevel(as.factor(total_cohort$race_category), ref = "white")
total_cohort$ethnicity_category <- relevel(as.factor(total_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
total_cohort_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | non_eng, data = total_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(total_cohort_table_one), file.path(sedation_tables_dir, paste0(institution, "_total_cohort_table_one_manuscript.html")))

#Create paralytic_early_flag
total_cohort[, paralytic_early_flag := fifelse(
  (time_to_first_paralytic > 0) | (push_paralytics == 1),
  1, 0)]

setnafill(total_cohort, cols = "paralytic_early_flag", fill = 0)

#Create Continuous Paralytic Flag
total_cohort[, paralysis_cont := fifelse(
  time_to_first_paralytic > 4 & time_to_first_paralytic <= 48,
  1L, 0L
)]
setnafill(total_cohort, cols = "paralysis_cont", fill = 0)
```

# Cohort for Top 5 non-English, non-Spanish Language
```{r}
top_five_cohort <- total_cohort[top_five_non_eng_non_spanish %in% c("english", "top_five_non_eng_non_spanish", "spanish")]
top_five_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(top_five_cohort$top_five_non_eng_non_spanish), ref = "english")
top_five_cohort$non_eng <- relevel(as.factor(top_five_cohort$non_eng), ref = "english")
top_five_cohort$sex_category <- relevel(as.factor(top_five_cohort$sex_category), ref = "male")
top_five_cohort$race_category <- relevel(as.factor(top_five_cohort$race_category), ref = "white")
top_five_cohort$ethnicity_category <- relevel(as.factor(top_five_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
top_five_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | top_five_non_eng_non_spanish, data = top_five_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(top_five_table_one), file.path(sedation_tables_dir, paste0(institution, "_top_five_cohort_table_one_manuscript.html")))
```

# Cohort for Sensitivity Analysis Involving All Spanish/English Pts regardless of ICU of admission
```{r}
english_spanish_all_icu_analysis <- total_cohort[language_category %in% c("english", "spanish")]
english_spanish_all_icu_analysis[, race_category := as.character(race_category)]
english_spanish_all_icu_analysis[, race_category := fifelse(
  race_category %in% c("asian"),
  "other/unknown",
  race_category
)]

# Identify reference points for model
english_spanish_all_icu_analysis$language_category <- relevel(as.factor(english_spanish_all_icu_analysis$language_category), ref = "english")
english_spanish_all_icu_analysis$sex_category <- relevel(as.factor(english_spanish_all_icu_analysis$sex_category), ref = "male")
english_spanish_all_icu_analysis$race_category <- relevel(as.factor(english_spanish_all_icu_analysis$race_category), ref = "white")
english_spanish_all_icu_analysis$ethnicity_category <- relevel(as.factor(english_spanish_all_icu_analysis$ethnicity_category), ref = "non-hispanic")


#Basic Table One
# Generate the summary table grouped by language
eng_span_all_icu_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = english_spanish_all_icu_analysis, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(eng_span_all_icu_table_one), file.path(sedation_tables_dir, paste0(institution, "_eng_span_all_icu_cohort_table_one_manuscript.html")))

```

# Primary Cohort for Analysis
```{r}
# Step 1: Count Spanish-speaking patients by hospital_icu_id
spanish_counts <- english_spanish_all_icu_analysis[
  language_category == "spanish", 
  .N, 
  by = hospital_icu_id
]

# Step 2: Identify hospital_icu_id values with 0 Spanish-speaking patients
# (We do this by first getting all unique hospital_icu_id values and checking which ones are missing)
all_hospitals <- unique(english_spanish_all_icu_analysis$hospital_icu_id)
hospitals_with_spanish <- spanish_counts$hospital_icu_id
hospitals_to_remove <- setdiff(all_hospitals, hospitals_with_spanish)

# Step 3: Filter out those hospitals from the original table
english_spanish_primary_analysis <- english_spanish_all_icu_analysis[
  !hospital_icu_id %in% hospitals_to_remove
]

english_spanish_primary_analysis <- english_spanish_primary_analysis %>%
  mutate(language_category = droplevels(language_category))

# Identify reference points for model
english_spanish_primary_analysis$language_category <- relevel(as.factor(english_spanish_primary_analysis$language_category), ref = "english")
english_spanish_primary_analysis$sex_category <- relevel(as.factor(english_spanish_primary_analysis$sex_category), ref = "male")
english_spanish_primary_analysis$race_category <- relevel(as.factor(english_spanish_primary_analysis$race_category), ref = "white")
english_spanish_primary_analysis$ethnicity_category <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category), ref = "non-hispanic")
english_spanish_primary_analysis$covid_status <- relevel(as.factor(english_spanish_primary_analysis$covid_status), ref = "not detected")
english_spanish_primary_analysis$ethnicity_category_final <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")
english_spanish_primary_analysis$push_paralytics <- relevel(as.factor(english_spanish_primary_analysis$push_paralytics), ref = "0")


# Create labels for readability
label(english_spanish_primary_analysis$imv_duration_days) <- "<u>Days Mechanically Ventilated</u>"
label(english_spanish_primary_analysis$race_category) <- "<u>Race</u>"
label(english_spanish_primary_analysis$ethnicity_category) <- "<u>Ethnicity</u>"
label(english_spanish_primary_analysis$sex_category) <- "<u>Sex</u>"
label(english_spanish_primary_analysis$age_at_admission) <- "<u>Age</u>"
label(english_spanish_primary_analysis$language_category) <- "Language"
label(english_spanish_primary_analysis$bmi) <- "<u>BMI</u>"
label(english_spanish_primary_analysis$hospital_id) <- "<u>Hospital</u>"
label(english_spanish_primary_analysis$icu_imv) <- "<u>ICU Type</u>"
label(english_spanish_primary_analysis$thirty_day_mortality) <- "<u>Thirty Day Mortality</u>"
label(english_spanish_primary_analysis$location_of_intubation) <- "<u>Location of Intubation</u>"
label(english_spanish_primary_analysis$elixhauser_score) <- "<u>Elixhauser Score</u>"
label(english_spanish_primary_analysis$SVM_SCORE) <- "<u>Social Vulnerability Metric</u>"
label(english_spanish_primary_analysis$laps2) <- "<u>LAPS2 Score</u>"
label(english_spanish_primary_analysis$time_to_first_rass) <- "<u>Time to First RASS</u>"
label(english_spanish_primary_analysis$rass_frequency_hrs) <- "<u>Frequency Of Rass Measurements</u>"
label(english_spanish_primary_analysis$covid_status) <- "<u>COVID Status</u>"

# Generate the summary table grouped by language
summary_table <- table1(
  ~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + location_of_intubation + elixhauser_score + SVM_SCORE + time_to_first_rass + rass_frequency_hrs 
  | language_category, 
  data = english_spanish_primary_analysis,
  factorVars = catVars,
  test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(summary_table), file.path(sedation_tables_dir, paste0(institution, "_primary_cohort_table_one_manuscript.html")))

vars <- c("age_at_admission", "bmi", "sex_category", "race_category", "ethnicity_category", "laps2", "hospital_id", "icu_imv", "location_of_intubation", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs","covid_status")
catVars <- c("sex_category", "race_category", "ethnicity_category", "hospital_id", "icu_imv", "location_of_intubation")

# Create table
table1_obj <- CreateTableOne(vars = vars, strata = "language_category",
                             data = english_spanish_primary_analysis,
                             factorVars = catVars)

# Print with SMD and nonparametric summary
print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi"))

print(table1_obj, smd = TRUE)

# Convert the table1 object to a data frame
table1_df <- print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi", "laps2", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs"), printToggle = FALSE)

# Save to CSV
# Save the combined results to a file
fwrite(table1_df, file.path(sedation_tables_dir, paste0(institution, "_IQR_SMD_Table1.csv")))

#Create paralytic_early_flag
english_spanish_primary_analysis[, paralytic_early_flag := fifelse(
  (time_to_first_paralytic > 0) | (push_paralytics == 1),
  1, 0)]

setnafill(english_spanish_primary_analysis, cols = "paralytic_early_flag", fill = 0)

#Create Continuous Paralytic Flag
english_spanish_primary_analysis[, paralysis_cont := fifelse(
  time_to_first_paralytic > 4 & time_to_first_paralytic <= 48,
  1L, 0L
)]
setnafill(english_spanish_primary_analysis, cols = "paralysis_cont", fill = 0)

english_spanish_primary_analysis[, benzos := fifelse(!is.na(loraz_median), 1, 0)]
```

#Cohort Intubated for Full 48 Hours
```{r}
full_48hr_cohort <- english_spanish_primary_analysis[  hours_mechanically_ventilated_72 >= 48]
full_48hr_cohort <- full_48hr_cohort[
  is.finite(ipw) & !is.na(ipw)]

# Identify reference points for model
full_48hr_cohort$language_category <- relevel(as.factor(full_48hr_cohort$language_category), ref = "english")
full_48hr_cohort$sex_category <- relevel(as.factor(full_48hr_cohort$sex_category), ref = "male")
full_48hr_cohort$race_category <- relevel(as.factor(full_48hr_cohort$race_category), ref = "white")
full_48hr_cohort$ethnicity_category <- relevel(as.factor(full_48hr_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
full_48hr_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = full_48hr_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(full_48hr_table_one), file.path(sedation_tables_dir, paste0(institution, "_full_48hr_cohort_table_one_manuscript.html")))
```

#Cohort with RASS Measurements within 2 Hours
```{r}
rass_2hrs_cohort <- english_spanish_primary_analysis[time_to_first_rass <= 2]

# Identify reference points for model
rass_2hrs_cohort$language_category <- relevel(as.factor(rass_2hrs_cohort$language_category), ref = "english")
rass_2hrs_cohort$sex_category <- relevel(as.factor(rass_2hrs_cohort$sex_category), ref = "male")
rass_2hrs_cohort$race_category <- relevel(as.factor(rass_2hrs_cohort$race_category), ref = "white")
rass_2hrs_cohort$ethnicity_category <- relevel(as.factor(rass_2hrs_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
rass_2hrs_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = rass_2hrs_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(rass_2hrs_table_one), file.path(sedation_tables_dir, paste0(institution, "_rass_2hrs_cohort_table_one_manuscript.html")))

```

#Cohort without Neuro_ICU patients
```{r}
no_neuro_cohort <- english_spanish_primary_analysis[
  !icu_imv %in% c("Neuro ICU", "Neurosurgical ICU", "Mixed Neuro ICU")]

# Identify reference points for model
no_neuro_cohort$language_category <- relevel(as.factor(no_neuro_cohort$language_category), ref = "english")
no_neuro_cohort$sex_category <- relevel(as.factor(no_neuro_cohort$sex_category), ref = "male")
no_neuro_cohort$race_category <- relevel(as.factor(no_neuro_cohort$race_category), ref = "white")
no_neuro_cohort$ethnicity_category <- relevel(as.factor(no_neuro_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
no_neuro_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = no_neuro_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(no_neuro_table_one), file.path(sedation_tables_dir, paste0(institution, "_no_neuro_cohort_table_one_manuscript.html")))

```

#Cohort without Paralytic Patients
```{r}
no_paralytic_cohort <- english_spanish_primary_analysis[
  (is.na(time_to_first_paralytic) & push_paralytics == 0)
]

# Identify reference points for model
no_paralytic_cohort$language_category <- relevel(as.factor(no_paralytic_cohort$language_category), ref = "english")
no_paralytic_cohort$sex_category <- relevel(as.factor(no_paralytic_cohort$sex_category), ref = "male")
no_paralytic_cohort$race_category <- relevel(as.factor(no_paralytic_cohort$race_category), ref = "white")
no_paralytic_cohort$ethnicity_category <- relevel(as.factor(no_paralytic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
no_paralytic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = no_paralytic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(no_paralytic_table_one), file.path(sedation_tables_dir, paste0(institution, "_no_paralytic_cohort_table_one_manuscript.html")))

```

#Cohort with only Paralytic Patients
```{r}
only_paralytic_cohort <- english_spanish_primary_analysis[!
  (is.na(time_to_first_paralytic) & push_paralytics == 0)
]

# Identify reference points for model
only_paralytic_cohort$language_category <- relevel(as.factor(only_paralytic_cohort$language_category), ref = "english")
only_paralytic_cohort$sex_category <- relevel(as.factor(only_paralytic_cohort$sex_category), ref = "male")
only_paralytic_cohort$race_category <- relevel(as.factor(only_paralytic_cohort$race_category), ref = "white")
only_paralytic_cohort$ethnicity_category <- relevel(as.factor(only_paralytic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
only_paralytic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + covid_status | language_category, data = only_paralytic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(only_paralytic_table_one), file.path(sedation_tables_dir, paste0(institution, "_only_paralytic_cohort_table_one_manuscript.html")))

```

#Cohort removing continuous paralytic patients but keeping push paralytics
```{r}
no_continuous_paralytic_cohort <- english_spanish_primary_analysis[
  (is.na(time_to_first_paralytic))]

# Identify reference points for model
no_continuous_paralytic_cohort$language_category <- relevel(as.factor(no_continuous_paralytic_cohort$language_category), ref = "english")
no_continuous_paralytic_cohort$sex_category <- relevel(as.factor(no_continuous_paralytic_cohort$sex_category), ref = "male")
no_continuous_paralytic_cohort$race_category <- relevel(as.factor(no_continuous_paralytic_cohort$race_category), ref = "white")
no_continuous_paralytic_cohort$ethnicity_category <- relevel(as.factor(no_continuous_paralytic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
no_continuous_paralytic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + imv_duration_days + thirty_day_mortality | language_category, data = no_continuous_paralytic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(no_continuous_paralytic_table_one), file.path(sedation_tables_dir, paste0(institution, "_no_continuous_paralytic_cohort_table_one_manuscript.html")))

```

#Cohort removing push paralytic patients but keeping continuous paralytics
```{r}
no_push_paralytic_cohort <- english_spanish_primary_analysis[
  (push_paralytics == 0)
]

# Identify reference points for model
no_push_paralytic_cohort$language_category <- relevel(as.factor(no_push_paralytic_cohort$language_category), ref = "english")
no_push_paralytic_cohort$sex_category <- relevel(as.factor(no_push_paralytic_cohort$sex_category), ref = "male")
no_push_paralytic_cohort$race_category <- relevel(as.factor(no_push_paralytic_cohort$race_category), ref = "white")
no_push_paralytic_cohort$ethnicity_category <- relevel(as.factor(no_push_paralytic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
no_push_paralytic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + imv_duration_days + thirty_day_mortality | language_category, data = no_push_paralytic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(no_push_paralytic_table_one), file.path(sedation_tables_dir, paste0(institution, "_no_push_paralytic_cohort_table_one_manuscript.html")))

```

#Pandemic Cohort
```{r}
pandemic_cohort <- english_spanish_primary_analysis[
  admission_dttm >= as.POSIXct("2020-03-01 00:00:00") &
  admission_dttm <= as.POSIXct("2023-05-11 23:59:59")
]

# Identify reference points for model
pandemic_cohort$language_category <- relevel(as.factor(pandemic_cohort$language_category), ref = "english")
pandemic_cohort$sex_category <- relevel(as.factor(pandemic_cohort$sex_category), ref = "male")
pandemic_cohort$race_category <- relevel(as.factor(pandemic_cohort$race_category), ref = "white")
pandemic_cohort$ethnicity_category <- relevel(as.factor(pandemic_cohort$ethnicity_category), ref = "non-hispanic")
pandemic_cohort$covid_status <- relevel(as.factor(pandemic_cohort$covid_status), ref = "not detected")

#Basic Table One
# Generate the summary table grouped by language
pandemic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + covid_status | language_category, data = pandemic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(pandemic_table_one), file.path(sedation_tables_dir, paste0(institution, "_pandemic_cohort_table_one_manuscript.html")))

```

#Pre-Pandemic Cohort
```{r}
pre_pandemic_cohort <- english_spanish_primary_analysis[
  admission_dttm < as.POSIXct("2020-03-01 00:00:00")]

# Identify reference points for model
pre_pandemic_cohort$language_category <- relevel(as.factor(pre_pandemic_cohort$language_category), ref = "english")
pre_pandemic_cohort$sex_category <- relevel(as.factor(pre_pandemic_cohort$sex_category), ref = "male")
pre_pandemic_cohort$race_category <- relevel(as.factor(pre_pandemic_cohort$race_category), ref = "white")
pre_pandemic_cohort$ethnicity_category <- relevel(as.factor(pre_pandemic_cohort$ethnicity_category), ref = "non-hispanic")
#pre_pandemic_cohort$covid_status <- relevel(as.factor(pre_pandemic_cohort$covid_status), ref = "Not Detected")

#Basic Table One
# Generate the summary table grouped by language
pre_pandemic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + covid_status | language_category, data = pre_pandemic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(pre_pandemic_table_one), file.path(sedation_tables_dir, paste0(institution, "_pre_pandemic_cohort_table_one_manuscript.html")))

```


#Endemic Cohort
```{r}
endemic_cohort <- english_spanish_primary_analysis[
  admission_dttm > as.POSIXct("2023-05-11 23:59:59")]

# Identify reference points for model
endemic_cohort$language_category <- relevel(as.factor(endemic_cohort$language_category), ref = "english")
endemic_cohort$sex_category <- relevel(as.factor(endemic_cohort$sex_category), ref = "male")
endemic_cohort$race_category <- relevel(as.factor(endemic_cohort$race_category), ref = "white")
endemic_cohort$ethnicity_category <- relevel(as.factor(endemic_cohort$ethnicity_category), ref = "non-hispanic")
endemic_cohort$covid_status <- relevel(as.factor(endemic_cohort$covid_status), ref = "not detected")

#Basic Table One
# Generate the summary table grouped by language
endemic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + covid_status | language_category, data = endemic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(endemic_table_one), file.path(sedation_tables_dir, paste0(institution, "_pre_pandemic_cohort_table_one_manuscript.html")))

```

#MICU Cohort
```{r}
micu_only_cohort <- english_spanish_primary_analysis[icu_imv == "Medical ICU"]

# Identify reference points for model
micu_only_cohort$language_category <- relevel(as.factor(micu_only_cohort$language_category), ref = "english")
micu_only_cohort$sex_category <- relevel(as.factor(micu_only_cohort$sex_category), ref = "male")
micu_only_cohort$race_category <- relevel(as.factor(micu_only_cohort$race_category), ref = "white")
micu_only_cohort$ethnicity_category <- relevel(as.factor(micu_only_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
micu_only_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = micu_only_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(micu_only_table_one), file.path(sedation_tables_dir, paste0(institution, "_micu_only_table_one_manuscript.html")))

```

#Hispanic Cohort
```{r}
hispanic_cohort <- english_spanish_primary_analysis[ethnicity_category == "hispanic"]

# Identify reference points for model
hispanic_cohort$language_category <- relevel(as.factor(hispanic_cohort$language_category), ref = "english")
hispanic_cohort$sex_category <- relevel(as.factor(hispanic_cohort$sex_category), ref = "male")
hispanic_cohort$race_category <- relevel(as.factor(hispanic_cohort$race_category), ref = "white")
hispanic_cohort$ethnicity_category <- relevel(as.factor(hispanic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
hispanic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = hispanic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(hispanic_table_one), file.path(sedation_tables_dir, paste0(institution, "_hispanic_table_one_manuscript.html")))

```

#Non-Hispanic / Unknown Cohort
```{r}
non_hispanic_cohort <- english_spanish_primary_analysis[ethnicity_category_final == "non-hispanic/unknown"]

# Identify reference points for model
non_hispanic_cohort$language_category <- relevel(as.factor(non_hispanic_cohort$language_category), ref = "english")
non_hispanic_cohort$sex_category <- relevel(as.factor(non_hispanic_cohort$sex_category), ref = "male")
non_hispanic_cohort$race_category <- relevel(as.factor(non_hispanic_cohort$race_category), ref = "white")
non_hispanic_cohort$ethnicity_category <- relevel(as.factor(non_hispanic_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
non_hispanic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = non_hispanic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(non_hispanic_table_one), file.path(sedation_tables_dir, paste0(institution, "_non_hispanic_table_one_manuscript.html")))

```

#Ethnicity/Language Re-Classification Cohort
```{r}
language_reclass_cohort <- copy(english_spanish_primary_analysis)
language_reclass_cohort[
  ethnicity_category == "unknown" & language_category == "spanish",
  ethnicity_category := "hispanic"
]

language_reclass_cohort[
  ethnicity_category == "unknown" & language_category == "english",
  ethnicity_category := "non-hispanic"
]

language_reclass_cohort[
  ethnicity_category == "non-hispanic" & language_category == "spanish",
  ethnicity_category := "hispanic"
]

# Identify reference points for model
language_reclass_cohort$language_category <- relevel(as.factor(language_reclass_cohort$language_category), ref = "english")
language_reclass_cohort$sex_category <- relevel(as.factor(language_reclass_cohort$sex_category), ref = "male")
language_reclass_cohort$race_category <- relevel(as.factor(language_reclass_cohort$race_category), ref = "white")
language_reclass_cohort$ethnicity_category <- relevel(as.factor(language_reclass_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
language_reclass_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = language_reclass_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(language_reclass_table_one), file.path(sedation_tables_dir, paste0(institution, "_language_reclass_table_one_manuscript.html")))

```

#Cohort wihout unknown ethnicity patients
```{r}
no_unknown_ethnicity_cohort <- english_spanish_primary_analysis[
  ethnicity_category != "Unknown"
]

# Identify reference points for model
no_unknown_ethnicity_cohort$language_category <- relevel(as.factor(no_unknown_ethnicity_cohort$language_category), ref = "english")
no_unknown_ethnicity_cohort$sex_category <- relevel(as.factor(no_unknown_ethnicity_cohort$sex_category), ref = "male")
no_unknown_ethnicity_cohort$race_category <- relevel(as.factor(no_unknown_ethnicity_cohort$race_category), ref = "white")
no_unknown_ethnicity_cohort$ethnicity_category <- relevel(as.factor(no_unknown_ethnicity_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
no_unknown_ethnicity_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + imv_duration_days + thirty_day_mortality | language_category, data = no_unknown_ethnicity_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(no_unknown_ethnicity_table_one), file.path(sedation_tables_dir, paste0(institution, "_no_unknown_ethnicity_table_one_manuscript.html")))

```

#Non-English Analysis Cohort without Paralytic Patients
```{r}
non_english_no_paralytic_cohort <- total_cohort[
  (is.na(time_to_first_paralytic) & push_paralytics == 0)
]

# Identify reference points for model
non_english_no_paralytic_cohort$language_category <- relevel(as.factor(non_english_no_paralytic_cohort$language_category), ref = "english")
non_english_no_paralytic_cohort$sex_category <- relevel(as.factor(non_english_no_paralytic_cohort$sex_category), ref = "male")
non_english_no_paralytic_cohort$race_category <- relevel(as.factor(non_english_no_paralytic_cohort$race_category), ref = "white")
non_english_no_paralytic_cohort$ethnicity_category <- relevel(as.factor(non_english_no_paralytic_cohort$ethnicity_category), ref = "non-hispanic")
non_english_no_paralytic_cohort$non_eng <- relevel(as.factor(non_english_no_paralytic_cohort$non_eng), ref = "english")

#Basic Table One
# Generate the summary table grouped by language
non_english_no_paralytic_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | non_eng, data = non_english_no_paralytic_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(non_english_no_paralytic_table_one), file.path(sedation_tables_dir, paste0(institution, "_non_english_no_paralytic_cohort_table_one.html")))
```

#STEP 2: PRIMARY ANALYSIS
```{r}
# Fit the unadjusted primary model
primary_model_unadjusted <- lm(percent_deep_sedation_48 ~ language_category, data = english_spanish_primary_analysis)

summary(primary_model_unadjusted)

# Extract coefficients, standard errors, variance, and p-values for the unadjusted model
primary_results_unadjusted <- data.frame(
  term = names(coef(primary_model_unadjusted)),
  estimate = coef(primary_model_unadjusted),
  std.error = summary(primary_model_unadjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_unadjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_unadjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

primary_model_adjusted <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(primary_model_adjusted)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
primary_results_adjusted <- data.frame(
  term = names(coef(primary_model_adjusted)),
  estimate = coef(primary_model_adjusted),
  std.error = summary(primary_model_adjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_adjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_adjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Combine the results into one data frame
primary_results_combined <- bind_rows(
  primary_results_unadjusted %>% mutate(model = "Unadjusted"),
  primary_results_adjusted %>% mutate(model = "Adjusted")
)

# Save the combined results to a file
fwrite(primary_results_combined, file.path(final_data, paste0(institution, "_primary_analysis_results.csv")))
```

#STEP 3: SENSITIVITY ANALYSES
#Sensitivity Analysis #1: RASS Imputation of -1 instead of -5 to start
```{r}
sensitivity_model_one <- lm(percent_deep_sedation_48_sensitivity ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(sensitivity_model_one)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_one_results <- data.frame(
  term = names(coef(sensitivity_model_one)),
  estimate = coef(sensitivity_model_one),
  std.error = summary(sensitivity_model_one)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_one)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_one)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)


# Save the results to a file
fwrite(sensitivity_model_one_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_rass_imputation.csv")))
```

#Sensitivity Analysis #2: First 72 hours of IMV instead of first 48 hours
```{r}
sensitivity_model_two <- lm(percent_deep_sedation_72 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(sensitivity_model_two)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_two_results <- data.frame(
  term = names(coef(sensitivity_model_two)),
  estimate = coef(sensitivity_model_two),
  std.error = summary(sensitivity_model_two)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_two)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_two)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_two_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_72hr.csv")))
```

#Sensitivity Analysis #3: Only include patients intubated for full 48-hours to deal with censoring
#Rerun Analysis #5: Only include patients intubated for full 48-hours to deal with censoring - improved robust SEs
```{r}

#Cohort Intubated for Full 48 Hours
full_48hr_cohort <- english_spanish_primary_analysis[  hours_mechanically_ventilated_72 >= 48]

#Trim IPWs
full_48hr_cohort <- full_48hr_cohort[
  is.finite(ipw) & !is.na(ipw)]
ipw_q <- quantile(full_48hr_cohort$ipw, c(0.01, 0.99), na.rm = TRUE)
full_48hr_cohort[, ipw := pmin(pmax(ipw, ipw_q[1]), ipw_q[2])]

# Identify reference points for model
full_48hr_cohort$language_category <- relevel(as.factor(full_48hr_cohort$language_category), ref = "english")
full_48hr_cohort$sex_category <- relevel(as.factor(full_48hr_cohort$sex_category), ref = "male")
full_48hr_cohort$race_category <- relevel(as.factor(full_48hr_cohort$race_category), ref = "white")
full_48hr_cohort$ethnicity_category <- relevel(as.factor(full_48hr_cohort$ethnicity_category), ref = "non-hispanic")

# Step 1: Fit Model with IPW weighting
sensitivity_model_three <- lm(
  percent_deep_sedation_48 ~ language_category + race_category + age_category +
    sex_category + laps2 + elixhauser_score + SVM_SCORE + bmi + factor(hospital_icu_id),
  data = full_48hr_cohort,
  weights = ipw
)

# Step 2: Calculate robust standard errors with a fallback
robust_se <- tryCatch(
  vcovHC(sensitivity_model_three, type = "HC3"),
  error = function(e) NULL
)

if (is.null(robust_se) || anyNA(robust_se)) {
  warning("HC3 robust SE failed at this site; falling back to HC0.")
  robust_se <- vcovHC(sensitivity_model_three, type = "HC0")
}

# Step 3: Coefficient table
sensitivity_model_three_sandwich <- coeftest(sensitivity_model_three, vcov = robust_se)

# Step 3: Output coefficient table with robust SEs
sensitivity_model_three_sandwich <- coeftest(sensitivity_model_three, vcov = robust_se)
print(sensitivity_model_three_sandwich)

# Step 4: Extract results from coeftest matrix
sensitivity_model_three_results <- data.frame(
  term = rownames(sensitivity_model_three_sandwich),
  estimate = sensitivity_model_three_sandwich[, "Estimate"],
  std.error = sensitivity_model_three_sandwich[, "Std. Error"],
  variance = sensitivity_model_three_sandwich[, "Std. Error"]^2,
  p.value = sensitivity_model_three_sandwich[, "Pr(>|t|)"]
)

# Step 5: Save the results
fwrite(sensitivity_model_three_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_full_48hr_rerun.csv")))
```
```{r}
# Step 1: Fit Model with IPW weighting
#sensitivity_model_three <- lm(
#  percent_deep_sedation_48 ~ language_category + race_category + age_category +
#    sex_category + laps2 + elixhauser_score + SVM_SCORE + bmi + factor(hospital_icu_id),
#  data = full_48hr_cohort,
#  weights = ipw
#)

# Step 2: Calculate robust standard errors using sandwich estimator
#robust_se <- vcovHC(sensitivity_model_three, type = "HC3")  # You can change to "HC0" if needed

# Step 3: Output coefficient table with robust SEs
#sensitivity_model_three_sandwich <- coeftest(sensitivity_model_three, vcov = robust_se)
#print(sensitivity_model_three_sandwich)

# Step 4: Extract results from coeftest matrix
#sensitivity_model_three_results <- data.frame(
#  term = rownames(sensitivity_model_three_sandwich),
#  estimate = sensitivity_model_three_sandwich[, "Estimate"],
#  std.error = sensitivity_model_three_sandwich[, "Std. Error"],
#  variance = sensitivity_model_three_sandwich[, "Std. Error"]^2,
#  p.value = sensitivity_model_three_sandwich[, "Pr(>|t|)"]
#)

# Step 5: Save the results
#fwrite(sensitivity_model_three_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_full_48hr.csv")))
```


#Sensitivity Analysis #4: Only include patients with a documented RASS within 2 hours of IMV instead of 8
```{r}
sensitivity_model_four <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = rass_2hrs_cohort)

summary(sensitivity_model_four)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_four_results <- data.frame(
  term = names(coef(sensitivity_model_four)),
  estimate = coef(sensitivity_model_four),
  std.error = summary(sensitivity_model_four)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_four)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_four)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_four_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_rass_2hrs.csv")))
```


#Sensitivity Analysis #5: Remove Neuro_ICU Patients
```{r}
sensitivity_model_five <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = no_neuro_cohort)

summary(sensitivity_model_five)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_five_results <- data.frame(
  term = names(coef(sensitivity_model_five)),
  estimate = coef(sensitivity_model_five),
  std.error = summary(sensitivity_model_five)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_five)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_five)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_five_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_no_neuro.csv")))
```

#Sensitivity Analysis #6: Remove patients who received pushes of paralytics
```{r}
sensitivity_model_six <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = no_push_paralytic_cohort)

summary(sensitivity_model_six)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_six_results <- data.frame(
  term = names(coef(sensitivity_model_six)),
  estimate = coef(sensitivity_model_six),
  std.error = summary(sensitivity_model_six)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_six)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_six)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_six_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_no_push_paralytics.csv")))
```

#Sensitivity Analysis #7: Alternative Definition of Deep Sedation - Weighted Average Rass over first 48_hours
```{r}
sensitivity_model_seven <- glm(deep_sedation_weighted ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(sensitivity_model_seven)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_seven <- summary(sensitivity_model_seven)$coefficients
odds_ratios <- exp(coef_summary_seven[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_seven[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_seven[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
sensitivity_model_seven_results <- data.frame(
    Variable = rownames(coef_summary_seven),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
sensitivity_model_seven_results

# Write the results to a CSV file
fwrite(sensitivity_model_seven_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_weighted_deep_sedation.csv")))
```

#Sensitivity Analysis #8: Hispanic Collinearity - Kappa statistic
```{r}
# Build contingency table
tab <- table(english_spanish_primary_analysis$ethnicity_category_final,
             english_spanish_primary_analysis$language_category)

# Run kappa
kappa_site <- psych::cohen.kappa(tab)

# Extract unweighted kappa and z
kappa_hat <- kappa_site$kappa
var_kappa <- kappa_site$var.kappa
se_kappa <- sqrt(kappa_site$var.kappa)


kappa_results_primary <- data.table(
  site = institution,     # change this label as needed
  kappa_hat = kappa_site$kappa,
  var_kappa = kappa_site$var.kappa,
  se_kappa = sqrt(kappa_site$var.kappa)
)

# Write the results to a CSV file
fwrite(kappa_results_primary, file = file.path(final_data, paste0(institution, "_kappa_results_primary.csv")))

#Sensitivity Analysis
# Build contingency table
tab_sensitivity <- table(no_unknown_ethnicity_cohort$ethnicity_category_final,
             no_unknown_ethnicity_cohort$language_category)

# Run kappa
kappa_site_sensitivity <- psych::cohen.kappa(tab_sensitivity)

# Extract unweighted kappa and z
kappa_hat_sensitivity <- kappa_site_sensitivity$kappa
var_kappa_sensitivity <- kappa_site_sensitivity$var.kappa
se_kappa_sensitivity <- sqrt(kappa_site_sensitivity$var.kappa)


kappa_results_sensitivity <- data.table(
  site = institution,     # change this label as needed
  kappa_hat_sensitivity = kappa_site_sensitivity$kappa,
  var_kappa_sensitivity = kappa_site_sensitivity$var.kappa,
  se_kappa_sensitivity = sqrt(kappa_site_sensitivity$var.kappa)
)

# Write the results to a CSV file
fwrite(kappa_results_sensitivity, file.path(final_data, paste0(institution, "_kappa_results_sensitivity.csv")))

#Sensitivity Analysis #2
# Build contingency table
tab_sensitivity_2 <- table(language_reclass_cohort$ethnicity_category,
             language_reclass_cohort$language_category)

# Run kappa
kappa_site_sensitivity_2 <- psych::cohen.kappa(tab_sensitivity_2)

# Extract unweighted kappa and z
kappa_hat_sensitivity_2 <- kappa_site_sensitivity_2$kappa
var_kappa_sensitivity_2 <- kappa_site_sensitivity_2$var.kappa
se_kappa_sensitivity_2 <- sqrt(kappa_site_sensitivity_2$var.kappa)


kappa_results_sensitivity_2 <- data.table(
  site = institution,     # change this label as needed
  kappa_hat_sensitivity_2 = kappa_site_sensitivity_2$kappa,
  var_kappa_sensitivity_2 = kappa_site_sensitivity_2$var.kappa,
  se_kappa_sensitivity_2 = sqrt(kappa_site_sensitivity_2$var.kappa)
)

# Write the results to a CSV file
fwrite(kappa_results_sensitivity_2, file.path(final_data, paste0(institution, "_kappa_results_sensitivity_2.csv")))
```


#Sensitivity Analysis #9: English vs Non-English
```{r}
sensitivity_model_nine <- lm(percent_deep_sedation_48 ~ non_eng + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = total_cohort)

summary(sensitivity_model_nine)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_nine_results <- data.frame(
  term = names(coef(sensitivity_model_nine)),
  estimate = coef(sensitivity_model_nine),
  std.error = summary(sensitivity_model_nine)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_nine)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_nine)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_nine_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_eng_non-eng.csv")))

```

#Sensitivity Analysis #10: Completely Remove Patients with any Paralytics
```{r}
sensitivity_model_ten <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = no_paralytic_cohort)

summary(sensitivity_model_ten)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_ten_results <- data.frame(
  term = names(coef(sensitivity_model_ten)),
  estimate = coef(sensitivity_model_ten),
  std.error = summary(sensitivity_model_ten)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_ten)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_ten)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_ten_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_no_paralysis.csv")))
```

#Sensitivity Analysis #11: Non-English, Non-Spanish Sensitivity
```{r}
sensitivity_model_eleven <- lm(percent_deep_sedation_48 ~ top_five_non_eng_non_spanish + race_category + ethnicity_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = top_five_cohort)

summary(sensitivity_model_eleven)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_eleven_results <- data.frame(
  term = names(coef(sensitivity_model_eleven)),
  estimate = coef(sensitivity_model_eleven),
  std.error = summary(sensitivity_model_eleven)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_eleven)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_eleven)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_eleven_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_top_five.csv")))
```

#Sensitivity Analysis #12: English, Spanish, All ICUs
```{r}
sensitivity_model_twelve <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_all_icu_analysis)

summary(sensitivity_model_twelve)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_twelve_results <- data.frame(
  term = names(coef(sensitivity_model_twelve)),
  estimate = coef(sensitivity_model_twelve),
  std.error = summary(sensitivity_model_twelve)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_twelve)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_twelve)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_twelve_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_all_icus.csv")))
```

#Sensitivity Analysis #13: Total Hours Deep Sedation
```{r}
sensitivity_model_thirteen <- lm(deep_sedation_count_total_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(sensitivity_model_thirteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_thirteen_results <- data.frame(
  term = names(coef(sensitivity_model_thirteen)),
  estimate = coef(sensitivity_model_thirteen),
  std.error = summary(sensitivity_model_thirteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_thirteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_thirteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_thirteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_total_hours_deep_sedation.csv")))
```


#Sensitivity Analysis #14: Inclusion of ethnicity_category
```{r}
sensitivity_model_fourteen <- lm(percent_deep_sedation_48 ~ language_category + ethnicity_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(sensitivity_model_fourteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_fourteen_results <- data.frame(
  term = names(coef(sensitivity_model_fourteen)),
  estimate = coef(sensitivity_model_fourteen),
  std.error = summary(sensitivity_model_fourteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_fourteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_fourteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_fourteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_ethnicity_inclusion.csv")))
```

#Sensitivity Analysis #15: Include ethnicity_category with ethnicity_category re-classified per preferred language
```{r}
sensitivity_model_fifteen <- lm(percent_deep_sedation_48 ~ language_category + ethnicity_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = language_reclass_cohort)

summary(sensitivity_model_fifteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_fifteen_results <- data.frame(
  term = names(coef(sensitivity_model_fifteen)),
  estimate = coef(sensitivity_model_fifteen),
  std.error = summary(sensitivity_model_fifteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_fifteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_fifteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_fifteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_ethnicity_reclassification.csv")))
```

#Sensitivity Analysis #16: Exclude Unknown Ethnicity Patients
```{r}
sensitivity_model_sixteen <- lm(percent_deep_sedation_48 ~ language_category + ethnicity_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = no_unknown_ethnicity_cohort)

summary(sensitivity_model_sixteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_sixteen_results <- data.frame(
  term = names(coef(sensitivity_model_sixteen)),
  estimate = coef(sensitivity_model_sixteen),
  std.error = summary(sensitivity_model_sixteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_sixteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_sixteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_sixteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_no_unknown_ethnicity.csv")))
```

#Sensitivity Analysis #17: Only Paralytic Patients
```{r}
sensitivity_model_seventeen <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = only_paralytic_cohort)

summary(sensitivity_model_seventeen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_seventeen_results <- data.frame(
  term = names(coef(sensitivity_model_seventeen)),
  estimate = coef(sensitivity_model_seventeen),
  std.error = summary(sensitivity_model_seventeen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_seventeen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_seventeen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_seventeen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_only_paralytic_patients.csv")))
```

#Sensitivity Analysis #18: Non-English, No Paralytic Patients
```{r}
sensitivity_model_eighteen <- lm(percent_deep_sedation_48 ~ non_eng + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = non_english_no_paralytic_cohort)

summary(sensitivity_model_eighteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_eighteen_results <- data.frame(
  term = names(coef(sensitivity_model_eighteen)),
  estimate = coef(sensitivity_model_eighteen),
  std.error = summary(sensitivity_model_eighteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_eighteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_eighteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_eighteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_non_english_no_paralytic_patients.csv")))
```

#Sensitivity Analysis #19: No Continuous Paralytics
```{r}
sensitivity_model_nineteen <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = no_continuous_paralytic_cohort)

summary(sensitivity_model_nineteen)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_nineteen_results <- data.frame(
  term = names(coef(sensitivity_model_nineteen)),
  estimate = coef(sensitivity_model_nineteen),
  std.error = summary(sensitivity_model_nineteen)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_nineteen)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_nineteen)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_nineteen_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_no_continuous_paralytic_patients.csv")))
```

#Sensitivity Analysis #20: Controlling for Paralytics
```{r}
sensitivity_model_twenty <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + paralytic_early_flag,
               data = english_spanish_primary_analysis)

summary(sensitivity_model_twenty)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
sensitivity_model_twenty_results <- data.frame(
  term = names(coef(sensitivity_model_twenty)),
  estimate = coef(sensitivity_model_twenty),
  std.error = summary(sensitivity_model_twenty)$coefficients[, "Std. Error"],
  variance = (summary(sensitivity_model_twenty)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(sensitivity_model_twenty)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(sensitivity_model_twenty_results, file.path(final_data, paste0(institution, "_sensitivity_analysis_controlling_for_paralytics.csv")))
```


#STEP FOUR: SECONDARY ANALYSES
#Secondary Analysis #1: Total time mechnically ventilated
#Need to decide imv_time short or long ###
```{r}
secondary_model_imv_time <- lm(imv_duration_hours_short ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_imv_time)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_imv_time_results <- data.frame(
  term = names(coef(secondary_model_imv_time)),
  estimate = coef(secondary_model_imv_time),
  std.error = summary(secondary_model_imv_time)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_imv_time)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_imv_time)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_imv_time_results, file.path(final_data, paste0(institution, "_secondary_analysis_imv_time.csv")))

```

#Secondary Analysis #2: 28-Day Ventilator-Free-Days
```{r}
english_spanish_primary_analysis$vfd_clean[english_spanish_primary_analysis$vfd_clean < 0] <- 0

secondary_model_vfd <- lm(vfd_clean ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_vfd)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_vfd_results <- data.frame(
  term = names(coef(secondary_model_vfd)),
  estimate = coef(secondary_model_vfd),
  std.error = summary(secondary_model_vfd)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_vfd)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_vfd)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_vfd_results, file.path(final_data, paste0(institution, "_secondary_analysis_vfd.csv")))

```


#Secondary Analysis #3: 28 Day ICU Free Days
```{r}
secondary_model_icufd <- lm(icufd_28 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_icufd)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_icufd_results <- data.frame(
  term = names(coef(secondary_model_icufd)),
  estimate = coef(secondary_model_icufd),
  std.error = summary(secondary_model_icufd)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_icufd)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_icufd)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_icufd_results, file.path(final_data, paste0(institution, "_secondary_analysis_icufd.csv")))

```


#Secondary Analysis #4: 30-Day Mortality
```{r}
secondary_model_30day_mortality <- glm(thirty_day_mortality ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_30day_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_30dm <- summary(secondary_model_30day_mortality)$coefficients
odds_ratios <- exp(coef_summary_30dm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_30dm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_30dm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_30day_mortality_results <- data.frame(
    Variable = rownames(coef_summary_30dm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_30day_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_30day_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_thirty_day_mortality.csv")))

```


#Secondary Analysis: In-Hospital Mortality
```{r}
secondary_model_inhospital_mortality <- glm(death_during_admission ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_inhospital_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_inhospitalm <- summary(secondary_model_inhospital_mortality)$coefficients
odds_ratios <- exp(coef_summary_inhospitalm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_inhospitalm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_inhospitalm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_inhospital_mortality_results <- data.frame(
    Variable = rownames(coef_summary_inhospitalm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_inhospital_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_inhospital_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_in_hospital_mortality.csv")))

```

#Secondary Analysis #5: Tracheostomy Odds
```{r}
secondary_model_trach <- glm(trached ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_trach)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_trach <- summary(secondary_model_trach)$coefficients
odds_ratios <- exp(coef_summary_trach[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_trach[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_trach[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_trach_results <- data.frame(
    Variable = rownames(coef_summary_trach),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_trach_results

# Write the results to a CSV file
fwrite(secondary_model_trach_results, file.path(final_data, paste0(institution, "_secondary_analysis_trach.csv")))

```

#Secondary Analysis #6: Reintubation Odds
```{r}
secondary_model_reintubation <- glm(reintubation ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_reintubation)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_reintubation <- summary(secondary_model_reintubation)$coefficients
odds_ratios <- exp(coef_summary_reintubation[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_reintubation[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_reintubation[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_reintubation_results <- data.frame(
    Variable = rownames(coef_summary_reintubation),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_reintubation_results

# Write the results to a CSV file
fwrite(secondary_model_reintubation_results, file.path(final_data, paste0(institution, "_secondary_analysis_reintubation.csv")))

```

#Secondary Analysis #7: Odds of Any Paralysis
```{r}
secondary_model_paralysis <- glm(paralytic_early_flag ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(secondary_model_paralysis)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(secondary_model_paralysis)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_results <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_results)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis.csv")))
```

#Secondary Analysis #8: 6-month hospital free days
```{r}
secondary_model_hfd_6m <- lm(hfd_6m ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_hfd_6m)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_hfd_6m_results <- data.frame(
  term = names(coef(secondary_model_hfd_6m)),
  estimate = coef(secondary_model_hfd_6m),
  std.error = summary(secondary_model_hfd_6m)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_hfd_6m)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_hfd_6m)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_hfd_6m_results, file.path(final_data, paste0(institution, "_secondary_analysis_hfd_6m.csv")))
```

#Secondary Analysis #9: 28-day Hospital free days
```{r}
secondary_model_hfd_28d <- lm(hfd_28d ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_hfd_28d)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_hfd_28d_results <- data.frame(
  term = names(coef(secondary_model_hfd_28d)),
  estimate = coef(secondary_model_hfd_28d),
  std.error = summary(secondary_model_hfd_28d)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_hfd_28d)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_hfd_28d)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_hfd_28d_results, file.path(final_data, paste0(institution, "_secondary_analysis_hfd_28d.csv")))

```

#Secondary Analysis #10: Odds of Push Paralytics
```{r}
secondary_model_push_paralytics <- glm(push_paralytics ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = no_continuous_paralytic_cohort, 
                                  family = binomial(link = "logit"))

summary(secondary_model_push_paralytics)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_push_paralytics <- summary(secondary_model_push_paralytics)$coefficients
odds_ratios <- exp(coef_summary_push_paralytics[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_push_paralytics[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_push_paralytics[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_push_paralytics_results <- data.frame(
    Variable = rownames(coef_summary_push_paralytics),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_push_paralytics_results

# Write the results to a CSV file
fwrite(secondary_model_push_paralytics_results, file.path(final_data, paste0(institution, "_secondary_analysis_push_paralytics.csv")))
```

#Secondary Analysis #11: Odds of Continuous Paralysis
```{r}
secondary_model_paralysis_cont <- glm(paralysis_cont ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_cont <- summary(secondary_model_paralysis_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis_cont[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_cont[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_cont[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_cont_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_cont),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_paralysis_cont_results

# Write the results to a CSV file
fwrite(secondary_model_paralysis_cont_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_cont.csv")))
```

#Secondary Analysis #11: MME
```{r}
secondary_model_mme <- lm(mme_median ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_mme)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_mme_results <- data.frame(
  term = names(coef(secondary_model_mme)),
  estimate = coef(secondary_model_mme),
  std.error = summary(secondary_model_mme)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_mme)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_mme)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_mme_results, file.path(final_data, paste0(institution, "_secondary_analysis_mme.csv")))
```

#Secondary Analysis #12: Benzos
```{r}
secondary_model_benzos <- lm(loraz_median ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_benzos)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_benzos_results <- data.frame(
  term = names(coef(secondary_model_benzos)),
  estimate = coef(secondary_model_benzos),
  std.error = summary(secondary_model_benzos)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_benzos)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_benzos)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_benzos_results, file.path(final_data, paste0(institution, "_secondary_analysis_benzos.csv")))
```

#Secondary Analysis #13: Propofol
```{r}
secondary_model_propofol <- lm(propofol_median ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = english_spanish_primary_analysis)

summary(secondary_model_propofol)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_propofol_results <- data.frame(
  term = names(coef(secondary_model_propofol)),
  estimate = coef(secondary_model_propofol),
  std.error = summary(secondary_model_propofol)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_propofol)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_propofol)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_propofol_results, file.path(final_data, paste0(institution, "_secondary_analysis_propofol.csv")))
```

#Secondary Analysis #14: Odds of Being Started on Benzos
```{r}
secondary_model_benzo_odds <- glm(benzos ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_benzo_odds)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_benzo_odds <- summary(secondary_model_benzo_odds)$coefficients
odds_ratios <- exp(coef_summary_benzo_odds[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_benzo_odds[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_benzo_odds[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_benzo_odds_results <- data.frame(
    Variable = rownames(coef_summary_benzo_odds),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)

# Write the results to a CSV file
fwrite(secondary_model_benzo_odds_results, file.path(final_data, paste0(institution, "_secondary_analysis_benzo_odds.csv")))
```

#Secondary Analysis #15: Opioid Only Analgesia
```{r}
secondary_model_opioid_only_analgesia <- glm(opioid_only_analgesia ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_opioid_only_analgesia)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_opioid_only_cont <- summary(secondary_model_opioid_only_analgesia)$coefficients
odds_ratios <- exp(coef_summary_opioid_only_cont[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_opioid_only_cont[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_opioid_only_cont[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_opioid_only_results <- data.frame(
    Variable = rownames(coef_summary_opioid_only_cont),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)

# Write the results to a CSV file
fwrite(secondary_model_opioid_only_results, file.path(final_data, paste0(institution, "_secondary_analysis_opioid_only.csv")))
```

#Secondary Analysis #16: Odds of Any Paralysis on Total Cohort non_eng_non_span vs span vs english
```{r}
secondary_model_paralysis_non_eng <- glm(paralytic_early_flag ~ non_eng_non_span + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = total_cohort, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_non_eng)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_non_eng <- summary(secondary_model_paralysis_non_eng)$coefficients
odds_ratios <- exp(coef_summary_paralysis_non_eng[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_non_eng[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_non_eng[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_non_eng_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_non_eng),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_paralysis_non_eng_results

# Write the results to a CSV file
fwrite(secondary_model_paralysis_non_eng_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_non_eng.csv")))
```

#Secondary Analysis #17: Odds of Any Paralysis Pre-Pandemic
```{r}
secondary_model_paralysis_pre_pandemic <- glm(paralytic_early_flag ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = pre_pandemic_cohort, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_pre_pandemic)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_pre_pandemic <- summary(secondary_model_paralysis_pre_pandemic)$coefficients
odds_ratios <- exp(coef_summary_paralysis_pre_pandemic[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_pre_pandemic[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_pre_pandemic[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_pre_pandemic_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_pre_pandemic),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_pre_pandemic_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_pre_pandemic.csv")))
```

#Secondary Analysis #18: Odds of Any Paralysis Pandemic and Beyond
```{r}
secondary_model_paralysis_pandemic <- glm(paralytic_early_flag ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = pandemic_cohort, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_pandemic)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_pandemic <- summary(secondary_model_paralysis_pandemic)$coefficients
odds_ratios <- exp(coef_summary_paralysis_pandemic[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_pandemic[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_pandemic[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_pandemic_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_pandemic),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_pandemic_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_pandemic.csv")))
```

#Secondary Analysis #19: Total time mechnically ventilated with covid
```{r}
secondary_model_imv_time_covid <- lm(imv_duration_hours_long ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status,
               data = english_spanish_primary_analysis)

summary(secondary_model_imv_time_covid)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_imv_time_covid_results <- data.frame(
  term = names(coef(secondary_model_imv_time_covid)),
  estimate = coef(secondary_model_imv_time_covid),
  std.error = summary(secondary_model_imv_time_covid)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_imv_time_covid)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_imv_time_covid)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_imv_time_covid_results, file.path(final_data, paste0(institution, "_secondary_analysis_imv_time_covid.csv")))

```

#STEP FIVE: SUBGROUP ANALYSES
#Subgroup Analysis #1: Hispanic Patients
```{r}
subgroup_model_hispanic <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = hispanic_cohort)

summary(subgroup_model_hispanic)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
subgroup_model_hispanic_results <- data.frame(
  term = names(coef(subgroup_model_hispanic)),
  estimate = coef(subgroup_model_hispanic),
  std.error = summary(subgroup_model_hispanic)$coefficients[, "Std. Error"],
  variance = (summary(subgroup_model_hispanic)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(subgroup_model_hispanic)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(subgroup_model_hispanic_results, file.path(final_data, paste0(institution, "_subgroup_analysis_hispanic.csv")))
```


#Subgroup Analysis #2: MICU Patients
```{r}

if(n_distinct(micu_only_cohort$hospital_icu_id) > 1) {
  
subgroup_model_micu <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = micu_only_cohort)
  
} else {

subgroup_model_micu <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + elixhauser_score + SVM_SCORE + laps2,
               data = micu_only_cohort)
  
}

summary(subgroup_model_micu)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
subgroup_model_micu_results <- data.frame(
  term = names(coef(subgroup_model_micu)),
  estimate = coef(subgroup_model_micu),
  std.error = summary(subgroup_model_micu)$coefficients[, "Std. Error"],
  variance = (summary(subgroup_model_micu)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(subgroup_model_micu)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(subgroup_model_micu_results, file.path(final_data, paste0(institution, "_subgroup_analysis_micu.csv")))

```


#Subgroup Analysis #3: COVID Pandemic Cohort
```{r}
subgroup_model_pandemic <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status,
               data = pandemic_cohort)

summary(subgroup_model_pandemic)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
subgroup_model_pandemic_results <- data.frame(
  term = names(coef(subgroup_model_pandemic)),
  estimate = coef(subgroup_model_pandemic),
  std.error = summary(subgroup_model_pandemic)$coefficients[, "Std. Error"],
  variance = (summary(subgroup_model_pandemic)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(subgroup_model_pandemic)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(subgroup_model_pandemic_results, file.path(final_data, paste0(institution, "_subgroup_analysis_pandemic.csv")))

```

#Subgroup Analysis #4: Non-Hispanic Patients
```{r}
subgroup_model_non_hispanic <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = non_hispanic_cohort)

summary(subgroup_model_non_hispanic)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
subgroup_model_non_hispanic_results <- data.frame(
  term = names(coef(subgroup_model_non_hispanic)),
  estimate = coef(subgroup_model_non_hispanic),
  std.error = summary(subgroup_model_non_hispanic)$coefficients[, "Std. Error"],
  variance = (summary(subgroup_model_non_hispanic)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(subgroup_model_non_hispanic)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(subgroup_model_non_hispanic_results, file.path(final_data, paste0(institution, "_subgroup_analysis_non_hispanic.csv")))
```

#Subgroup Analysis #5: Pre Pandemic Cohort
```{r}
subgroup_model_pre_pandemic <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2,
               data = pre_pandemic_cohort)

summary(subgroup_model_pre_pandemic)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
subgroup_model_pre_pandemic_results <- data.frame(
  term = names(coef(subgroup_model_pre_pandemic)),
  estimate = coef(subgroup_model_pre_pandemic),
  std.error = summary(subgroup_model_pre_pandemic)$coefficients[, "Std. Error"],
  variance = (summary(subgroup_model_pre_pandemic)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(subgroup_model_pre_pandemic)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(subgroup_model_pre_pandemic_results, file.path(final_data, paste0(institution, "_subgroup_analysis_pre_pandemic.csv")))

```

# Data Cleaning Summary
```{r}
data_cleaning_summary_analysis <- data.table(
  Stage = c(
    "Patients Removed for Speaking Language Other than English or Spanish", 
    "Patients Removed Because ICU never had Spanish speaking patient"
 ),
  Table_Name = c(
    "total_cohort - english_spanish_primary_all_icu_analysis", 
    "english_spanish_primary_all_icu_analysis - english_spanish_primary_analysis"
),
  Row_Count = c(
    nrow(total_cohort) - nrow(english_spanish_all_icu_analysis), 
    nrow(english_spanish_all_icu_analysis) - nrow(english_spanish_primary_analysis)
  )
)

fwrite(data_cleaning_summary_analysis, file.path(final_data, paste0(institution, "_analysis_data_cleaning_summary.csv")))
```




#PARALYTIC ANALYSES

#Whole Cohort
```{r}
paralysis_all_three_lang_cont_total_cohort <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = total_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_total_cohort)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_total_cohort)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_total_cohort <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_total_cohort)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_total_cohort, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_all_three_lang_total_cohort.csv")))


```

#AHRF All Paralysis
```{r}
#Acute Hypoxemic Respiratory Failure cohort
ahrf_cohort <- total_cohort[arHf_flag == TRUE]


ahrf_cohort$language_category <- relevel(as.factor(ahrf_cohort$language_category), ref = "english")
ahrf_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(ahrf_cohort$top_five_non_eng_non_spanish), ref = "english")
ahrf_cohort$non_eng <- relevel(as.factor(ahrf_cohort$non_eng), ref = "english")
ahrf_cohort$three_language_cat <- relevel(as.factor(ahrf_cohort$three_language_cat), ref = "english")
ahrf_cohort$sex_category <- relevel(as.factor(ahrf_cohort$sex_category), ref = "male")
ahrf_cohort$race_category <- relevel(as.factor(ahrf_cohort$race_category), ref = "white")
ahrf_cohort$ethnicity_category <- relevel(as.factor(ahrf_cohort$ethnicity_category), ref = "non-hispanic")
ahrf_cohort$covid_status <- relevel(as.factor(ahrf_cohort$covid_status), ref = "not detected")


paralysis_all_three_lang <- glm(paralytic_early_flag ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = ahrf_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_results <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_results)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_results, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_TOTAL_paralysis_all_three_lang_ahrf.csv")))

```


#AHRF Continuous Paralysis
```{r}
paralysis_all_three_lang_cont <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = ahrf_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_all_three_lang_ahrf.csv")))

```

#AHRF Push Paralytics
```{r}
paralysis_all_three_lang_push <- glm(push_paralytics ~ three_language_cat + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = ahrf_cohort, 
                                  family = binomial(link = "logit"))

summary(paralysis_all_three_lang_push)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_push_paralytics <- summary(paralysis_all_three_lang_push)$coefficients
odds_ratios <- exp(coef_summary_push_paralytics[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_push_paralytics[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_push_paralytics[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_push_paralytics_results <- data.frame(
    Variable = rownames(coef_summary_push_paralytics),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_push_paralytics_results)

# Write the results to a CSV file
fwrite(secondary_model_push_paralytics_results, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_PUSH_paralysis_all_three_lang_ahrf.csv")))

```



#AHRF Continuous Paralysis - Top 5 Languages
```{r}
paralysis_top_five_lang_cont <- glm(paralysis_cont ~ language_category + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = ahrf_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_top_five_lang_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_top_five_lang_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_top_five_lang_cont_results <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_top_five_lang_cont_results)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_top_five_lang_cont_results, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_top_five_lang_ahrf.csv")))

```

#ARF
```{r}
#Acute Hypoxemic Respiratory Failure cohort
arf_cohort <- total_cohort[arf_flag == TRUE]


arf_cohort$language_category <- relevel(as.factor(arf_cohort$language_category), ref = "english")
arf_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(arf_cohort$top_five_non_eng_non_spanish), ref = "english")
arf_cohort$non_eng <- relevel(as.factor(arf_cohort$non_eng), ref = "english")
arf_cohort$three_language_cat <- relevel(as.factor(arf_cohort$three_language_cat), ref = "english")
arf_cohort$sex_category <- relevel(as.factor(arf_cohort$sex_category), ref = "male")
arf_cohort$race_category <- relevel(as.factor(arf_cohort$race_category), ref = "white")
arf_cohort$ethnicity_category <- relevel(as.factor(arf_cohort$ethnicity_category), ref = "non-hispanic")
arf_cohort$covid_status <- relevel(as.factor(arf_cohort$covid_status), ref = "not detected")


paralysis_all_three_lang_arf <- glm(paralytic_early_flag ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = arf_cohort, 
                                  family = binomial(link = "logit"))

summary(paralysis_all_three_lang_arf)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_arf)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_arf_results <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_arf_results)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_arf_results, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_paralysis_all_three_lang_arf.csv")))

```


#ARF Continuous Paralysis
```{r}
paralysis_all_three_lang_cont <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = arf_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_arf <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_arf)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_arf, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_three_lang_lang_arf.csv")))

```

#Subgroup Analysis: COVID Pandemic Cohort
```{r}
paralysis_all_three_lang_cont_pandemic <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = pandemic_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_pandemic)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_pandemic)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_pandemic <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_pandemic)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_pandemic, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_three_lang_pandemic.csv")))

```

#Subgroup Analysis: Pre-Pandemic Cohort
```{r}
paralysis_all_three_lang_cont_prepandemic_unadjusted <- glm(paralysis_cont ~ three_language_cat, data = pre_pandemic_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_prepandemic_unadjusted)


paralysis_all_three_lang_cont_prepandemic <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = pre_pandemic_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_prepandemic)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_prepandemic)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_prepandemic <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_prepandemic)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_prepandemic, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_three_lang_prepandemic.csv")))

```


#Subgroup Analysis: Endemic Cohort
```{r}
paralysis_all_three_lang_cont_endemic_unadjusted <- glm(paralysis_cont ~ three_language_cat, data = endemic_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_endemic_unadjusted)


paralysis_all_three_lang_cont_endemic <- glm(paralysis_cont ~ three_language_cat + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + covid_status, data = pandemic_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_endemic)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_endemic)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_endemic <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_endemic)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_endemic, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_three_lang_endemic.csv")))

```


#Sample Figure for CRWIP
```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)

nmba_monthly <- total_cohort[
  !is.na(begin_imv) & !is.na(paralysis_cont) & !is.na(non_eng),
  .(
    n = .N,
    nmba_n = sum(paralysis_cont == 1, na.rm = TRUE),
    nmba_rate = mean(paralysis_cont == 1, na.rm = TRUE)
  ),
  by = .(
    month = floor_date(begin_imv, unit = "month"),
    non_eng
  )
]

nmba_quarterly <- total_cohort %>%
  mutate(
    quarter = floor_date(begin_imv, unit = "quarter")
  ) %>%
  group_by(quarter, non_eng) %>%
  summarise(
    n = n(),
    nmba_rate = mean(paralysis_cont == 1, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(nmba_quarterly, aes(x = quarter, y = nmba_rate, color = non_eng)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = as.Date("2020-03-01"), linetype = "dashed") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    x = "Quarter of IMV initiation",
    y = "Proportion receiving continuous NMBA within 48 hours",
    color = "Language group",
    title = "Quarterly early NMBA rates by language group"
  ) +
  theme_minimal(base_size = 14)


ggplot(nmba_monthly, aes(x = month, y = nmba_rate, color = non_eng)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(se = FALSE, linewidth = 1.2, method = "loess", span = 0.4) +
  geom_vline(
    xintercept = as.POSIXct("2020-03-01"),
    linetype = "dashed",
    linewidth = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = as.POSIXct("2020-03-01"),
    y = max(nmba_monthly$nmba_rate, na.rm = TRUE),
    label = "COVID onset",
    hjust = -0.1,
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Month",
    y = "NMBA within 48h",
    color = "Language",
    title = "NMBA trends by language group"
  ) +
  theme_minimal(base_size = 14)


#Pre-COVID

nmba_pre <- nmba_monthly %>%
  filter(month < as.Date("2020-03-01"))

ggplot(nmba_pre, aes(x = month, y = nmba_rate, color = non_eng)) +
  geom_point(alpha = 0.3, size = 1.5) +
  geom_smooth(se = FALSE, method = "loess", span = 0.4, linewidth = 1.2) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Month",
    y = "NMBA within 48h",
    color = "Language",
    title = "Pre-COVID NMBA trends by language group"
  ) +
  theme_minimal(base_size = 14)

```


#Sample Figure for CRWIP 2
```{r}
total_cohort[, covid_binary := fifelse(
  covid_status == "detected", 1, 0
)]

total_cohort <- total_cohort %>%
  mutate(
    month = lubridate::floor_date(begin_imv, "month"),
    time_num = as.numeric(month - min(month)) / 30  # approx months
  )

model_adj <- glm(
  paralysis_cont ~ non_eng + time_num + non_eng:time_num +
    race_category + ethnicity_category + age_category +
    bmi + sex_category + elixhauser_score + SVM_SCORE + laps2 +
    factor(hospital_icu_id) + covid_binary,
  data = total_cohort,
  family = binomial()
)

month_lookup <- total_cohort %>%
  distinct(time_num, month) %>%
  arrange(time_num)

newdata <- expand.grid(
  time_num = sort(unique(total_cohort$time_num)),
  non_eng = c("english", "non_english")
)

# Fill in other variables with typical values
newdata$age_category <- "65-74 Years"   # or most common
newdata$sex_category <- "male"
newdata$race_category <- "white"
newdata$ethnicity_category <- "non-hispanic"
newdata$covid_binary <- 0
newdata$bmi <- mean(total_cohort$bmi, na.rm = TRUE)
newdata$elixhauser_score <- mean(total_cohort$elixhauser_score, na.rm = TRUE)
newdata$SVM_SCORE <- mean(total_cohort$SVM_SCORE, na.rm = TRUE)
newdata$laps2 <- mean(total_cohort$laps2, na.rm = TRUE)
newdata$hospital_icu_id <- levels(factor(total_cohort$hospital_icu_id))[1]

newdata$pred <- predict(model_adj, newdata, type = "response")

newdata <- newdata %>%
  left_join(month_lookup, by = "time_num")


ggplot(newdata, aes(x = month, y = pred, color = non_eng)) +
  geom_line(linewidth = 1.2) +
  geom_vline(xintercept = as.Date("2020-03-01"), linetype = "dashed") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Month",
    y = "Adjusted probability of NMBA within 48h",
    color = "Language",
    title = "Adjusted NMBA use over time by language group"
  ) +
  theme_minimal(base_size = 14)

```



#Sample Figure for CRWIP 2
```{r}
arf_cohort[, covid_binary := fifelse(
  covid_status == "detected", 1, 0
)]

arf_cohort <- arf_cohort %>%
  mutate(
    month = lubridate::floor_date(begin_imv, "month"),
    time_num = as.numeric(month - min(month)) / 30  # approx months
  )

model_adj <- glm(
  paralysis_cont ~ non_eng + time_num + non_eng:time_num +
    race_category + ethnicity_category + age_category +
    bmi + sex_category + elixhauser_score + SVM_SCORE + laps2 +
    factor(hospital_icu_id) + covid_binary,
  data = arf_cohort,
  family = binomial()
)

month_lookup <- arf_cohort %>%
  distinct(time_num, month) %>%
  arrange(time_num)

newdata <- expand.grid(
  time_num = sort(unique(arf_cohort$time_num)),
  non_eng = c("english", "non_english")
)

# Fill in other variables with typical values
newdata$age_category <- "65-74 Years"   # or most common
newdata$sex_category <- "male"
newdata$race_category <- "white"
newdata$ethnicity_category <- "non-hispanic"
newdata$covid_binary <- 0
newdata$bmi <- mean(total_cohort$bmi, na.rm = TRUE)
newdata$elixhauser_score <- mean(total_cohort$elixhauser_score, na.rm = TRUE)
newdata$SVM_SCORE <- mean(total_cohort$SVM_SCORE, na.rm = TRUE)
newdata$laps2 <- mean(total_cohort$laps2, na.rm = TRUE)
newdata$hospital_icu_id <- levels(factor(total_cohort$hospital_icu_id))[1]

newdata$pred <- predict(model_adj, newdata, type = "response")

newdata <- newdata %>%
  left_join(month_lookup, by = "time_num")


ggplot(newdata, aes(x = month, y = pred, color = non_eng)) +
  geom_line(linewidth = 1.2) +
  geom_vline(xintercept = as.Date("2020-03-01"), linetype = "dashed") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Month",
    y = "Adjusted probability of NMBA within 48h",
    color = "Language",
    title = "Adjusted NMBA use over time by language group"
  ) +
  theme_minimal(base_size = 14)

```