# Project: Disparities in Sedation Practice by Patient Preferred Language
# Author: Alex Ortiz

# Setup Step - Setup libraries
```{r setup}
# Install Packages

packages <- c("zoo", "data.table", "dplyr", "lubridate", "tidyr", "fst", "arrow", "table1", "lme4", "metafor", "yaml", "pscl", "MASS", "htmltools", "car", "WeightIt", "MatchIt", "purrr", "survival", "mediation", "cmprsk", "gfoRmula", "nnet", "medflex", "bruceR", "tidyverse", "survey", "ipw", "geepack", "readr", "collapse", "stringr", "rprojroot", "tableone", "lmtest")
install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)

# Load necessary libraries
library(data.table)
library(dplyr)
library(zoo)  # for na.locf function
library(lubridate)
library(tidyr)
library(fst)
library(arrow)
library(table1)
library(metafor)
library(yaml)
library(htmltools)
library(MASS)
library(pscl)
library(car)
library(stringr)
library(rprojroot)
library(jsonlite)
library(tableone)
library(lmtest)


# Read the YAML from the config folder
project_root <- file.path( find_root(rprojroot::has_dir("deep_sedation_language_final")),
  "deep_sedation_language_final")
knitr::opts_knit$set(root.dir = file.path(project_root, "config"))
config <- yaml::read_yaml(file.path(project_root, "config", "deep_sedation_language_config.yaml"))

# Define directories based on YAML configuration
data_dir <- config$data_dir
output_dir <- config$output_dir
intermediate_dir <- file.path(output_dir, "intermediate_data_DO_NOT_UPLOAD/")  # Define output_dir based on data_dir
final_data_dir <- file.path(output_dir, "UPLOAD_ME/")
cohort_characteristics <- file.path(final_data_dir, "cohort_characteristics")
sedation_project <- file.path(final_data_dir, "sedation_project")
sedation_tables_dir <- file.path(sedation_project, "final_tables/")  
sedation_final_data <- file.path(sedation_project, "final_data/")
paralytics_project <- file.path(final_data_dir, "paralytics_project")
paralytics_tables_dir <- file.path(sedation_project, "final_tables/")  
paralytics_final_data <- file.path(sedation_project, "final_data/")
#second_dir <- file.path(output_dir, "secondary_project/")
file_type <- config$file_type
institution <- config$institution
time_zone <- config$time_zone

dirs_to_create <- c(intermediate_dir, final_data_dir, sedation_project, sedation_tables_dir, sedation_final_data, paralytics_project, paralytics_tables_dir, paralytics_final_data, cohort_characteristics)
for (dir in dirs_to_create) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)
  }
}

proj_end_date
```

#Upload Data
```{r}
read_data <- function(file_path, select = NULL) {
  if (grepl("\\.csv$|\\.csv\\.gz$", file_path)) {
    # Use fread with the select argument for CSV and CSV.GZ files
    return(fread(file_path, select = all_of(select)))
  } else if (grepl("\\.parquet$", file_path)) {
    # Use read_parquet with col_select for Parquet files
    return(arrow::read_parquet(file_path, col_select = all_of(select)))
  } else if (grepl("\\.fst$", file_path)) {
    # Use read_fst with columns for FST files
    return(fst::read_fst(file_path, columns = all_of(select)))
  } else {
    stop("Unsupported file format. Only CSV, Parquet, and FST are supported.")
  }
}

# Upload Analytic File
total_cohort <- read_parquet(paste0(intermediate_dir, "analytic_file", institution, ".parquet"))

total_cohort$icu_imv <- factor(
  total_cohort$icu_imv,
  levels = c("cardiac_icu", "general_icu", "medical_icu", 
             "mixed_cardiothoracic_icu", "mixed_neuro_icu", "neuro_icu", "neurosurgical_icu", "surgical_icu", "trauma_icu", "burn_icu", "cardiothoracic_surgical_icu"),
  labels = c("Cardiac ICU", "General ICU", "Medical ICU", 
             "Mixed Cardiothoracic ICU", "Mixed Neuro ICU", "Neuro ICU", "Neurosurgical ICU", "Surgical ICU", "Trauma ICU", "Burn ICU", "Cardiothoracic Surgical ICU")
)

catVars <- c("sex_category", "ethnicity_category", "race_category", "hospital_id", "icu_imv", "location_of_intubation")

```

#IPW Calculation
```{r}
ipw_outcome <- total_cohort[, .(
  reached_48h = as.integer(imv_duration_hours_final >= 48)
), by = hospitalization_id]

# Merge the reached_48h flag back to your dataset
ipw_sensitivity_analysis <- merge(total_cohort, ipw_outcome, by = "hospitalization_id")

# Use patient-level covariates to predict censoring
ipw_model <- glm(reached_48h ~ age_category + sex_category + race_category + ethnicity_category + laps2 + factor(hospital_icu_id) + elixhauser_score,
                 data = unique(ipw_sensitivity_analysis, by = "hospitalization_id"),
                 family = binomial())

total_cohort[, prob_reach_48h := predict(ipw_model, newdata = total_cohort, type = "response")]
total_cohort[, ipw := 1 / prob_reach_48h]
```

# Total Cohort for Non-English Analysis
```{r}
# Identify reference points for model
total_cohort$language_category <- relevel(as.factor(total_cohort$language_category), ref = "english")
total_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(total_cohort$top_five_non_eng_non_spanish), ref = "english")
total_cohort$non_eng <- relevel(as.factor(total_cohort$non_eng), ref = "english")
total_cohort$three_language_cat <- relevel(as.factor(total_cohort$three_language_cat), ref = "english")
total_cohort$sex_category <- relevel(as.factor(total_cohort$sex_category), ref = "male")
total_cohort$race_category <- relevel(as.factor(total_cohort$race_category), ref = "white")
total_cohort$ethnicity_category <- relevel(as.factor(total_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
total_cohort_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | non_eng, data = total_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(total_cohort_table_one), file.path(sedation_tables_dir, paste0(institution, "_total_cohort_table_one_manuscript.html")))

#Create paralytic_early_flag
total_cohort[, paralytic_early_flag := fifelse(
  (time_to_first_paralytic > 0) | (push_paralytics == 1),
  1, 0)]

setnafill(total_cohort, cols = "paralytic_early_flag", fill = 0)

#Create Continuous Paralytic Flag
total_cohort[, paralysis_cont := fifelse(
  time_to_first_paralytic > 0 & time_to_first_paralytic <= 48,
  1L, 0L
)]
setnafill(total_cohort, cols = "paralysis_cont", fill = 0)

total_cohort[, benzos := fifelse(!is.na(loraz_median), 1, 0)]
```

# Cohort for Top 5 non-English, non-Spanish Language
```{r}
top_five_cohort <- total_cohort[top_five_non_eng_non_spanish %in% c("english", "top_five_non_eng_non_spanish", "spanish")]
top_five_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(top_five_cohort$top_five_non_eng_non_spanish), ref = "english")
top_five_cohort$non_eng <- relevel(as.factor(top_five_cohort$non_eng), ref = "english")
top_five_cohort$sex_category <- relevel(as.factor(top_five_cohort$sex_category), ref = "male")
top_five_cohort$race_category <- relevel(as.factor(top_five_cohort$race_category), ref = "white")
top_five_cohort$ethnicity_category <- relevel(as.factor(top_five_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
# Generate the summary table grouped by language
top_five_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | top_five_non_eng_non_spanish, data = top_five_cohort, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(top_five_table_one), file.path(sedation_tables_dir, paste0(institution, "_top_five_cohort_table_one_manuscript.html")))
```

# Cohort for Sensitivity Analysis Involving All Spanish/English Pts regardless of ICU of admission
```{r}
english_spanish_all_icu_analysis <- total_cohort[language_category %in% c("english", "spanish")]
english_spanish_all_icu_analysis[, race_category := as.character(race_category)]
english_spanish_all_icu_analysis[, race_category := fifelse(
  race_category %in% c("asian"),
  "other/unknown",
  race_category
)]

# Identify reference points for model
english_spanish_all_icu_analysis$language_category <- relevel(as.factor(english_spanish_all_icu_analysis$language_category), ref = "english")
english_spanish_all_icu_analysis$sex_category <- relevel(as.factor(english_spanish_all_icu_analysis$sex_category), ref = "male")
english_spanish_all_icu_analysis$race_category <- relevel(as.factor(english_spanish_all_icu_analysis$race_category), ref = "white")
english_spanish_all_icu_analysis$ethnicity_category <- relevel(as.factor(english_spanish_all_icu_analysis$ethnicity_category), ref = "non-hispanic")


#Basic Table One
# Generate the summary table grouped by language
eng_span_all_icu_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = english_spanish_all_icu_analysis, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(eng_span_all_icu_table_one), file.path(sedation_tables_dir, paste0(institution, "_eng_span_all_icu_cohort_table_one_manuscript.html")))

```

# Primary Cohort for Analysis
```{r}
# Step 1: Count Spanish-speaking patients by hospital_icu_id
spanish_counts <- english_spanish_all_icu_analysis[
  language_category == "spanish", 
  .N, 
  by = hospital_icu_id
]

# Step 2: Identify hospital_icu_id values with 0 Spanish-speaking patients
# (We do this by first getting all unique hospital_icu_id values and checking which ones are missing)
all_hospitals <- unique(english_spanish_all_icu_analysis$hospital_icu_id)
hospitals_with_spanish <- spanish_counts$hospital_icu_id
hospitals_to_remove <- setdiff(all_hospitals, hospitals_with_spanish)

# Step 3: Filter out those hospitals from the original table
english_spanish_primary_analysis <- english_spanish_all_icu_analysis[
  !hospital_icu_id %in% hospitals_to_remove
]

english_spanish_primary_analysis <- english_spanish_primary_analysis %>%
  mutate(language_category = droplevels(language_category))

# Identify reference points for model
english_spanish_primary_analysis$language_category <- relevel(as.factor(english_spanish_primary_analysis$language_category), ref = "english")
english_spanish_primary_analysis$sex_category <- relevel(as.factor(english_spanish_primary_analysis$sex_category), ref = "male")
english_spanish_primary_analysis$race_category <- relevel(as.factor(english_spanish_primary_analysis$race_category), ref = "white")
english_spanish_primary_analysis$ethnicity_category <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category), ref = "non-hispanic")
#english_spanish_primary_analysis$covid_status <- relevel(as.factor(english_spanish_primary_analysis$covid_status), ref = "not detected")
english_spanish_primary_analysis$ethnicity_category_final <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")
english_spanish_primary_analysis$push_paralytics <- relevel(as.factor(english_spanish_primary_analysis$push_paralytics), ref = "0")


# Create labels for readability
label(english_spanish_primary_analysis$imv_duration_hours) <- "<u>Hours Mechanically Ventilated</u>"
label(english_spanish_primary_analysis$race_category) <- "<u>Race</u>"
label(english_spanish_primary_analysis$ethnicity_category) <- "<u>Ethnicity</u>"
label(english_spanish_primary_analysis$sex_category) <- "<u>Sex</u>"
label(english_spanish_primary_analysis$age_at_admission) <- "<u>Age</u>"
label(english_spanish_primary_analysis$language_category) <- "Language"
label(english_spanish_primary_analysis$bmi) <- "<u>BMI</u>"
label(english_spanish_primary_analysis$hospital_id) <- "<u>Hospital</u>"
label(english_spanish_primary_analysis$icu_imv) <- "<u>ICU Type</u>"
label(english_spanish_primary_analysis$mortality_28day) <- "<u>Twenty-Eight Day Mortality</u>"
label(english_spanish_primary_analysis$location_of_intubation) <- "<u>Location of Intubation</u>"
label(english_spanish_primary_analysis$elixhauser_score) <- "<u>Elixhauser Score</u>"
label(english_spanish_primary_analysis$SVM_SCORE) <- "<u>Social Vulnerability Metric</u>"
label(english_spanish_primary_analysis$laps2) <- "<u>LAPS2 Score</u>"
label(english_spanish_primary_analysis$time_to_first_rass) <- "<u>Time to First RASS</u>"
label(english_spanish_primary_analysis$rass_frequency_hrs) <- "<u>Frequency Of Rass Measurements</u>"
#label(english_spanish_primary_analysis$covid_status) <- "<u>COVID Status</u>"

# Generate the summary table grouped by language
summary_table <- table1(
  ~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + location_of_intubation + elixhauser_score + SVM_SCORE + time_to_first_rass + rass_frequency_hrs 
  | language_category, 
  data = english_spanish_primary_analysis,
  factorVars = catVars,
  test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(summary_table), file.path(sedation_tables_dir, paste0(institution, "_primary_cohort_table_one_manuscript.html")))

vars <- c("age_at_admission", "bmi", "sex_category", "race_category", "ethnicity_category", "laps2", "hospital_id", "icu_imv", "location_of_intubation", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs")
catVars <- c("sex_category", "race_category", "ethnicity_category", "hospital_id", "icu_imv", "location_of_intubation")

# Create table
table1_obj <- CreateTableOne(vars = vars, strata = "language_category",
                             data = english_spanish_primary_analysis,
                             factorVars = catVars)

# Print with SMD and nonparametric summary
print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi"))

print(table1_obj, smd = TRUE)

# Convert the table1 object to a data frame
table1_df <- print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi", "laps2", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs"), printToggle = FALSE)

# Save to CSV
# Save the combined results to a file
fwrite(table1_df, file.path(sedation_tables_dir, paste0(institution, "_IQR_SMD_Table1.csv")))
```

#STEP TWO: PRIMARY ANALYSIS
```{r}
# Fit the unadjusted primary model
primary_model_unadjusted <- lm(percent_deep_sedation_48 ~ language_category, data = english_spanish_primary_analysis)

summary(primary_model_unadjusted)

# Extract coefficients, standard errors, variance, and p-values for the unadjusted model
primary_results_unadjusted <- data.frame(
  term = names(coef(primary_model_unadjusted)),
  estimate = coef(primary_model_unadjusted),
  std.error = summary(primary_model_unadjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_unadjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_unadjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

primary_model_adjusted <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(primary_model_adjusted)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
primary_results_adjusted <- data.frame(
  term = names(coef(primary_model_adjusted)),
  estimate = coef(primary_model_adjusted),
  std.error = summary(primary_model_adjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_adjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_adjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Combine the results into one data frame
primary_results_combined <- bind_rows(
  primary_results_unadjusted %>% mutate(model = "Unadjusted"),
  primary_results_adjusted %>% mutate(model = "Adjusted")
)

# Save the combined results to a file
fwrite(primary_results_combined, file.path(final_data, paste0(institution, "_primary_analysis_results.csv")))
```


#STEP FOUR: SECONDARY ANALYSES
#Secondary Analysis: Total time mechnically ventilated
```{r}
secondary_model_imv_time <- lm(imv_duration_hours_final ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_imv_time)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_imv_time_results <- data.frame(
  term = names(coef(secondary_model_imv_time)),
  estimate = coef(secondary_model_imv_time),
  std.error = summary(secondary_model_imv_time)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_imv_time)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_imv_time)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_imv_time_results, file.path(final_data, paste0(institution, "_secondary_analysis_imv_time.csv")))

```


#Secondary Analysis: 28-Day Ventilator-Free-Days
```{r}

secondary_model_vfd <- lm(vent_free_days_28 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_vfd)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_vfd_results <- data.frame(
  term = names(coef(secondary_model_vfd)),
  estimate = coef(secondary_model_vfd),
  std.error = summary(secondary_model_vfd)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_vfd)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_vfd)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_vfd_results, file.path(final_data, paste0(institution, "_secondary_analysis_vfd.csv")))

```


#Secondary Analysis: 28-Day Mortality
```{r}
secondary_model_28day_mortality <- glm(mortality_28day ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_28day_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_30dm <- summary(secondary_model_28day_mortality)$coefficients
odds_ratios <- exp(coef_summary_30dm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_30dm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_30dm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_30day_mortality_results <- data.frame(
    Variable = rownames(coef_summary_30dm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_30day_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_30day_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_28_day_mortality.csv")))

```



#Secondary Analysis: In-Hospital Mortality
```{r}
secondary_model_inhospital_mortality <- glm(death_during_admission ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_inhospital_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_inhospitalm <- summary(secondary_model_inhospital_mortality)$coefficients
odds_ratios <- exp(coef_summary_inhospitalm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_inhospitalm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_inhospitalm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_inhospital_mortality_results <- data.frame(
    Variable = rownames(coef_summary_inhospitalm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_inhospital_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_inhospital_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_in_hospital_mortality.csv")))

```


#Secondary Analysis: Tracheostomy Odds
```{r}
secondary_model_trach <- glm(trached ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_trach)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_trach <- summary(secondary_model_trach)$coefficients
odds_ratios <- exp(coef_summary_trach[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_trach[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_trach[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_trach_results <- data.frame(
    Variable = rownames(coef_summary_trach),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_trach_results

# Write the results to a CSV file
fwrite(secondary_model_trach_results, file.path(final_data, paste0(institution, "_secondary_analysis_trach.csv")))

```


#Secondary Analysis: Reintubation Odds
```{r}
secondary_model_reintubation <- glm(reintubation_final ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_reintubation)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_reintubation <- summary(secondary_model_reintubation)$coefficients
odds_ratios <- exp(coef_summary_reintubation[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_reintubation[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_reintubation[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_reintubation_results <- data.frame(
    Variable = rownames(coef_summary_reintubation),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_reintubation_results

# Write the results to a CSV file
fwrite(secondary_model_reintubation_results, file.path(final_data, paste0(institution, "_secondary_analysis_reintubation.csv")))

```

#PARALYTIC ANALYSES

#Whole Cohort
```{r}
paralysis_all_three_lang_cont_total_cohort <- glm(paralysis_cont ~ language_category + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = total_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_total_cohort)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_total_cohort)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_total_cohort <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_total_cohort)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_total_cohort, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_all_three_lang_total_cohort.csv")))


```

#Secondary Analysis: Odds of Starting Continuous Paralysis in the first 48 hours
```{r}
secondary_model_paralysis_cont <- glm(paralysis_cont ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_cont <- summary(secondary_model_paralysis_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis_cont[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_cont[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_cont[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_cont_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_cont),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_paralysis_cont_results

# Write the results to a CSV file
fwrite(secondary_model_paralysis_cont_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_cont.csv")))
```

