# Project: Disparities in Sedation Practice by Patient Preferred Language
# Author: Alex Ortiz

# Setup Step - Setup libraries
```{r setup}
# Install Packages

packages <- c("zoo", "data.table", "dplyr", "lubridate", "tidyr", "fst", "arrow", "table1", "lme4", "metafor", "yaml", "pscl", "MASS", "htmltools", "car", "WeightIt", "MatchIt", "purrr", "survival", "mediation", "cmprsk", "gfoRmula", "nnet", "medflex", "bruceR", "tidyverse", "survey", "ipw", "geepack", "readr", "collapse", "stringr", "rprojroot", "tableone", "lmtest")
install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)

# Load necessary libraries
library(data.table)
library(dplyr)
library(zoo)  # for na.locf function
library(lubridate)
library(tidyr)
library(fst)
library(arrow)
library(table1)
library(metafor)
library(yaml)
library(htmltools)
library(MASS)
library(pscl)
library(car)
library(stringr)
library(rprojroot)
library(jsonlite)
library(tableone)
library(lmtest)

# Read the YAML from the config folder
project_root <- file.path( find_root(rprojroot::has_dir("deep_sedation_language_final")),
  "deep_sedation_language_final")
knitr::opts_knit$set(root.dir = file.path(project_root, "config"))
config <- yaml::read_yaml(file.path(project_root, "config", "deep_sedation_language_config.yaml"))

# Define directories based on YAML configuration
data_dir <- config$data_dir
local_dir <- config$local_dir
output_dir <- file.path(local_dir, "output/")
intermediate_dir <- file.path(local_dir, "intermediate_data_DO_NOT_UPLOAD/")  # Define output_dir based on data_dir
final_data_dir <- file.path(output_dir, "FINAL_DATA_UPLOAD_ME/")
cohort_characteristics <- file.path(final_data_dir, "cohort_characteristics/")
sedation_project <- file.path(final_data_dir, "sedation_project/")
sedation_tables_dir <- file.path(sedation_project, "final_tables/")
sedation_final_data <- file.path(sedation_project, "final_data/")
paralytics_project <- file.path(final_data_dir, "paralytics_project/")
paralytics_tables_dir <- file.path(sedation_project, "final_tables/")  
paralytics_final_data <- file.path(sedation_project, "final_data/")
file_type <- config$file_type
institution <- config$institution
time_zone <- config$time_zone

dirs_to_create <- c(output_dir, intermediate_dir, final_data_dir, sedation_project, sedation_tables_dir, sedation_final_data, paralytics_project, paralytics_tables_dir, paralytics_final_data, cohort_characteristics)
for (dir in dirs_to_create) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)
  }
}
```

#Upload Data
```{r}
# Upload Analytic File
total_cohort <- read_parquet(paste0(intermediate_dir, "analytic_file", institution, ".parquet"))
setDT(total_cohort)

# Upload hospital system language information
language_data <- fread(paste0(cohort_characteristics, institution, "_hospital_system_language_counts", ".csv"))
setDT(language_data)

cat_vars <- c("sex_category", "ethnicity_category", "race_category", "hospital_id", "icu_imv", "location_of_intubation", "admission_type_category", "arf_flag", "arHf_flag")

cont_vars <- c("age_at_admission", "bmi", "elixhauser_score", "SVM_SCORE", "laps2", "time_to_first_rass", "rass_frequency_hrs", "observation_year")

```

#Function to Create Table 1s
```{r}
#Table 1: Descriptive
make_table1_federated <- function(dt,
                                  cat_vars = NULL,
                                  cont_vars = NULL,
                                  by = NULL,
                                  site_name = NA_character_) {
  dt <- as.data.table(copy(dt))
  out_list <- list()
  
  #-----------------------------#
  # CATEGORICAL VARIABLES
  #-----------------------------#
  if (!is.null(cat_vars) && length(cat_vars) > 0) {
    for (v in cat_vars) {
      
      if (!is.null(by)) {
        tmp <- dt[, .(n = .N), by = c(by, v)]
        
        totals <- dt[, .(N = .N), by = by]
        
        tmp <- merge(tmp, totals, by = by, all.x = TRUE)
        
        setnames(tmp, old = by, new = "group")
        setnames(tmp, old = v, new = "level")
        
        tmp[, `:=`(
          site = site_name,
          variable = v,
          variable_type = "categorical",
          level = as.character(level),
          pct = 100 * n / N,
          mean = NA_real_,
          sd = NA_real_,
          median = NA_real_,
          p25 = NA_real_,
          p75 = NA_real_,
          min = NA_real_,
          max = NA_real_
        )]
        
      } else {
        tmp <- dt[, .(n = .N), by = v]
        total_n <- nrow(dt)
        
        setnames(tmp, old = v, new = "level")
        
        tmp[, `:=`(
          site = site_name,
          group = "overall",
          variable = v,
          variable_type = "categorical",
          level = as.character(level),
          N = total_n,
          pct = 100 * n / N,
          mean = NA_real_,
          sd = NA_real_,
          median = NA_real_,
          p25 = NA_real_,
          p75 = NA_real_,
          min = NA_real_,
          max = NA_real_
        )]
      }
      
      tmp <- tmp[, .(
        site, group, variable, variable_type, level,
        n, N, pct,
        mean, sd, median, p25, p75, min, max
      )]
      
      out_list[[paste0("cat_", v)]] <- tmp
    }
  }
  
  #-----------------------------#
  # CONTINUOUS VARIABLES
  #-----------------------------#
  if (!is.null(cont_vars) && length(cont_vars) > 0) {
    for (v in cont_vars) {
      
      if (!is.null(by)) {
        tmp <- dt[, .(
          n = sum(!is.na(get(v))),
          N = .N,
          pct = NA_real_,
          mean = if (all(is.na(get(v)))) NA_real_ else mean(get(v), na.rm = TRUE),
          sd = if (sum(!is.na(get(v))) <= 1) NA_real_ else sd(get(v), na.rm = TRUE),
          median = if (all(is.na(get(v)))) NA_real_ else median(get(v), na.rm = TRUE),
          p25 = if (all(is.na(get(v)))) NA_real_ else as.numeric(quantile(get(v), 0.25, na.rm = TRUE)),
          p75 = if (all(is.na(get(v)))) NA_real_ else as.numeric(quantile(get(v), 0.75, na.rm = TRUE)),
          min = if (all(is.na(get(v)))) NA_real_ else min(get(v), na.rm = TRUE),
          max = if (all(is.na(get(v)))) NA_real_ else max(get(v), na.rm = TRUE)
        ), by = by]
        
        setnames(tmp, old = by, new = "group")
        
        tmp[, `:=`(
          site = site_name,
          variable = v,
          variable_type = "continuous",
          level = NA_character_
        )]
        
      } else {
        tmp <- dt[, .(
          n = sum(!is.na(get(v))),
          N = .N,
          pct = NA_real_,
          mean = if (all(is.na(get(v)))) NA_real_ else mean(get(v), na.rm = TRUE),
          sd = if (sum(!is.na(get(v))) <= 1) NA_real_ else sd(get(v), na.rm = TRUE),
          median = if (all(is.na(get(v)))) NA_real_ else median(get(v), na.rm = TRUE),
          p25 = if (all(is.na(get(v)))) NA_real_ else as.numeric(quantile(get(v), 0.25, na.rm = TRUE)),
          p75 = if (all(is.na(get(v)))) NA_real_ else as.numeric(quantile(get(v), 0.75, na.rm = TRUE)),
          min = if (all(is.na(get(v)))) NA_real_ else min(get(v), na.rm = TRUE),
          max = if (all(is.na(get(v)))) NA_real_ else max(get(v), na.rm = TRUE)
        )]
        
        tmp[, `:=`(
          site = site_name,
          group = "overall",
          variable = v,
          variable_type = "continuous",
          level = NA_character_
        )]
      }
      
      tmp <- tmp[, .(
        site, group, variable, variable_type, level,
        n, N, pct,
        mean, sd, median, p25, p75, min, max
      )]
      
      out_list[[paste0("cont_", v)]] <- tmp
    }
  }
  
  out <- rbindlist(out_list, fill = TRUE, use.names = TRUE)
  
  setcolorder(out, c(
    "site", "group", "variable", "variable_type", "level",
    "n", "N", "pct", "mean", "sd", "median", "p25", "p75", "min", "max"
  ))
  
  return(out[])
}

# Table 1: Comparative

make_table1_comparisons <- function(dt,
                                    by,
                                    cont_vars = NULL,
                                    cat_vars = NULL,
                                    site_name = NA_character_,
                                    group_levels = NULL,
                                    fisher_if_expected_lt5 = TRUE) {
  dt <- as.data.table(copy(dt))
  
  if (!(by %in% names(dt))) {
    stop(sprintf("Grouping variable '%s' not found in dt.", by))
  }
  
  # Keep only non-missing comparison groups
  dt <- dt[!is.na(get(by))]
  
  # Force grouping variable to character first
  dt[, (by) := as.character(get(by))]
  
  # Determine the two groups
  observed_groups <- unique(dt[[by]])
  observed_groups <- observed_groups[!is.na(observed_groups)]
  
  if (!is.null(group_levels)) {
    if (length(group_levels) != 2) {
      stop("group_levels must have length 2.")
    }
    if (!all(group_levels %in% observed_groups)) {
      stop("Not all specified group_levels are present in the data.")
    }
    groups <- group_levels
  } else {
    if (length(observed_groups) != 2) {
      stop(sprintf(
        "Grouping variable '%s' must have exactly 2 non-missing groups. Found: %s",
        by,
        paste(observed_groups, collapse = ", ")
      ))
    }
    groups <- sort(observed_groups)
  }
  
  g1 <- groups[1]
  g2 <- groups[2]
  comparison_label <- paste0(by, ": ", g2, " vs ", g1)
  
  out_list <- list()
  
  #--------------------------------------------------#
  # Helper: continuous variable comparison
  #--------------------------------------------------#
  compare_continuous <- function(x, group, var_name) {
    dsub <- data.table(x = x, group = group)
    dsub <- dsub[!is.na(group)]
    
    x1 <- dsub[group == g1, x]
    x2 <- dsub[group == g2, x]
    
    x1_nonmiss <- x1[!is.na(x1)]
    x2_nonmiss <- x2[!is.na(x2)]
    
    n1 <- length(x1_nonmiss)
    n2 <- length(x2_nonmiss)
    
    mean1 <- if (n1 > 0) mean(x1_nonmiss) else NA_real_
    mean2 <- if (n2 > 0) mean(x2_nonmiss) else NA_real_
    sd1   <- if (n1 > 1) sd(x1_nonmiss) else NA_real_
    sd2   <- if (n2 > 1) sd(x2_nonmiss) else NA_real_
    med1  <- if (n1 > 0) median(x1_nonmiss) else NA_real_
    med2  <- if (n2 > 0) median(x2_nonmiss) else NA_real_
    p25_1 <- if (n1 > 0) as.numeric(quantile(x1_nonmiss, 0.25, na.rm = TRUE)) else NA_real_
    p25_2 <- if (n2 > 0) as.numeric(quantile(x2_nonmiss, 0.25, na.rm = TRUE)) else NA_real_
    p75_1 <- if (n1 > 0) as.numeric(quantile(x1_nonmiss, 0.75, na.rm = TRUE)) else NA_real_
    p75_2 <- if (n2 > 0) as.numeric(quantile(x2_nonmiss, 0.75, na.rm = TRUE)) else NA_real_
    
    # Mean difference and CI from Welch t-test
    estimate <- NA_real_
    ci_lower <- NA_real_
    ci_upper <- NA_real_
    p_value  <- NA_real_
    test_used <- NA_character_
    
    if (n1 >= 2 && n2 >= 2) {
      tt <- try(t.test(x2_nonmiss, x1_nonmiss), silent = TRUE)
      if (!inherits(tt, "try-error")) {
        estimate <- unname(mean2 - mean1)
        ci_lower <- unname(tt$conf.int[1])
        ci_upper <- unname(tt$conf.int[2])
        p_value  <- tt$p.value
        test_used <- "welch_t_test"
      }
    }
    
    # SMD = difference in means / pooled SD
    pooled_sd <- NA_real_
    smd <- NA_real_
    if (n1 >= 2 && n2 >= 2 && !is.na(sd1) && !is.na(sd2)) {
      pooled_sd <- sqrt(((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2) / (n1 + n2 - 2))
      if (!is.na(pooled_sd) && pooled_sd > 0) {
        smd <- (mean2 - mean1) / pooled_sd
      }
    }
    
    data.table(
      site = site_name,
      variable = var_name,
      variable_type = "continuous",
      comparison = comparison_label,
      level = NA_character_,
      group1 = g1,
      group2 = g2,
      group1_n = n1,
      group2_n = n2,
      group1_mean = mean1,
      group2_mean = mean2,
      group1_sd = sd1,
      group2_sd = sd2,
      group1_median = med1,
      group2_median = med2,
      group1_p25 = p25_1,
      group2_p25 = p25_2,
      group1_p75 = p75_1,
      group2_p75 = p75_2,
      statistic = "mean_difference",
      estimate = estimate,
      ci_lower = ci_lower,
      ci_upper = ci_upper,
      p_value = p_value,
      smd = smd,
      effect_measure = "mean_difference",
      test = test_used,
      note = NA_character_
    )
  }
  
  #--------------------------------------------------#
  # Helper: multilevel categorical SMD
  #--------------------------------------------------#
  calc_multilevel_smd <- function(x, group, g1, g2) {
    dsub <- data.table(
      x = as.character(x),
      group = as.character(group)
    )
    
    dsub <- dsub[!is.na(group) & group %in% c(g1, g2)]
    dsub[is.na(x), x := "missing"]
    
    levs <- sort(unique(dsub$x))
    
    # Need at least 2 levels
    if (length(levs) < 2) return(NA_real_)
    
    # Use first level as reference, keep remaining K-1 levels
    ref <- levs[1]
    keep_levs <- levs[levs != ref]
    
    make_indicator_matrix <- function(vals, levels_keep) {
      out <- sapply(levels_keep, function(lv) as.numeric(vals == lv))
      if (is.null(dim(out))) {
        out <- matrix(out, ncol = 1)
        colnames(out) <- levels_keep
      }
      out
    }
    
    x1 <- dsub[group == g1, x]
    x2 <- dsub[group == g2, x]
    
    if (length(x1) < 2 || length(x2) < 2) return(NA_real_)
    
    M1 <- make_indicator_matrix(x1, keep_levs)
    M2 <- make_indicator_matrix(x2, keep_levs)
    
    p1 <- colMeans(M1)
    p2 <- colMeans(M2)
    d  <- p2 - p1
    
    S1 <- stats::cov(M1)
    S2 <- stats::cov(M2)
    
    if (is.null(dim(S1))) S1 <- matrix(S1, nrow = 1, ncol = 1)
    if (is.null(dim(S2))) S2 <- matrix(S2, nrow = 1, ncol = 1)
    
    S_pooled <- (S1 + S2) / 2
    
    # Try regular inverse first; generalized inverse as fallback
    S_inv <- try(solve(S_pooled), silent = TRUE)
    if (inherits(S_inv, "try-error")) {
      if (!requireNamespace("MASS", quietly = TRUE)) return(NA_real_)
      S_inv <- try(MASS::ginv(S_pooled), silent = TRUE)
      if (inherits(S_inv, "try-error")) return(NA_real_)
    }
    
    smd <- as.numeric(sqrt(t(d) %*% S_inv %*% d))
    smd
  }
  
  #--------------------------------------------------#
  # Helper: categorical variable comparison
  #--------------------------------------------------#
  compare_categorical <- function(x, group, var_name) {
    dsub <- data.table(x = x, group = group)
    dsub <- dsub[!is.na(group)]
    
    # treat missing category as its own level if present
    dsub[, x := as.character(x)]
    dsub[is.na(x), x := "missing"]
    
    tab <- table(dsub$x, dsub$group)
    
    # Ensure both groups are present as columns in correct order
    needed_cols <- c(g1, g2)
    if (!all(needed_cols %in% colnames(tab))) {
      full_tab <- matrix(
        0,
        nrow = nrow(tab),
        ncol = 2,
        dimnames = list(rownames(tab), needed_cols)
      )
      common_cols <- intersect(colnames(tab), needed_cols)
      full_tab[, common_cols] <- tab[, common_cols, drop = FALSE]
      tab <- full_tab
    } else {
      tab <- tab[, needed_cols, drop = FALSE]
    }
    
    n_levels <- nrow(tab)
    n1 <- sum(tab[, g1])
    n2 <- sum(tab[, g2])
    
    overall_smd <- calc_multilevel_smd(
      x = dsub$x,
      group = dsub$group,
      g1 = g1,
      g2 = g2
    )
    
    p_value <- NA_real_
    test_used <- NA_character_
    
    if (n_levels >= 1) {
      chisq_try <- try(chisq.test(tab, correct = FALSE), silent = TRUE)
      
      if (!inherits(chisq_try, "try-error")) {
        expected <- chisq_try$expected
        
        if (fisher_if_expected_lt5 && any(expected < 5)) {
          fisher_try <- try(fisher.test(tab), silent = TRUE)
          if (!inherits(fisher_try, "try-error")) {
            p_value <- fisher_try$p.value
            test_used <- "fisher_exact"
          } else {
            p_value <- chisq_try$p.value
            test_used <- "chi_square"
          }
        } else {
          p_value <- chisq_try$p.value
          test_used <- "chi_square"
        }
      }
    }
    
    # Overall categorical row for all categorical variables
    out <- data.table(
      site = site_name,
      variable = var_name,
      variable_type = "categorical",
      comparison = comparison_label,
      level = NA_character_,
      group1 = g1,
      group2 = g2,
      group1_n = n1,
      group2_n = n2,
      group1_mean = NA_real_,
      group2_mean = NA_real_,
      group1_sd = NA_real_,
      group2_sd = NA_real_,
      group1_median = NA_real_,
      group2_median = NA_real_,
      group1_p25 = NA_real_,
      group2_p25 = NA_real_,
      group1_p75 = NA_real_,
      group2_p75 = NA_real_,
      statistic = "overall_difference",
      estimate = NA_real_,
      ci_lower = NA_real_,
      ci_upper = NA_real_,
      p_value = p_value,
      smd = overall_smd,
      effect_measure = "categorical_smd",
      test = test_used,
      note = NA_character_
    )
    
    # If binary categorical, also compute risk difference and binary SMD
    if (n_levels == 2) {
      levels_x <- rownames(tab)
      ref_level <- levels_x[1]
      target_level <- levels_x[2]
      
      a <- as.numeric(tab[target_level, g2])  # target in group2
      b <- as.numeric(tab[ref_level,   g2])   # ref in group2
      c <- as.numeric(tab[target_level, g1])  # target in group1
      d <- as.numeric(tab[ref_level,   g1])   # ref in group1
      
      p1 <- if ((c + d) > 0) c / (c + d) else NA_real_
      p2 <- if ((a + b) > 0) a / (a + b) else NA_real_
      
      # Risk difference
      rd <- if (!is.na(p1) && !is.na(p2)) p2 - p1 else NA_real_
      rd_se <- if (!is.na(p1) && !is.na(p2) && (c + d) > 0 && (a + b) > 0) {
        sqrt((p1 * (1 - p1) / (c + d)) + (p2 * (1 - p2) / (a + b)))
      } else {
        NA_real_
      }
      rd_lcl <- if (!is.na(rd_se)) rd - 1.96 * rd_se else NA_real_
      rd_ucl <- if (!is.na(rd_se)) rd + 1.96 * rd_se else NA_real_
      
      # Binary SMD
      pooled_p <- if (!is.na(p1) && !is.na(p2)) (p1 + p2) / 2 else NA_real_
      smd_binary <- if (!is.na(pooled_p) && pooled_p > 0 && pooled_p < 1) {
        (p2 - p1) / sqrt(pooled_p * (1 - pooled_p))
      } else {
        NA_real_
      }
      
      out_rd <- data.table(
        site = site_name,
        variable = var_name,
        variable_type = "categorical",
        comparison = comparison_label,
        level = target_level,
        group1 = g1,
        group2 = g2,
        group1_n = n1,
        group2_n = n2,
        group1_mean = p1,
        group2_mean = p2,
        group1_sd = NA_real_,
        group2_sd = NA_real_,
        group1_median = NA_real_,
        group2_median = NA_real_,
        group1_p25 = NA_real_,
        group2_p25 = NA_real_,
        group1_p75 = NA_real_,
        group2_p75 = NA_real_,
        statistic = "risk_difference",
        estimate = rd,
        ci_lower = rd_lcl,
        ci_upper = rd_ucl,
        p_value = p_value,
        smd = smd_binary,
        effect_measure = "risk_difference",
        test = test_used,
        note = sprintf(
          "Binary categorical variable. level='%s' treated as event.",
          target_level
        )
      )
      
      out <- rbind(out, out_rd, fill = TRUE)
      
    } else if (n_levels > 2) {
      out[, note := "Multi-level categorical variable: overall p-value and omnibus SMD reported; CI not populated."]
    }
    
    out
  }
  
  #--------------------------------------------------#
  # Run continuous variables
  #--------------------------------------------------#
  if (!is.null(cont_vars) && length(cont_vars) > 0) {
    for (v in cont_vars) {
      if (!(v %in% names(dt))) {
        warning(sprintf("Continuous variable '%s' not found. Skipping.", v))
        next
      }
      out_list[[paste0("cont_", v)]] <- compare_continuous(
        x = dt[[v]],
        group = dt[[by]],
        var_name = v
      )
    }
  }
  
  #--------------------------------------------------#
  # Run categorical variables
  #--------------------------------------------------#
  if (!is.null(cat_vars) && length(cat_vars) > 0) {
    for (v in cat_vars) {
      if (!(v %in% names(dt))) {
        warning(sprintf("Categorical variable '%s' not found. Skipping.", v))
        next
      }
      out_list[[paste0("cat_", v)]] <- compare_categorical(
        x = dt[[v]],
        group = dt[[by]],
        var_name = v
      )
    }
  }
  
  out <- rbindlist(out_list, fill = TRUE, use.names = TRUE)
  
  setcolorder(out, c(
    "site", "variable", "variable_type", "comparison", "level",
    "group1", "group2",
    "group1_n", "group2_n",
    "group1_mean", "group2_mean",
    "group1_sd", "group2_sd",
    "group1_median", "group2_median",
    "group1_p25", "group2_p25",
    "group1_p75", "group2_p75",
    "statistic", "estimate", "ci_lower", "ci_upper",
    "p_value", "smd", "effect_measure", "test", "note"
  ))
  
  return(out[])
}
```

#IPW Calculation
```{r}
ipw_outcome <- total_cohort[, .(
  reached_48h = as.integer(imv_duration_hours_final >= 48)
), by = hospitalization_id]

# Merge the reached_48h flag back to your dataset
ipw_sensitivity_analysis <- merge(total_cohort, ipw_outcome, by = "hospitalization_id")

# Use patient-level covariates to predict censoring
ipw_model <- glm(reached_48h ~ age_at_admission + sex_category + race_category + ethnicity_category + laps2 + factor(hospital_icu_id) + elixhauser_score + factor(observation_year),
                 data = unique(ipw_sensitivity_analysis, by = "hospitalization_id"),
                 family = binomial())

total_cohort[, prob_reach_48h := predict(ipw_model, newdata = total_cohort, type = "response")]
total_cohort[, ipw := 1 / prob_reach_48h]
```

# Build Language Categorizations
```{r}
# 1: English v Non-English
total_cohort[, lang_cat_1 := fifelse(
  language_category == "english", "english",
  fifelse(!is.na(language_category), "non-english", NA_character_)
)]

# 2: English v Spanish v all other non-English
total_cohort[, lang_cat_2 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  !is.na(language_category), "other non-english",
  default = NA_character_
)]

# 3: English v Spanish v third most frequent hospital system wide language - all others NA
# Define third_lang
# Skip "Unknown or NA" if it appears in the 3rd position
third_lang <- language_data[
  !tolower(language_category) %in% c("unknown or na"),
  language_category
][3]

# Define the variable
total_cohort[, lang_cat_3 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category == third_lang, third_lang,
  default = NA_character_
)]

# 4: English v Spanish v Chinese v Vietnamese (three non-English languages with the most evidence supporting the existence of health disparities) - all others NA
total_cohort[, lang_cat_4 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category == "chinese", "chinese",
  language_category == "vietnamese", "vietnamese",
  default = NA_character_
)]

# 5: English v Spanish v Broad Regional Groupings
# South Asian, East Asian, Southeast Asian, Sub-Saharan African, non-colonial European
# Others that do not fit are NA

# -----------------------------
# Define language groupings
# -----------------------------

south_asian_langs <- c(
  "gujarati",
  "hindi",
  "urdu",
  "punjabi",
  "bengali",
  "nepali, marathi, or other indic languages",
  "telugu",
  "tamil",
  "malayalam, kannada, or other dravidian languages"
)

east_asian_langs <- c(
  "chinese",
  "japanese",
  "korean"
)

southeast_asian_langs <- c(
  "vietnamese",
  "khmer",
  "thai, lao, or other tai-kadai languages"
)

sub_saharan_african_langs <- c(
  "amharic, somali, or other afro-asiatic languages",
  "yoruba, twi, igbo, or other languages of western africa",
  "swahili or other languages of central, eastern, and southern africa"
)

non_colonial_european_langs <- c(
  "german",
  "greek",
  "russian",
  "polish",
  "italian",
  "serbo-croatian",
  "ukrainian or other slavic languages",
  "yiddish, pennsylvania dutch, or other west germanic languages",
  "other european indo-european languages")
  
mena_langs <- c(
  "arabic",
  "persian",
  "hebrew"
)

total_cohort[, lang_cat_5 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category == "french", "french",
  language_category == "portugese", "portugese",
  language_category %in% south_asian_langs, "south_asian",
  language_category %in% east_asian_langs, "east_asian",
  language_category %in% southeast_asian_langs, "southeast_asian",
  language_category %in% sub_saharan_african_langs, "sub_saharan_african",
  language_category %in% non_colonial_european_langs, "non_colonial_european",
  language_category %in% mena_langs, "mena",
  default = NA_character_
)]

# 6: English v Spanish - all others NA
total_cohort[, lang_cat_6 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  default = NA_character_
)]


```

# Total Cohort for Non-English Analysis
```{r}
# Identify reference points for model
total_cohort$language_category <- relevel(as.factor(total_cohort$language_category), ref = "english")
#total_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(total_cohort$top_five_non_eng_non_spanish), ref = "english")
#total_cohort$non_eng <- relevel(as.factor(total_cohort$non_eng), ref = "english")
#total_cohort$three_language_cat <- relevel(as.factor(total_cohort$three_language_cat), ref = "english")
total_cohort$sex_category <- relevel(as.factor(total_cohort$sex_category), ref = "male")
total_cohort$race_category <- relevel(as.factor(total_cohort$race_category), ref = "white")
total_cohort$ethnicity_category <- relevel(as.factor(total_cohort$ethnicity_category), ref = "non-hispanic")

#Descriptive Table One
table1_english_v_nonenglish_cohort_csv <- make_table1_federated(
  dt = total_cohort,
  cat_vars = cat_vars,
  cont_vars = cont_vars,
  by = "non_eng",   
  site_name = institution
)

fwrite(
  table1_english_v_nonenglish_cohort_csv,
  file.path(sedation_tables_dir, paste0(institution, "_table1_english_v_nonenglish_cohort_descriptive.csv"))
)

#Comparison Table One
table1_english_v_nonenglish_cohort_comparisons <- make_table1_comparisons(
  dt = total_cohort,
  by = "non_eng",
  cont_vars = cont_vars,
  cat_vars = cat_vars,
  site_name = institution,
  group_levels = c("english", "non_english")
)

fwrite(
  table1_english_v_nonenglish_cohort_comparisons,
  file.path(sedation_tables_dir, paste0(institution, "_table1_english_v_nonenglish_cohort_comparisons.csv"))
)
```

# Cohort for Top 5 non-English, non-Spanish Language
```{r}
#top_five_cohort <- total_cohort[top_five_non_eng_non_spanish %in% c("english", "top_five_non_eng_non_spanish", "spanish")]
#top_five_cohort$top_five_non_eng_non_spanish <- relevel(as.factor(top_five_cohort$top_five_non_eng_non_spanish), ref = "english")
#top_five_cohort$non_eng <- relevel(as.factor(top_five_cohort$non_eng), ref = "english")
top_five_cohort$sex_category <- relevel(as.factor(top_five_cohort$sex_category), ref = "male")
top_five_cohort$race_category <- relevel(as.factor(top_five_cohort$race_category), ref = "white")
top_five_cohort$ethnicity_category <- relevel(as.factor(top_five_cohort$ethnicity_category), ref = "non-hispanic")

#Basic Table One
table1_top5_cohort_csv <- make_table1_federated(
  dt = total_cohort,
  cat_vars = cat_vars,
  cont_vars = cont_vars,
  by = "top_five_non_eng_non_spanish",   # or NULL if overall only
  site_name = institution
)

fwrite(
  table1_total_cohort_csv,
  file.path(sedation_tables_dir, paste0(institution, "_table1_federated.csv"))
)
```

# Cohort for Sensitivity Analysis Involving All Spanish/English Pts regardless of ICU of admission
```{r}
english_spanish_all_icu_analysis <- total_cohort[language_category %in% c("english", "spanish")]
english_spanish_all_icu_analysis[, race_category := as.character(race_category)]
english_spanish_all_icu_analysis[, race_category := fifelse(
  race_category %in% c("asian"),
  "other/unknown",
  race_category
)]

# Identify reference points for model
english_spanish_all_icu_analysis$language_category <- relevel(as.factor(english_spanish_all_icu_analysis$language_category), ref = "english")
english_spanish_all_icu_analysis$sex_category <- relevel(as.factor(english_spanish_all_icu_analysis$sex_category), ref = "male")
english_spanish_all_icu_analysis$race_category <- relevel(as.factor(english_spanish_all_icu_analysis$race_category), ref = "white")
english_spanish_all_icu_analysis$ethnicity_category <- relevel(as.factor(english_spanish_all_icu_analysis$ethnicity_category), ref = "non-hispanic")


#Basic Table One
# Generate the summary table grouped by language
eng_span_all_icu_table_one <- table1(~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv | language_category, data = english_spanish_all_icu_analysis, factorVars = catVars, test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(eng_span_all_icu_table_one), file.path(sedation_tables_dir, paste0(institution, "_eng_span_all_icu_cohort_table_one_manuscript.html")))

```

# Primary Cohort for Analysis
```{r}
# Step 1: Count Spanish-speaking patients by hospital_icu_id
spanish_counts <- english_spanish_all_icu_analysis[
  language_category == "spanish", 
  .N, 
  by = hospital_icu_id
]

# Step 2: Identify hospital_icu_id values with 0 Spanish-speaking patients
# (We do this by first getting all unique hospital_icu_id values and checking which ones are missing)
all_hospitals <- unique(english_spanish_all_icu_analysis$hospital_icu_id)
hospitals_with_spanish <- spanish_counts$hospital_icu_id
hospitals_to_remove <- setdiff(all_hospitals, hospitals_with_spanish)

# Step 3: Filter out those hospitals from the original table
english_spanish_primary_analysis <- english_spanish_all_icu_analysis[
  !hospital_icu_id %in% hospitals_to_remove
]

english_spanish_primary_analysis <- english_spanish_primary_analysis %>%
  mutate(language_category = droplevels(language_category))

# Identify reference points for model
english_spanish_primary_analysis$language_category <- relevel(as.factor(english_spanish_primary_analysis$language_category), ref = "english")
english_spanish_primary_analysis$sex_category <- relevel(as.factor(english_spanish_primary_analysis$sex_category), ref = "male")
english_spanish_primary_analysis$race_category <- relevel(as.factor(english_spanish_primary_analysis$race_category), ref = "white")
english_spanish_primary_analysis$ethnicity_category <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category), ref = "non-hispanic")
#english_spanish_primary_analysis$covid_status <- relevel(as.factor(english_spanish_primary_analysis$covid_status), ref = "not detected")
english_spanish_primary_analysis$ethnicity_category_final <- relevel(as.factor(english_spanish_primary_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")
english_spanish_primary_analysis$push_paralytics <- relevel(as.factor(english_spanish_primary_analysis$push_paralytics), ref = "0")


# Create labels for readability
label(english_spanish_primary_analysis$imv_duration_hours) <- "<u>Hours Mechanically Ventilated</u>"
label(english_spanish_primary_analysis$race_category) <- "<u>Race</u>"
label(english_spanish_primary_analysis$ethnicity_category) <- "<u>Ethnicity</u>"
label(english_spanish_primary_analysis$sex_category) <- "<u>Sex</u>"
label(english_spanish_primary_analysis$age_at_admission) <- "<u>Age</u>"
label(english_spanish_primary_analysis$language_category) <- "Language"
label(english_spanish_primary_analysis$bmi) <- "<u>BMI</u>"
label(english_spanish_primary_analysis$hospital_id) <- "<u>Hospital</u>"
label(english_spanish_primary_analysis$icu_imv) <- "<u>ICU Type</u>"
label(english_spanish_primary_analysis$mortality_28day) <- "<u>Twenty-Eight Day Mortality</u>"
label(english_spanish_primary_analysis$location_of_intubation) <- "<u>Location of Intubation</u>"
label(english_spanish_primary_analysis$elixhauser_score) <- "<u>Elixhauser Score</u>"
label(english_spanish_primary_analysis$SVM_SCORE) <- "<u>Social Vulnerability Metric</u>"
label(english_spanish_primary_analysis$laps2) <- "<u>LAPS2 Score</u>"
label(english_spanish_primary_analysis$time_to_first_rass) <- "<u>Time to First RASS</u>"
label(english_spanish_primary_analysis$rass_frequency_hrs) <- "<u>Frequency Of Rass Measurements</u>"
#label(english_spanish_primary_analysis$covid_status) <- "<u>COVID Status</u>"

# Generate the summary table grouped by language
summary_table <- table1(
  ~ age_at_admission + sex_category + ethnicity_category + race_category + bmi + laps2 + hospital_id + icu_imv + location_of_intubation + elixhauser_score + SVM_SCORE + time_to_first_rass + rass_frequency_hrs 
  | language_category, 
  data = english_spanish_primary_analysis,
  factorVars = catVars,
  test = TRUE)

# Convert to HTML and save as file
save_html(as.tags(summary_table), file.path(sedation_tables_dir, paste0(institution, "_primary_cohort_table_one_manuscript.html")))

vars <- c("age_at_admission", "bmi", "sex_category", "race_category", "ethnicity_category", "laps2", "hospital_id", "icu_imv", "location_of_intubation", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs")
catVars <- c("sex_category", "race_category", "ethnicity_category", "hospital_id", "icu_imv", "location_of_intubation")

# Create table
table1_obj <- CreateTableOne(vars = vars, strata = "language_category",
                             data = english_spanish_primary_analysis,
                             factorVars = catVars)

# Print with SMD and nonparametric summary
print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi"))

print(table1_obj, smd = TRUE)

# Convert the table1 object to a data frame
table1_df <- print(table1_obj, smd = TRUE, nonnormal = c("age_at_admission", "bmi", "laps2", "elixhauser_score", "SVM_SCORE", "time_to_first_rass", "rass_frequency_hrs"), printToggle = FALSE)

# Save to CSV
# Save the combined results to a file
fwrite(table1_df, file.path(sedation_tables_dir, paste0(institution, "_IQR_SMD_Table1.csv")))
```

#STEP TWO: PRIMARY ANALYSIS
```{r}
# Fit the unadjusted primary model
primary_model_unadjusted <- lm(percent_deep_sedation_48 ~ language_category, data = english_spanish_primary_analysis)

summary(primary_model_unadjusted)

# Extract coefficients, standard errors, variance, and p-values for the unadjusted model
primary_results_unadjusted <- data.frame(
  term = names(coef(primary_model_unadjusted)),
  estimate = coef(primary_model_unadjusted),
  std.error = summary(primary_model_unadjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_unadjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_unadjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

primary_model_adjusted <- lm(percent_deep_sedation_48 ~ language_category + race_category + age_category + sex_category + bmi + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(primary_model_adjusted)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
primary_results_adjusted <- data.frame(
  term = names(coef(primary_model_adjusted)),
  estimate = coef(primary_model_adjusted),
  std.error = summary(primary_model_adjusted)$coefficients[, "Std. Error"],
  variance = (summary(primary_model_adjusted)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(primary_model_adjusted)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Combine the results into one data frame
primary_results_combined <- bind_rows(
  primary_results_unadjusted %>% mutate(model = "Unadjusted"),
  primary_results_adjusted %>% mutate(model = "Adjusted")
)

# Save the combined results to a file
fwrite(primary_results_combined, file.path(final_data, paste0(institution, "_primary_analysis_results.csv")))
```


#STEP FOUR: SECONDARY ANALYSES
#Secondary Analysis: Total time mechnically ventilated
```{r}
secondary_model_imv_time <- lm(imv_duration_hours_final ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_imv_time)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_imv_time_results <- data.frame(
  term = names(coef(secondary_model_imv_time)),
  estimate = coef(secondary_model_imv_time),
  std.error = summary(secondary_model_imv_time)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_imv_time)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_imv_time)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_imv_time_results, file.path(final_data, paste0(institution, "_secondary_analysis_imv_time.csv")))

```


#Secondary Analysis: 28-Day Ventilator-Free-Days
```{r}

secondary_model_vfd <- lm(vent_free_days_28 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_vfd)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_vfd_results <- data.frame(
  term = names(coef(secondary_model_vfd)),
  estimate = coef(secondary_model_vfd),
  std.error = summary(secondary_model_vfd)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_vfd)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_vfd)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_vfd_results, file.path(final_data, paste0(institution, "_secondary_analysis_vfd.csv")))

```


#Secondary Analysis: 28-Day Mortality
```{r}
secondary_model_28day_mortality <- glm(mortality_28day ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_28day_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_30dm <- summary(secondary_model_28day_mortality)$coefficients
odds_ratios <- exp(coef_summary_30dm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_30dm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_30dm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_30day_mortality_results <- data.frame(
    Variable = rownames(coef_summary_30dm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_30day_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_30day_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_28_day_mortality.csv")))

```



#Secondary Analysis: In-Hospital Mortality
```{r}
secondary_model_inhospital_mortality <- glm(death_during_admission ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_inhospital_mortality)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_inhospitalm <- summary(secondary_model_inhospital_mortality)$coefficients
odds_ratios <- exp(coef_summary_inhospitalm[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_inhospitalm[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_inhospitalm[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_inhospital_mortality_results <- data.frame(
    Variable = rownames(coef_summary_inhospitalm),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_inhospital_mortality_results

# Write the results to a CSV file
fwrite(secondary_model_inhospital_mortality_results, file.path(final_data, paste0(institution, "_secondary_analysis_in_hospital_mortality.csv")))

```


#Secondary Analysis: Tracheostomy Odds
```{r}
secondary_model_trach <- glm(trached ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_trach)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_trach <- summary(secondary_model_trach)$coefficients
odds_ratios <- exp(coef_summary_trach[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_trach[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_trach[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_trach_results <- data.frame(
    Variable = rownames(coef_summary_trach),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_trach_results

# Write the results to a CSV file
fwrite(secondary_model_trach_results, file.path(final_data, paste0(institution, "_secondary_analysis_trach.csv")))

```


#Secondary Analysis: Reintubation Odds
```{r}
secondary_model_reintubation <- glm(reintubation_final ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_reintubation)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_reintubation <- summary(secondary_model_reintubation)$coefficients
odds_ratios <- exp(coef_summary_reintubation[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_reintubation[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_reintubation[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_reintubation_results <- data.frame(
    Variable = rownames(coef_summary_reintubation),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_reintubation_results

# Write the results to a CSV file
fwrite(secondary_model_reintubation_results, file.path(final_data, paste0(institution, "_secondary_analysis_reintubation.csv")))

```


#Exploratory Analysis: Propofol Dosing
```{r}

secondary_model_propofol <- lm(propofol_avg_mcgkgmin_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_propofol)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_propofol_results <- data.frame(
  term = names(coef(secondary_model_propofol)),
  estimate = coef(secondary_model_propofol),
  std.error = summary(secondary_model_propofol)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_propofol)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_propofol)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_propofol_results, file.path(final_data, paste0(institution, "_secondary_analysis_vfd.csv")))

```

#Exploratory Analysis: Precedex Dosing
```{r}

secondary_model_precedex <- lm(dexmedetomidine_avg_mcgkghr_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year),
               data = english_spanish_primary_analysis)

summary(secondary_model_precedex)

# Extract coefficients, standard errors, variance, and p-values for the adjusted model
secondary_model_precedex_results <- data.frame(
  term = names(coef(secondary_model_precedex)),
  estimate = coef(secondary_model_precedex),
  std.error = summary(secondary_model_precedex)$coefficients[, "Std. Error"],
  variance = (summary(secondary_model_precedex)$coefficients[, "Std. Error"])^2,  # Calculate variance
  p.value = summary(secondary_model_precedex)$coefficients[, "Pr(>|t|)"]  # Extract p-values
)

# Save the results to a file
fwrite(secondary_model_precedex_results, file.path(final_data, paste0(institution, "_secondary_analysis_vfd.csv")))

```


#PARALYTIC ANALYSES

#Whole Cohort
```{r}
paralysis_all_three_lang_cont_total_cohort <- glm(continuous_nmba_48 ~ language_category + race_category + ethnicity_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2 + factor(observation_year), data = total_cohort, 
                                  family = binomial(link = "logit"))

#Removed covid_status
summary(paralysis_all_three_lang_cont_total_cohort)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis <- summary(paralysis_all_three_lang_cont_total_cohort)$coefficients
odds_ratios <- exp(coef_summary_paralysis[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_all_three_lang_cont_results_total_cohort <- data.frame(
    Variable = rownames(coef_summary_paralysis),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
view(secondary_model_paralysis_all_three_lang_cont_results_total_cohort)

# Write the results to a CSV file
fwrite(secondary_model_paralysis_all_three_lang_cont_results_total_cohort, file.path(paralytics_final_data, paste0(institution, "_secondary_analysis_CONTINUOUS_paralysis_all_three_lang_total_cohort.csv")))


```

#Secondary Analysis: Odds of Starting Continuous Paralysis in the first 48 hours
```{r}
secondary_model_paralysis_cont <- glm(continuous_nmba_48 ~ language_category + race_category + age_category + bmi + sex_category + factor(hospital_icu_id) + elixhauser_score + SVM_SCORE + laps2, data = english_spanish_primary_analysis, 
                                  family = binomial(link = "logit"))

summary(secondary_model_paralysis_cont)

# Extract coefficients and calculate odds ratios, standard errors, and variance
coef_summary_paralysis_cont <- summary(secondary_model_paralysis_cont)$coefficients
odds_ratios <- exp(coef_summary_paralysis_cont[, "Estimate"])  # Exponentiated coefficients
standard_errors <- coef_summary_paralysis_cont[, "Std. Error"]
variance <- standard_errors^2  # Calculate variance
p_values <- coef_summary_paralysis_cont[, "Pr(>|z|)"]  # Extract p-values

# Create a data frame to hold the results
secondary_model_paralysis_cont_results <- data.frame(
    Variable = rownames(coef_summary_paralysis_cont),
    Odds_Ratio = odds_ratios,
    Std_Error = standard_errors,
    Variance = variance,  # Add variance
    P_Value = p_values    # Add p-values
)
secondary_model_paralysis_cont_results

# Write the results to a CSV file
fwrite(secondary_model_paralysis_cont_results, file.path(final_data, paste0(institution, "_secondary_analysis_paralysis_cont.csv")))
```

#Plot
library(data.table)

dt <- copy(secondary_model_paralysis_all_three_lang_cont_results_total_cohort)
setDT(dt)

dt[, log_or := log(Odds_Ratio)]
dt[, lower := exp(log_or - 1.96 * Std_Error)]
dt[, upper := exp(log_or + 1.96 * Std_Error)]

dt_lang <- dt[grepl("^language_category", Variable)]

# Clean up labels (remove prefix)
dt_lang[, language_label := gsub("^language_category", "", Variable)]

# Optional: make nicer labels
dt_lang[language_label == "", language_label := "Reference (English)"]

dt_lang[, `:=`(
  lower_plot = pmax(lower, 0.1),
  upper_plot = pmin(upper, 10)
)]
library(ggplot2)

ggplot(dt_lang, aes(x = Odds_Ratio, y = reorder(language_label, Odds_Ratio))) +
  geom_errorbarh(aes(xmin = lower_plot, xmax = upper_plot), height = 0.2, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10() +
  coord_cartesian(xlim = c(0.1, 20)) +
  labs(
    title = "Odds of Paralytic Use by Language",
    x = "Odds Ratio",
    y = "Language Category"
  ) +
  theme_minimal()
  
```{r}
library(data.table)
library(ggplot2)

dt_plot <- total_cohort[
  icu_imv != "mixed_neuro_icu" &
  imv_duration_hours_final >= 24
]

ggplot(dt_plot, aes(x = vent_free_days_28)) +
  geom_histogram(binwidth = 1, boundary = 0, closed = "left") +
  labs(
    title = "Distribution of Ventilator-Free Days (28-day)",
    x = "Ventilator-Free Days",
    y = "Count"
  ) +
  theme_minimal()

ggplot(dt_plot, aes(x = vent_free_days_28_alt)) +
  geom_histogram(binwidth = 1, boundary = -1, closed = "left") +
  labs(
    title = "Distribution of Ventilator-Free Days (28-day, Death = -1)",
    x = "Ventilator-Free Days (Alt)",
    y = "Count"
  ) +
  theme_minimal()
```

  