#START
```{r setup}
# Install Packages

packages <- c("tidyverse", "zoo", "data.table", "dplyr", "lubridate", "tidyr", "fst", "arrow", "table1", "lme4", "metafor", "yaml", "pscl", "MASS", "htmltools", "car", "WeightIt", "MatchIt", "purrr", "survival", "mediation", "cmprsk", "gfoRmula", "nnet", "medflex", "bruceR", "survey", "ipw", "geepack", "readr", "collapse", "stringr", "rprojroot", "tableone", "lmtest", "broom", "sandwich", "pROC", "ResourceSelection")
install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)

# Load necessary libraries
library(data.table)
library(dplyr)
library(zoo)  # for na.locf function
library(lubridate)
library(tidyr)
library(fst)
library(arrow)
library(table1)
library(metafor)
library(yaml)
library(htmltools)
library(MASS)
library(pscl)
library(car)
library(stringr)
library(rprojroot)
library(tableone)
library(lmtest)
library(broom)
library(sandwich)
library(pROC)
library(ResourceSelection)


# Define paths
project_root <- file.path(
  find_root(rprojroot::has_dir("deep_sedation_language_final")),
  "deep_sedation_language_final"
)

config_dir <- file.path(project_root, "config")
knitr::opts_knit$set(root.dir = config_dir)

# Look for both .yaml and .yml
config_file <- c(
  file.path(config_dir, "deep_sedation_language_config.yaml"),
  file.path(config_dir, "deep_sedation_language_config.yml")
)

# Keep only existing files
config_file <- config_file[file.exists(config_file)]

# Error if neither exists
if (length(config_file) == 0) {
  stop("No config file found (.yaml or .yml)")
}

# If both exist, take the first (yaml prioritized here)
config <- yaml::read_yaml(config_file[1])

# Define directories based on YAML configuration
data_dir <- config$data_dir
local_dir <- config$local_dir
output_dir <- file.path(local_dir, "output/")
intermediate_dir <- file.path(local_dir, "intermediate_data_DO_NOT_UPLOAD/")  # Define output_dir based on data_dir
final_data_dir <- file.path(output_dir, "FINAL_DATA_UPLOAD_ME/")
rerun_data_dir <- file.path(output_dir, "RERUN_DATA_UPLOAD_ME/")
rerun_tables_dir <- file.path(rerun_data_dir, "rerun_tables")
rerun_final_data <- file.path(rerun_data_dir, "rerun_data")
rerun_paralytics <- file.path(rerun_data_dir, "rerun_paralytics_data")
cohort_characteristics <- file.path(final_data_dir, "cohort_characteristics/")
sedation_project <- file.path(final_data_dir, "sedation_project/")
sedation_tables_dir <- file.path(sedation_project, "final_tables/")
sedation_final_data <- file.path(sedation_project, "final_data/")
sedation_sensitivity <- file.path(sedation_final_data, "sensitivity_analyses/")
sedation_secondary <- file.path(sedation_final_data, "secondary_analyses/")
sedation_exploratory <- file.path(sedation_final_data, "exploratory_analyses/")
sedation_subgroup <- file.path(sedation_final_data, "subgroup_analyses/")
paralytics_project <- file.path(final_data_dir, "paralytics_project/")
paralytics_tables_dir <- file.path(paralytics_project, "final_tables/")  
paralytics_final_data <- file.path(paralytics_project, "final_data/")
paralytics_exploratory <- file.path(paralytics_final_data, "exploratory_analyses/")
paralytics_exploratory_tables <- file.path(paralytics_tables_dir, "exploratory_tables/")
arf_project <- file.path(final_data_dir, "arf_project/")
arf_tables_dir <- file.path(arf_project, "final_tables/")  
arf_final_data <- file.path(arf_project, "final_data/")
measurements_project <- file.path(final_data_dir, "measurements_project/")
measurements_tables_dir <- file.path(measurements_project, "final_tables/")  
measurements_final_data <- file.path(measurements_project, "final_data/")
measurements_model_data <- file.path(measurements_final_data, "model_data/")

file_type <- config$file_type
institution <- config$institution
time_zone <- config$time_zone

# Create all directories
dirs_to_create <- c(
  output_dir,
  intermediate_dir,
  final_data_dir,
  cohort_characteristics,
  sedation_project,
  sedation_tables_dir,
  sedation_final_data,
  sedation_sensitivity,
  sedation_secondary,
  sedation_exploratory,
  sedation_subgroup,
  paralytics_project,
  paralytics_tables_dir,
  paralytics_final_data,
  paralytics_exploratory,
  paralytics_exploratory_tables,
  arf_project,
  arf_tables_dir,
  arf_final_data,
  measurements_project,
  measurements_tables_dir,
  measurements_final_data,
  measurements_model_data,
  rerun_data_dir,
  rerun_tables_dir,
  rerun_final_data,
  rerun_paralytics
)

for (dir in dirs_to_create) {
  dir.create(dir, recursive = TRUE, showWarnings = FALSE)
}
```

#Upload Data
```{r}
# Upload Analytic File
total_cohort <- read_parquet(paste0(intermediate_dir, "analytic_file", institution, ".parquet"))
setDT(total_cohort)
total_cohort[
  race_category == "asian",
  race_category := "other/unknown"
]
total_cohort[, observation_year := as.double(observation_year)]
# Upload hospital system language information
language_data <- fread(paste0(cohort_characteristics, "/", institution, "_hospital_system_language_counts", ".csv"))
setDT(language_data)

cat_vars <- c("hospital_id", "icu_imv", "hospital_type")

cont_vars <- c()

```


#PART ZERO: Fix Hospitalization
```{r}
# This block adjusts for hospitalizations that are continuous within a health system.
# It looks at patient_id, discharge_dttm, and admission_dttm for hospitalization_ids
# that are close in time, creates new chronological hospitalization_ids, and builds
# a crosswalk (hosp_map) used by read_data() to remap all downstream tables.

read_data_start <- function(file_path, select = NULL) {
  if (grepl("\\.csv$|\\.csv\\.gz$", file_path)) {
    dt <- fread(file_path, select = select)
  } else if (grepl("\\.parquet$", file_path)) {
    if (is.null(select)) {
      dt <- arrow::read_parquet(file_path)
    } else {
      dt <- arrow::read_parquet(file_path, col_select = all_of(select))
    }
  } else if (grepl("\\.fst$", file_path)) {
    if (is.null(select)) {
      dt <- fst::read_fst(file_path)
    } else {
      dt <- fst::read_fst(file_path, columns = all_of(select))
    }
  } else {
    stop("Unsupported file format. Only CSV, Parquet, and FST are supported.")
  }

  dt <- data.table::as.data.table(dt)

  dttm_cols <- names(dt)[grepl("dttm$", names(dt))]
  local_tz  <- time_zone
  message("Using timezone: ", local_tz)

  for (var in dttm_cols) {
    col <- dt[[var]]

    if (inherits(col, "POSIXct")) {
      existing_tz <- attr(col, "tzone")

      # Already matches local tz — do nothing
      if (!is.null(existing_tz) && nzchar(existing_tz) && existing_tz == local_tz) {
        message("✅ ", var, ": tzone already '", local_tz, "', no change")
        next
      }

      # Blank or NULL — timestamps already in local time, just stamp the label
      if (is.null(existing_tz) || !nzchar(existing_tz)) {
        dt[, (var) := lubridate::force_tz(get(var), tzone = local_tz)]
        message("🔧 ", var, ": tzone was blank/null, stamped as '", local_tz, "' (no clock shift)")
        next
      }

      # UTC or GMT — timestamps are genuinely UTC, convert with clock shift
      if (existing_tz %in% c("UTC", "GMT")) {
        dt[, (var) := lubridate::with_tz(get(var), tzone = local_tz)]
        message("🌍 ", var, ": tzone was '", existing_tz, "', converted to '", local_tz, "' (clock shifted)")
        next
      }

      # Different real timezone — warn and leave alone
      warning("⚠️  ", var, ": tzone is '", existing_tz, "', which differs from YAML time_zone '",
              local_tz, "'. Leaving as-is. Check your source data.")
      next
    }

    if (inherits(col, "Date")) {
      dt[, (var) := lubridate::force_tz(as.POSIXct(get(var)), tzone = local_tz)]
      message("📅 ", var, ": Date -> POSIXct midnight, stamped as '", local_tz, "'")
      next
    }

    raw_vec <- if (is.factor(col)) as.character(col) else col
    suppressWarnings(parsed <- lubridate::ymd_hms(raw_vec, quiet = TRUE))
    bad_idx <- is.na(parsed) & !is.na(raw_vec) & nzchar(trimws(raw_vec))
    if (any(bad_idx)) {
      suppressWarnings(parsed_date <- lubridate::ymd(raw_vec[bad_idx], quiet = TRUE))
      parsed[bad_idx] <- as.POSIXct(parsed_date)
    }
    dt[, (var) := lubridate::force_tz(parsed, tzone = local_tz)]
    message("🔧 ", var, ": parsed from character/factor, stamped as '", local_tz, "'")
  }

  dt %>% glimpse()
  return(dt)
}

# Load and clean hospitalization table
clif_hospitalization_start <-
  read_data_start(paste0(data_dir, "clif_hospitalization", file_type)) |>
  dplyr::select(hospitalization_id, admission_dttm, discharge_dttm, age_at_admission,
                patient_id, discharge_name, discharge_category) |>
  mutate(across(where(is.factor), as.character)) |>
  mutate(across(where(is.character), str_to_lower)) |>
  as.data.table()

# Identify linked encounters (admissions within 6 hours of a discharge)
hospital_blocks <- clif_hospitalization_start[, .(patient_id, hospitalization_id, admission_dttm, discharge_dttm)]
setorder(hospital_blocks, patient_id, admission_dttm)

hospital_blocks[, next_diff_time := as.numeric(difftime(
  shift(admission_dttm, type = "lead"), discharge_dttm, units = "hours"
)), by = patient_id]

hospital_blocks[, prev_diff_time := as.numeric(difftime(
  admission_dttm, shift(discharge_dttm, type = "lag"), units = "hours"
)), by = patient_id]

hospital_blocks[, linked := as.integer(next_diff_time < 6 | prev_diff_time < 6)]

# Number link groups within patient using cumsum
hospital_blocks[, link_group := {
  is_new_group <- is.na(shift(linked)) | linked == 0 |
                  (linked == 1 & (is.na(prev_diff_time) | prev_diff_time >= 6))
  cumsum(is_new_group)
}, by = patient_id]

# Merge link group back to full hospitalization table
clif_hospitalization_start <- hospital_blocks[
  clif_hospitalization_start,
  on = .(patient_id, hospitalization_id)
]

# Unlinked hospitalizations get their own block
clif_hospitalization_start[is.na(linked), link_group := .I]
clif_hospitalization_start[, encounter_block := .GRP, by = .(patient_id, link_group)]

# Collapse to one row per encounter block
clif_hospitalization <- clif_hospitalization_start[
  , .(
    hospitalization_id  = hospitalization_id[1],
    age_at_admission    = age_at_admission[1],
    admission_dttm      = admission_dttm[1],
    discharge_dttm      = discharge_dttm[.N],
    discharge_name      = discharge_name[.N],
    discharge_category  = discharge_category[.N]
  ),
  by = .(patient_id, encounter_block)
][, clif_hospitalizations_joined_id := .I]

# Build hosp_map crosswalk for downstream remapping in read_data()
hosp_map <- clif_hospitalization_start[
  clif_hospitalization,
  on = .(patient_id, encounter_block),
  .(
    old_hosp_id              = x.hospitalization_id,
    new_hosp_id              = clif_hospitalizations_joined_id,
    final_admit              = i.admission_dttm,
    final_discharge          = i.discharge_dttm,
    final_discharge_name     = i.discharge_name,
    final_discharge_category = i.discharge_category
  )
]
```


# PART ONE: Upload clif tables
```{r}
read_data <- function(file_path, select = NULL) {

  # ---- 1) Read ----
  if (grepl("\\.csv$|\\.csv\\.gz$", file_path)) {
    dt <- data.table::fread(file_path, select = select)

  } else if (grepl("\\.parquet$", file_path)) {
    if (is.null(select)) {
      dt <- arrow::read_parquet(file_path)
    } else {
      dt <- arrow::read_parquet(file_path, col_select = tidyselect::all_of(select))
    }

  } else if (grepl("\\.fst$", file_path)) {
    if (is.null(select)) {
      dt <- fst::read_fst(file_path)
    } else {
      dt <- fst::read_fst(file_path, columns = tidyselect::all_of(select))
    }

  } else {
    stop("Unsupported file format. Only CSV, Parquet, and FST are supported.")
  }

  dt <- data.table::as.data.table(dt)

  # ---- 2) Optional remap using hosp_map ----
  if (exists("hosp_map", envir = .GlobalEnv) && "hospitalization_id" %in% names(dt)) {

    hm <- data.table::as.data.table(get("hosp_map", envir = .GlobalEnv))
    if (!"old_hosp_id" %in% names(hm)) stop("hosp_map missing old_hosp_id")

    hm[, old_hosp_id := as.character(old_hosp_id)]
    if ("new_hosp_id" %in% names(hm)) hm[, new_hosp_id := as.character(new_hosp_id)]

    dt[, hospitalization_id := as.character(hospitalization_id)]

    has_dates <- all(c("admission_dttm", "discharge_dttm") %in% names(dt))

    if (has_dates) {
      message("🔄 remapping hosp_id, admit & discharge")
      dt <- hm[dt, on = .(old_hosp_id = hospitalization_id)]
      dt[, hospitalization_id := data.table::fcoalesce(new_hosp_id, old_hosp_id)]

      if ("final_admit"              %in% names(dt)) dt[, admission_dttm       := data.table::fcoalesce(final_admit,              admission_dttm)]
      if ("final_discharge"          %in% names(dt)) dt[, discharge_dttm       := data.table::fcoalesce(final_discharge,          discharge_dttm)]
      if ("final_discharge_name"     %in% names(dt)) dt[, discharge_name       := data.table::fcoalesce(final_discharge_name,     discharge_name)]
      if ("final_discharge_category" %in% names(dt)) dt[, discharge_category   := data.table::fcoalesce(final_discharge_category, discharge_category)]

    } else {
      message("🔄 remapping just hosp_id")
      dt <- hm[dt, on = .(old_hosp_id = hospitalization_id)]
      dt[, hospitalization_id := data.table::fcoalesce(new_hosp_id, old_hosp_id)]
    }

    drop_cols <- intersect(
      c("new_hosp_id", "old_hosp_id", "final_admit", "final_discharge",
        "final_discharge_name", "final_discharge_category"),
      names(dt)
    )
    if (length(drop_cols)) dt[, (drop_cols) := NULL]
  }

  # ---- 3) Normalize all *_dttm columns ----
  dttm_cols <- names(dt)[grepl("dttm$", names(dt))]
  local_tz  <- time_zone
  message("Using timezone: ", local_tz)

  for (var in dttm_cols) {
    col <- dt[[var]]

    if (inherits(col, "POSIXct")) {
      existing_tz <- attr(col, "tzone")

      # Already matches local tz — do nothing
      if (!is.null(existing_tz) && nzchar(existing_tz) && existing_tz == local_tz) {
        message("✅ ", var, ": tzone already '", local_tz, "', no change")
        next
      }

      # Blank or NULL — timestamps already in local time, just stamp the label
      if (is.null(existing_tz) || !nzchar(existing_tz)) {
        dt[, (var) := lubridate::force_tz(get(var), tzone = local_tz)]
        message("🔧 ", var, ": tzone was blank/null, stamped as '", local_tz, "' (no clock shift)")
        next
      }

      # UTC or GMT — timestamps are genuinely UTC, convert with clock shift
      if (existing_tz %in% c("UTC", "GMT")) {
        dt[, (var) := lubridate::with_tz(get(var), tzone = local_tz)]
        message("🌍 ", var, ": tzone was '", existing_tz, "', converted to '", local_tz, "' (clock shifted)")
        next
      }

      # Different real timezone — warn and leave alone
      warning("⚠️  ", var, ": tzone is '", existing_tz, "', which differs from YAML time_zone '",
              local_tz, "'. Leaving as-is. Check your source data.")
      next
    }

    # Date columns
    if (inherits(col, "Date")) {
      if (identical(var, "death_dttm")) {
        dt[, (var) := lubridate::force_tz(
          as.POSIXct(get(var)) + lubridate::hours(23) + lubridate::minutes(59),
          tzone = local_tz
        )]
        message("📅 ", var, ": Date -> POSIXct 23:59, stamped as '", local_tz, "'")
      } else {
        dt[, (var) := lubridate::force_tz(as.POSIXct(get(var)), tzone = local_tz)]
        message("📅 ", var, ": Date -> POSIXct midnight, stamped as '", local_tz, "'")
      }
      next
    }

    # Character / factor — parse then stamp
    raw_vec <- if (is.factor(col)) as.character(col) else col

    suppressWarnings(parsed <- lubridate::ymd_hms(raw_vec, quiet = TRUE))

    bad_idx <- is.na(parsed) & !is.na(raw_vec) & nzchar(trimws(raw_vec))
    if (any(bad_idx)) {
      suppressWarnings(parsed_date <- lubridate::ymd(raw_vec[bad_idx], quiet = TRUE))
      offset <- if (identical(var, "death_dttm")) lubridate::hours(23) + lubridate::minutes(59) else 0
      parsed[bad_idx] <- as.POSIXct(parsed_date) + offset
    }

    dt[, (var) := lubridate::force_tz(parsed, tzone = local_tz)]
    message("🔧 ", var, ": parsed from character/factor, stamped as '", local_tz, "'")
  }

  # ---- 4) Peek & return ----
  if (nrow(dt) > 0) utils::str(head(dt))
  return(dt)
}

#Hospitalization
hospitalization <- as.data.table(read_data(paste0(data_dir, "clif_hospitalization", file_type)))
hospitalization <- unique(hospitalization, by = "hospitalization_id")
#Filter Out Incomplete Hospitalizations
hospitalization <- hospitalization[!is.na(admission_dttm) & !is.na(discharge_dttm) & discharge_dttm >= admission_dttm]
hospitalization[, admission_type_category := tolower(admission_type_category)]
hospitalization[, discharge_category := tolower(discharge_category)]


#ADT 
adt <- as.data.table(read_data(paste0(data_dir, "clif_adt", file_type), select = c("hospitalization_id", "hospital_id", "hospital_type", "in_dttm", "out_dttm", "location_category", "location_type")))
setorder(adt, hospitalization_id, in_dttm) 

adt_hospital_type_counts <- adt[
  , .N,
  by = .(hospital_id, hospital_type)
][
  order(hospital_id, hospital_type)
]

hospital_id_type_final <- adt[
  , .N,
  by = .(hospital_id, hospital_type)
][
  order(hospital_id, -N)
][
  , .SD[1],
  by = hospital_id
]

# Add hospital_type to total_cohort
total_cohort <- hospital_id_type_final[
  total_cohort,
  on = "hospital_id"
]
total_cohort[, N := NULL]

```


#Function to Create Table 1s
```{r}
#Function to Create Table 1s
make_table1_federated <- function(dt,
                                  cat_vars = NULL,
                                  cont_vars = NULL,
                                  by = NULL,
                                  site_name = NA_character_) {
  dt <- as.data.table(copy(dt))
  out_list <- list()
  
  #-----------------------------#
  # CATEGORICAL VARIABLES
  #-----------------------------#
  if (!is.null(cat_vars) && length(cat_vars) > 0) {
    for (v in cat_vars) {
      
      if (!is.null(by)) {
        tmp <- dt[, .(n = .N), by = c(by, v)]
        
        totals <- dt[, .(N = .N), by = by]
        
        tmp <- merge(tmp, totals, by = by, all.x = TRUE)
        
        setnames(tmp, old = by, new = "group")
        setnames(tmp, old = v, new = "level")
        
        tmp[, `:=`(
          site = site_name,
          variable = v,
          variable_type = "categorical",
          level = as.character(level),
          pct = 100 * n / N,
          mean = NA_real_,
          sd = NA_real_,
          median = NA_real_,
          p25 = NA_real_,
          p75 = NA_real_
        )]
        
      } else {
        tmp <- dt[, .(n = .N), by = v]
        total_n <- nrow(dt)
        
        setnames(tmp, old = v, new = "level")
        
        tmp[, `:=`(
          site = site_name,
          group = "overall",
          variable = v,
          variable_type = "categorical",
          level = as.character(level),
          N = total_n,
          pct = 100 * n / N,
          mean = NA_real_,
          sd = NA_real_,
          median = NA_real_,
          p25 = NA_real_,
          p75 = NA_real_
        )]
      }
      
      tmp <- tmp[, .(
        site, group, variable, variable_type, level,
        n, N, pct,
        mean, sd, median, p25, p75
      )]
      
      out_list[[paste0("cat_", v)]] <- tmp
    }
  }
  
  #-----------------------------#
  # CONTINUOUS VARIABLES
  #-----------------------------#
  if (!is.null(cont_vars) && length(cont_vars) > 0) {
    for (v in cont_vars) {
      
      if (!is.null(by)) {
        tmp <- dt[, .(
          n      = sum(!is.na(get(v))),
          N      = .N,
          pct    = NA_real_,
          mean   = as.numeric(if (all(is.na(get(v)))) NA_real_ else mean(get(v), na.rm = TRUE)),
          sd     = as.numeric(if (sum(!is.na(get(v))) <= 1) NA_real_ else sd(get(v), na.rm = TRUE)),
          median = as.numeric(if (all(is.na(get(v)))) NA_real_ else median(get(v), na.rm = TRUE)),
          p25    = as.numeric(if (all(is.na(get(v)))) NA_real_ else quantile(get(v), 0.25, na.rm = TRUE)),
          p75    = as.numeric(if (all(is.na(get(v)))) NA_real_ else quantile(get(v), 0.75, na.rm = TRUE))
        ), by = by]
        
        setnames(tmp, old = by, new = "group")
        
        tmp[, `:=`(
          site = site_name,
          variable = v,
          variable_type = "continuous",
          level = NA_character_
        )]
        
      } else {
        tmp <- dt[, .(
          n      = sum(!is.na(get(v))),
          N      = .N,
          pct    = NA_real_,
          mean   = as.numeric(if (all(is.na(get(v)))) NA_real_ else mean(get(v), na.rm = TRUE)),
          sd     = as.numeric(if (sum(!is.na(get(v))) <= 1) NA_real_ else sd(get(v), na.rm = TRUE)),
          median = as.numeric(if (all(is.na(get(v)))) NA_real_ else median(get(v), na.rm = TRUE)),
          p25    = as.numeric(if (all(is.na(get(v)))) NA_real_ else quantile(get(v), 0.25, na.rm = TRUE)),
          p75    = as.numeric(if (all(is.na(get(v)))) NA_real_ else quantile(get(v), 0.75, na.rm = TRUE))
        )]
        
        tmp[, `:=`(
          site = site_name,
          group = "overall",
          variable = v,
          variable_type = "continuous",
          level = NA_character_
        )]
      }
      
      tmp <- tmp[, .(
        site, group, variable, variable_type, level,
        n, N, pct,
        mean, sd, median, p25, p75
      )]
      
      out_list[[paste0("cont_", v)]] <- tmp
    }
  }
  
  out <- rbindlist(out_list, fill = TRUE, use.names = TRUE)
  
  setcolorder(out, c(
    "site", "group", "variable", "variable_type", "level",
    "n", "N", "pct", "mean", "sd", "median", "p25", "p75"
  ))
  
  #-----------------------------#
  # CENSORING
  #-----------------------------#
  out[variable_type == "categorical" & n < 10,
      `:=`(pct = NA_real_)]
  out[variable_type == "categorical",
      n_censored := ifelse(!is.na(n) & n < 10, "<10", as.character(n))]
  out[variable_type == "categorical" & n < 10,
    `:=`(n = NA_real_, N = NA_real_)]
  
  out[variable_type == "continuous" & n < 10,
      `:=`(mean = NA_real_, sd = NA_real_, median = NA_real_,
           p25 = NA_real_, p75 = NA_real_)]
  out[variable_type == "continuous",
      n_censored := ifelse(!is.na(n) & n < 10, "<10", as.character(n))]
  out[variable_type == "continuous" & n < 10,
    `:=`(n = NA_real_, N = NA_real_)] 

  #-----------------------------#
  # COMPLEMENTARY SUPPRESSION
  #-----------------------------#
  cat_out <- out[variable_type == "categorical"]
  if (nrow(cat_out) > 0) {
    suppressed_combos <- unique(cat_out[
      n_censored == "<10",
      .(variable, group)
    ])
    
    if (nrow(suppressed_combos) > 0) {
      suppressed_combos[, suppress_flag := TRUE]
      out <- merge(out, suppressed_combos, by = c("variable", "group"), all.x = TRUE)
      
      out[variable_type == "categorical" & !is.na(suppress_flag),
          `:=`(
            pct        = NA_real_,
            N          = NA_real_,
            n_censored = ifelse(n_censored == "<10", "<10", "suppressed"),
            n          = NA_real_
          )]
      
      out[, suppress_flag := NULL]
    }
  }

  cont_out <- out[variable_type == "continuous"]
  if (nrow(cont_out) > 0) {
    suppressed_cont_vars <- cont_out[
      n_censored == "<10",
      unique(variable)
    ]
    
    if (length(suppressed_cont_vars) > 0) {
      out[variable_type == "continuous" & variable %in% suppressed_cont_vars,
          `:=`(
            mean       = NA_real_,
            sd         = NA_real_,
            median     = NA_real_,
            p25        = NA_real_,
            p75        = NA_real_,
            N          = NA_real_,
            n_censored = ifelse(n_censored == "<10", "<10", "suppressed"),
            n          = NA_real_
          )]
    }
  }
  
  return(out[])
}

make_table1_comparisons <- function(dt, by, cont_vars = NULL, cat_vars = NULL,
                                    site_name = NA_character_, group_levels = NULL,
                                    reference_level = NULL, fisher_if_expected_lt5 = TRUE) {
  dt <- as.data.table(copy(dt))
  
  if (!(by %in% names(dt))) stop(sprintf("Grouping variable '%s' not found in dt.", by))
  dt <- dt[!is.na(get(by))]
  dt[, (by) := as.character(get(by))]
  
  observed_groups <- unique(dt[[by]])
  observed_groups <- observed_groups[!is.na(observed_groups)]
  
  if (!is.null(reference_level)) {
    if (!reference_level %in% observed_groups)
      stop(sprintf("reference_level '%s' not found in data.", reference_level))
    ref <- reference_level
    comparators <- if (!is.null(group_levels)) group_levels[group_levels != ref] else sort(observed_groups[observed_groups != ref])
  } else {
    if (!is.null(group_levels)) {
      if (length(group_levels) != 2) stop("group_levels must have length 2.")
      ref <- group_levels[1]; comparators <- group_levels[2]
    } else {
      if (length(observed_groups) != 2)
        stop(sprintf("Found %s groups — supply reference_level for 3+ groups.", paste(observed_groups, collapse = ", ")))
      grps <- sort(observed_groups); ref <- grps[1]; comparators <- grps[2]
    }
  }

  #--------------------------------------------------#
  # Helper: multilevel categorical SMD
  #--------------------------------------------------#
  calc_multilevel_smd <- function(x, group, g1, g2) {
    dsub <- data.table(x = as.character(x), group = as.character(group))
    dsub <- dsub[!is.na(group) & group %in% c(g1, g2)]
    dsub[is.na(x), x := "missing"]
    levs <- sort(unique(dsub$x))
    if (length(levs) < 2) return(list(smd = NA_real_, smd_se = NA_real_))
    ref_lev <- levs[1]
    keep_levs <- levs[levs != ref_lev]
    make_indicator_matrix <- function(vals, levels_keep) {
      out <- sapply(levels_keep, function(lv) as.numeric(vals == lv))
      if (is.null(dim(out))) { out <- matrix(out, ncol = 1); colnames(out) <- levels_keep }
      out
    }
    x1 <- dsub[group == g1, x]; x2 <- dsub[group == g2, x]
    n1 <- length(x1); n2 <- length(x2)
    if (n1 < 2 || n2 < 2) return(list(smd = NA_real_, smd_se = NA_real_))
    M1 <- make_indicator_matrix(x1, keep_levs)
    M2 <- make_indicator_matrix(x2, keep_levs)
    p1 <- colMeans(M1); p2 <- colMeans(M2); d <- p2 - p1
    S1 <- stats::cov(M1); S2 <- stats::cov(M2)
    if (is.null(dim(S1))) S1 <- matrix(S1, nrow = 1, ncol = 1)
    if (is.null(dim(S2))) S2 <- matrix(S2, nrow = 1, ncol = 1)
    S_pooled <- (S1 + S2) / 2
    S_inv <- try(solve(S_pooled), silent = TRUE)
    if (inherits(S_inv, "try-error")) {
      if (!requireNamespace("MASS", quietly = TRUE)) return(list(smd = NA_real_, smd_se = NA_real_))
      S_inv <- try(MASS::ginv(S_pooled), silent = TRUE)
      if (inherits(S_inv, "try-error")) return(list(smd = NA_real_, smd_se = NA_real_))
    }
    smd_val <- as.numeric(sqrt(t(d) %*% S_inv %*% d))
    # Yang & Dalton (2012) variance approximation for multivariate SMD:
    #   Var(d_M) ≈ (n1+n2)/(n1*n2) * k + d_M^2 / (2*(n1+n2))
    # where k is the number of non-reference categories.
    k <- length(keep_levs)
    smd_se_val <- sqrt((n1 + n2) / (n1 * n2) * k + smd_val^2 / (2 * (n1 + n2)))
    list(smd = smd_val, smd_se = smd_se_val)
  }

  #--------------------------------------------------#
  # Helper: continuous variable comparison
  #--------------------------------------------------#
  compare_continuous <- function(x, group, var_name, g1, g2, comparison_label) {
    dsub <- data.table(x = x, group = group)
    dsub <- dsub[!is.na(group)]
    
    x1 <- dsub[group == g1, x]
    x2 <- dsub[group == g2, x]
    x1_nonmiss <- x1[!is.na(x1)]
    x2_nonmiss <- x2[!is.na(x2)]
    
    n1 <- length(x1_nonmiss); n2 <- length(x2_nonmiss)
    mean1 <- if (n1 > 0) mean(x1_nonmiss) else NA_real_
    mean2 <- if (n2 > 0) mean(x2_nonmiss) else NA_real_
    sd1   <- if (n1 > 1) sd(x1_nonmiss) else NA_real_
    sd2   <- if (n2 > 1) sd(x2_nonmiss) else NA_real_
    med1  <- if (n1 > 0) median(x1_nonmiss) else NA_real_
    med2  <- if (n2 > 0) median(x2_nonmiss) else NA_real_
    p25_1 <- if (n1 > 0) as.numeric(quantile(x1_nonmiss, 0.25, na.rm = TRUE)) else NA_real_
    p25_2 <- if (n2 > 0) as.numeric(quantile(x2_nonmiss, 0.25, na.rm = TRUE)) else NA_real_
    p75_1 <- if (n1 > 0) as.numeric(quantile(x1_nonmiss, 0.75, na.rm = TRUE)) else NA_real_
    p75_2 <- if (n2 > 0) as.numeric(quantile(x2_nonmiss, 0.75, na.rm = TRUE)) else NA_real_
    
    estimate <- NA_real_; ci_lower <- NA_real_
    ci_upper <- NA_real_; p_value  <- NA_real_
    test_used <- NA_character_
    
    if (n1 >= 2 && n2 >= 2) {
      tt <- try(t.test(x2_nonmiss, x1_nonmiss), silent = TRUE)
      if (!inherits(tt, "try-error")) {
        estimate  <- unname(mean2 - mean1)
        ci_lower  <- unname(tt$conf.int[1])
        ci_upper  <- unname(tt$conf.int[2])
        p_value   <- tt$p.value
        test_used <- "welch_t_test"
      }
    }
    
    pooled_sd <- NA_real_; smd <- NA_real_; smd_se <- NA_real_
    if (n1 >= 2 && n2 >= 2 && !is.na(sd1) && !is.na(sd2)) {
      pooled_sd <- sqrt(((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2) / (n1 + n2 - 2))
      if (!is.na(pooled_sd) && pooled_sd > 0) {
        smd <- (mean2 - mean1) / pooled_sd
        # Hedges large-sample SE (this is the form metafor::escalc uses)
        smd_se <- sqrt((n1 + n2) / (n1 * n2) + smd^2 / (2 * (n1 + n2)))
      }
    }
    
    data.table(
      site = site_name, variable = var_name, variable_type = "continuous",
      comparison = comparison_label, level = NA_character_,
      group1 = g1, group2 = g2, group1_n = n1, group2_n = n2,
      group1_mean = mean1, group2_mean = mean2,
      group1_sd = sd1, group2_sd = sd2,
      group1_median = med1, group2_median = med2,
      group1_p25 = p25_1, group2_p25 = p25_2,
      group1_p75 = p75_1, group2_p75 = p75_2,
      statistic = "mean_difference", estimate = estimate,
      ci_lower = ci_lower, ci_upper = ci_upper, p_value = p_value,
      smd = smd, smd_se = smd_se, effect_measure = "mean_difference",
      test = test_used, note = NA_character_
    )
  }

  #--------------------------------------------------#
  # Helper: categorical variable comparison
  #--------------------------------------------------#
  compare_categorical <- function(x, group, var_name, g1, g2, comparison_label) {
    dsub <- data.table(x = x, group = group)
    dsub <- dsub[!is.na(group)]
    dsub[, x := as.character(x)]
    dsub[is.na(x), x := "missing"]
    
    tab <- table(dsub$x, dsub$group)
    needed_cols <- c(g1, g2)
    if (!all(needed_cols %in% colnames(tab))) {
      full_tab <- matrix(0, nrow = nrow(tab), ncol = 2,
                         dimnames = list(rownames(tab), needed_cols))
      common_cols <- intersect(colnames(tab), needed_cols)
      full_tab[, common_cols] <- tab[, common_cols, drop = FALSE]
      tab <- full_tab
    } else {
      tab <- tab[, needed_cols, drop = FALSE]
    }
    
    n_levels <- nrow(tab)
    n1 <- sum(tab[, g1])
    n2 <- sum(tab[, g2])
    
    overall_smd_result <- calc_multilevel_smd(dsub$x, dsub$group, g1, g2)
    overall_smd    <- overall_smd_result$smd
    overall_smd_se <- overall_smd_result$smd_se
    
    p_value <- NA_real_; test_used <- NA_character_
    if (n_levels >= 1) {
      chisq_try <- try(chisq.test(tab, correct = FALSE), silent = TRUE)
      if (!inherits(chisq_try, "try-error")) {
        expected <- chisq_try$expected
        if (fisher_if_expected_lt5 && any(expected < 5)) {
          fisher_try <- try(fisher.test(tab), silent = TRUE)
          if (!inherits(fisher_try, "try-error")) {
            p_value <- fisher_try$p.value; test_used <- "fisher_exact"
          } else {
            p_value <- chisq_try$p.value; test_used <- "chi_square"
          }
        } else {
          p_value <- chisq_try$p.value; test_used <- "chi_square"
        }
      }
    }
    
    out <- data.table(
      site = site_name, variable = var_name, variable_type = "categorical",
      comparison = comparison_label, level = NA_character_,
      group1 = g1, group2 = g2, group1_n = n1, group2_n = n2,
      group1_mean = NA_real_, group2_mean = NA_real_,
      group1_sd = NA_real_, group2_sd = NA_real_,
      group1_median = NA_real_, group2_median = NA_real_,
      group1_p25 = NA_real_, group2_p25 = NA_real_,
      group1_p75 = NA_real_, group2_p75 = NA_real_,
      statistic = "overall_difference", estimate = NA_real_,
      ci_lower = NA_real_, ci_upper = NA_real_, p_value = p_value,
      smd = overall_smd, smd_se = overall_smd_se, effect_measure = "categorical_smd",
      test = test_used, note = NA_character_
    )
    
    if (n_levels == 2) {
      levels_x <- rownames(tab)
      ref_level <- levels_x[1]; target_level <- levels_x[2]
      a <- as.numeric(tab[target_level, g2])
      b <- as.numeric(tab[ref_level,   g2])
      c <- as.numeric(tab[target_level, g1])
      d <- as.numeric(tab[ref_level,   g1])
      p1 <- if ((c + d) > 0) c / (c + d) else NA_real_
      p2 <- if ((a + b) > 0) a / (a + b) else NA_real_
      rd <- if (!is.na(p1) && !is.na(p2)) p2 - p1 else NA_real_
      rd_se <- if (!is.na(p1) && !is.na(p2) && (c + d) > 0 && (a + b) > 0)
        sqrt((p1 * (1 - p1) / (c + d)) + (p2 * (1 - p2) / (a + b))) else NA_real_
      rd_lcl <- if (!is.na(rd_se)) rd - 1.96 * rd_se else NA_real_
      rd_ucl <- if (!is.na(rd_se)) rd + 1.96 * rd_se else NA_real_
      pooled_p <- if (!is.na(p1) && !is.na(p2)) (p1 + p2) / 2 else NA_real_
      smd_binary <- if (!is.na(pooled_p) && pooled_p > 0 && pooled_p < 1)
        (p2 - p1) / sqrt(pooled_p * (1 - pooled_p)) else NA_real_
      # Hedges-style SE for the binary SMD. n1=(c+d), n2=(a+b) here.
      n1_bin <- c + d; n2_bin <- a + b
      smd_binary_se <- if (!is.na(smd_binary) && n1_bin > 1 && n2_bin > 1)
        sqrt((n1_bin + n2_bin) / (n1_bin * n2_bin) +
               smd_binary^2 / (2 * (n1_bin + n2_bin))) else NA_real_
      
      out_rd <- data.table(
        site = site_name, variable = var_name, variable_type = "categorical",
        comparison = comparison_label, level = target_level,
        group1 = g1, group2 = g2, group1_n = n1, group2_n = n2,
        group1_mean = p1, group2_mean = p2,
        group1_sd = NA_real_, group2_sd = NA_real_,
        group1_median = NA_real_, group2_median = NA_real_,
        group1_p25 = NA_real_, group2_p25 = NA_real_,
        group1_p75 = NA_real_, group2_p75 = NA_real_,
        statistic = "risk_difference", estimate = rd,
        ci_lower = rd_lcl, ci_upper = rd_ucl, p_value = p_value,
        smd = smd_binary, smd_se = smd_binary_se, effect_measure = "risk_difference",
        test = test_used,
        note = sprintf("Binary categorical variable. level='%s' treated as event.", target_level)
      )
      out <- rbind(out, out_rd, fill = TRUE)
      
    } else if (n_levels > 2) {
      out[, note := "Multi-level categorical variable: overall p-value and omnibus SMD reported; CI not populated."]
    }
    
    out
  }

  #--------------------------------------------------#
  # Loop over pairwise comparisons
  #--------------------------------------------------#
  all_comparisons <- lapply(comparators, function(comp) {
    g1 <- ref
    g2 <- comp
    comparison_label <- paste0(by, ": ", g2, " vs ", g1)
    dt_pair <- dt[get(by) %in% c(g1, g2)]
    out_list <- list()
    
    if (!is.null(cont_vars)) {
      for (v in cont_vars) {
        if (!(v %in% names(dt_pair))) { warning(sprintf("'%s' not found. Skipping.", v)); next }
        out_list[[paste0("cont_", v)]] <- compare_continuous(dt_pair[[v]], dt_pair[[by]], v, g1, g2, comparison_label)
      }
    }
    
    if (!is.null(cat_vars)) {
      for (v in cat_vars) {
        if (!(v %in% names(dt_pair))) { warning(sprintf("'%s' not found. Skipping.", v)); next }
        out_list[[paste0("cat_", v)]] <- compare_categorical(dt_pair[[v]], dt_pair[[by]], v, g1, g2, comparison_label)
      }
    }
    
    rbindlist(out_list, fill = TRUE, use.names = TRUE)
  })

  out <- rbindlist(all_comparisons, fill = TRUE, use.names = TRUE)
  
  #-----------------------------#
  # CENSORING
  #-----------------------------#
  # Add censored count columns before suppressing raw values
  out[, group1_n_censored := ifelse(group1_n < 10, "<10", as.character(group1_n))]
  out[, group2_n_censored := ifelse(group2_n < 10, "<10", as.character(group2_n))]
  
  # Suppress all statistics when either group is below threshold
  out[group1_n < 10 | group2_n < 10,
      `:=`(
        estimate      = NA_real_,
        ci_lower      = NA_real_,
        ci_upper      = NA_real_,
        p_value       = NA_real_,
        smd           = NA_real_,
        smd_se        = NA_real_,
        group1_mean   = NA_real_,
        group2_mean   = NA_real_,
        group1_sd     = NA_real_,
        group2_sd     = NA_real_,
        group1_median = NA_real_,
        group2_median = NA_real_,
        group1_p25    = NA_real_,
        group2_p25    = NA_real_,
        group1_p75    = NA_real_,
        group2_p75    = NA_real_,
        note          = "Suppressed: cell size < 10"
      )]
  
  # Suppress raw counts after creating censored versions
  out[group1_n < 10, group1_n := NA_real_]
  out[group2_n < 10, group2_n := NA_real_]

  #-----------------------------#
  # COMPLEMENTARY SUPPRESSION
  #-----------------------------#
  # In the comparisons table, the structure is one row per variable-comparison pair
  # (not per level), so if either group_n is suppressed we suppress the whole
  # variable-comparison combination, which the censoring block above already does.
  # The additional step here suppresses variables entirely if they appear in ANY
  # suppressed comparison, preventing back-calculation across comparisons.

  suppressed_vars <- out[
    group1_n_censored == "<10" | group2_n_censored == "<10",
    unique(variable)
  ]

  if (length(suppressed_vars) > 0) {
    out[variable %in% suppressed_vars,
        `:=`(
          estimate      = NA_real_,
          ci_lower      = NA_real_,
          ci_upper      = NA_real_,
          p_value       = NA_real_,
          smd           = NA_real_,
          group1_mean   = NA_real_,
          group2_mean   = NA_real_,
          group1_sd     = NA_real_,
          group2_sd     = NA_real_,
          group1_median = NA_real_,
          group2_median = NA_real_,
          group1_p25    = NA_real_,
          group2_p25    = NA_real_,
          group1_p75    = NA_real_,
          group2_p75    = NA_real_,
          group1_n_censored = ifelse(group1_n_censored == "<10", "<10", "suppressed"),
          group2_n_censored = ifelse(group2_n_censored == "<10", "<10", "suppressed"),
          group1_n      = NA_real_,
          group2_n      = NA_real_,
          note          = ifelse(
            is.na(note) | note == "",
            "Suppressed: complementary cell size < 10",
            note
          )
        )]
  }

  desired_order <- c(
    "site", "variable", "variable_type", "comparison", "level",
    "group1", "group2", "group1_n", "group1_n_censored",
    "group2_n", "group2_n_censored",
    "group1_mean", "group2_mean", "group1_sd", "group2_sd",
    "group1_median", "group2_median", "group1_p25", "group2_p25",
    "group1_p75", "group2_p75",
    "statistic", "estimate", "ci_lower", "ci_upper",
    "p_value", "smd", "smd_se", "effect_measure", "test", "note"
  )
  existing_order <- desired_order[desired_order %in% names(out)]
  setcolorder(out, existing_order)
  
  return(out[])
}

################################################################################
# Initialize a global log for skipped/failed table1s (with project column)
################################################################################
table1_log <- data.table(
  project         = character(),
  table_name      = character(),
  cohort_name     = character(),
  by_var          = character(),
  table_type      = character(),
  n_rows          = integer(),
  n_groups        = integer(),
  status          = character(),
  reason          = character(),
  output_file     = character()
)

################################################################################
# Wrapper: descriptive table1 with safety checks and logging
################################################################################
make_table1_safe <- function(dt,
                             cat_vars,
                             cont_vars,
                             by,
                             site_name,
                             output_path,
                             table_name,
                             cohort_name,
                             project = NA_character_) {                    # NEW
  
  if (is.null(dt) || nrow(dt) == 0) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "descriptive", n_rows = 0L, n_groups = 0L,
      status = "skipped_empty", reason = "Cohort has 0 rows",
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " table1: cohort is empty")
    return(NULL)
  }
  
  if (!by %in% names(dt)) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "descriptive", n_rows = nrow(dt), n_groups = 0L,
      status = "failed", reason = paste0("by variable '", by, "' not in data"),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " table1: by variable '", by, "' not in data")
    return(NULL)
  }
  
  n_groups <- length(unique(dt[[by]][!is.na(dt[[by]])]))
  if (n_groups < 1) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "descriptive", n_rows = nrow(dt), n_groups = 0L,
      status = "skipped_empty",
      reason = paste0("by variable '", by, "' has no non-NA values"),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " table1: by variable has no non-NA values")
    return(NULL)
  }
  
  result <- tryCatch({
    tbl <- make_table1_federated(
      dt = dt, cat_vars = cat_vars, cont_vars = cont_vars,
      by = by, site_name = site_name
    )
    fwrite(tbl, output_path)
    
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "descriptive", n_rows = nrow(dt), n_groups = n_groups,
      status = "success", reason = NA_character_,
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    tbl
    
  }, error = function(e) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "descriptive", n_rows = nrow(dt), n_groups = n_groups,
      status = "failed", reason = conditionMessage(e),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " table1: ", conditionMessage(e))
    return(NULL)
  })
  
  return(result)
}

################################################################################
# Wrapper: comparison table1 with safety checks and logging
################################################################################
make_table1_comparisons_safe <- function(dt,
                                         by,
                                         reference_level,
                                         cont_vars,
                                         cat_vars,
                                         site_name,
                                         output_path,
                                         table_name,
                                         cohort_name,
                                         project = NA_character_) {        # NEW
  
  if (is.null(dt) || nrow(dt) == 0) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = 0L, n_groups = 0L,
      status = "skipped_empty", reason = "Cohort has 0 rows",
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " comparisons: cohort is empty")
    return(NULL)
  }
  
  if (!by %in% names(dt)) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = nrow(dt), n_groups = 0L,
      status = "failed", reason = paste0("by variable '", by, "' not in data"),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " comparisons: by variable '", by, "' not in data")
    return(NULL)
  }
  
  observed_groups <- unique(as.character(dt[[by]]))
  observed_groups <- observed_groups[!is.na(observed_groups)]
  n_groups <- length(observed_groups)
  
  if (n_groups < 2) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = nrow(dt), n_groups = n_groups,
      status = "skipped_empty",
      reason = paste0("Only ", n_groups, " group(s) present; need >=2 for comparisons"),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " comparisons: only ", n_groups,
            " group(s) in '", by, "'")
    return(NULL)
  }
  
  if (!reference_level %in% observed_groups) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = nrow(dt), n_groups = n_groups,
      status = "skipped_empty",
      reason = paste0("Reference level '", reference_level,
                      "' not present in data. Observed: ",
                      paste(observed_groups, collapse = ", ")),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " comparisons: reference level '",
            reference_level, "' not in data")
    return(NULL)
  }
  
  result <- tryCatch({
    tbl <- make_table1_comparisons(
      dt = dt, by = by, reference_level = reference_level,
      cont_vars = cont_vars, cat_vars = cat_vars, site_name = site_name
    )
    fwrite(tbl, output_path)
    
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = nrow(dt), n_groups = n_groups,
      status = "success", reason = NA_character_,
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    tbl
    
  }, error = function(e) {
    log_row <- data.table(
      project = project,                                                   # NEW
      table_name = table_name, cohort_name = cohort_name, by_var = by,
      table_type = "comparisons", n_rows = nrow(dt), n_groups = n_groups,
      status = "failed", reason = conditionMessage(e),
      output_file = output_path
    )
    table1_log <<- rbind(table1_log, log_row, fill = TRUE)
    message("Skipping ", table_name, " comparisons: ", conditionMessage(e))
    return(NULL)
  })
  
  return(result)
}

################################################################################
# Combined wrapper: build descriptive + comparison table1s in one call
################################################################################
make_table1_pair <- function(dt,
                             by,
                             reference_level,
                             cat_vars,
                             cont_vars,
                             site_name,
                             tables_dir,
                             institution_prefix,
                             cohort_name,
                             file_stem,
                             project = NA_character_) {                    # NEW
  
  desc_path <- file.path(tables_dir,
                         paste0(institution_prefix, "_table1_", file_stem, "_descriptive.csv"))
  comp_path <- file.path(tables_dir,
                         paste0(institution_prefix, "_table1_", file_stem, "_comparisons.csv"))
  
  desc <- make_table1_safe(
    dt = dt, cat_vars = cat_vars, cont_vars = cont_vars, by = by,
    site_name = site_name, output_path = desc_path,
    table_name = paste0(file_stem, "_descriptive"),
    cohort_name = cohort_name,
    project = project                                                      # NEW
  )
  
  comp <- make_table1_comparisons_safe(
    dt = dt, by = by, reference_level = reference_level,
    cont_vars = cont_vars, cat_vars = cat_vars, site_name = site_name,
    output_path = comp_path,
    table_name = paste0(file_stem, "_comparisons"),
    cohort_name = cohort_name,
    project = project                                                      # NEW
  )
  
  invisible(list(descriptive = desc, comparisons = comp))
}
```

# Function to Run LM and GLM Models
```{r}
################################################################################
# Marginal standardization (g-computation) for binary GLM outcomes
# Returns marginal predicted probabilities under each exposure level and
# their difference (risk difference) with bootstrap CI.
# This is the correct estimand for meta-analyzing binary outcomes across sites.
################################################################################
extract_glm_marginal <- function(model,
                                 data,
                                 exposure      = "lang_cat_6",
                                 ref_level     = "english",
                                 target_level  = "spanish",
                                 n_boot        = 500,
                                 seed          = 42,
                                 site_name     = NA_character_,
                                 model_name    = NA_character_,
                                 cohort_name   = NA_character_) {
  
  data_used <- as.data.table(copy(data))
  
  if (!exposure %in% names(data_used)) {
    warning("Exposure variable '", exposure, "' not found in data. Returning NULL.")
    return(NULL)
  }
  
  original_levels <- levels(data_used[[exposure]])
  if (is.null(original_levels)) {
    original_levels <- sort(unique(data_used[[exposure]]))
  }
  
  # Extract the model formula and family so we can refit on each bootstrap sample
  model_formula <- formula(model)
  model_family  <- family(model)
  model_weights <- model$prior.weights  # NULL for unweighted models
  
  # ---- Core marginal standardization on any dataset + fitted model ----------
  marginal_rd_from_model <- function(fit, dat) {
    dat_target <- copy(dat)
    dat_ref    <- copy(dat)
    dat_target[[exposure]] <- factor(target_level, levels = original_levels)
    dat_ref[[exposure]]    <- factor(ref_level,    levels = original_levels)
    p_target <- mean(predict(fit, newdata = dat_target, type = "response"), na.rm = TRUE)
    p_ref    <- mean(predict(fit, newdata = dat_ref,    type = "response"), na.rm = TRUE)
    list(p_target = p_target, p_ref = p_ref, rd = p_target - p_ref)
  }
  
  # ---- Point estimate using original model ----------------------------------
  point_est <- marginal_rd_from_model(model, data_used)
  
  # ---- Bootstrap: refit model on each resample then standardize -------------
  set.seed(seed)
  n <- nrow(data_used)
  
  boot_rds <- vapply(seq_len(n_boot), function(i) {
    boot_idx <- sample(n, n, replace = TRUE)
    boot_dat <- data_used[boot_idx]
    
    # Refit the model on the bootstrap resample
    boot_fit <- tryCatch({
      if (is.null(model_weights)) {
        do.call(glm, list(formula = model_formula, data = boot_dat,
                          family = model_family))
      } else {
        do.call(glm, list(formula = model_formula, data = boot_dat,
                          family = model_family,
                          weights = model_weights[boot_idx]))
      }
    }, error = function(e) NULL)
    
    if (is.null(boot_fit)) return(NA_real_)
    
    # Standardize using the bootstrap-fitted model over the bootstrap sample
    tryCatch(
      marginal_rd_from_model(boot_fit, boot_dat)$rd,
      error = function(e) NA_real_
    )
  }, numeric(1))
  
  n_failed <- sum(is.na(boot_rds))
  if (n_failed > n_boot * 0.05) {
    warning(sprintf(
      "extract_glm_marginal: %d of %d bootstrap iterations failed (%.1f%%).",
      n_failed, n_boot, 100 * n_failed / n_boot
    ))
  }
  
  boot_rds <- boot_rds[!is.na(boot_rds)]
  boot_se  <- sd(boot_rds)
  boot_ci  <- quantile(boot_rds, c(0.025, 0.975), na.rm = TRUE)
  
  data.table(
    site             = site_name,
    model_name       = model_name,
    cohort_name      = cohort_name,
    exposure         = exposure,
    reference_level  = ref_level,
    target_level     = target_level,
    n_obs            = n,
    n_boot_attempted = n_boot,
    n_boot_succeeded = length(boot_rds),
    p_ref            = point_est$p_ref,
    p_target         = point_est$p_target,
    risk_difference  = point_est$rd,
    rd_se            = boot_se,
    rd_ci_lower      = unname(boot_ci[1]),
    rd_ci_upper      = unname(boot_ci[2]),
    rd_variance      = boot_se^2
  )
}
################################################################################
# Updated extract_glm_results — keeps the original OR output for internal
# use but adds a flag so callers know to use marginal results for publication
################################################################################
extract_glm_results <- function(model) {
  coef_table <- summary(model)$coefficients
  odds_ratios <- exp(coef_table[, "Estimate"])
  data.frame(
    Variable   = rownames(coef_table),
    Odds_Ratio = odds_ratios,
    Std_Error  = coef_table[, "Std. Error"],
    Variance   = coef_table[, "Std. Error"]^2,
    P_Value    = coef_table[, "Pr(>|z|)"]
  )
}

################################################################################
# extract_lm_results — unchanged
################################################################################
extract_lm_results <- function(model) {
  coef_table <- summary(model)$coefficients
  data.frame(
    term      = rownames(coef_table),
    estimate  = coef_table[, "Estimate"],
    std.error = coef_table[, "Std. Error"],
    variance  = coef_table[, "Std. Error"]^2,
    p.value   = coef_table[, "Pr(>|t|)"]
  )
}
################################################################################
# Initialize a global log for dropped terms (with project column)
################################################################################
model_dropped_terms_log <- data.table(
  project           = character(),
  model_name        = character(),
  outcome           = character(),
  cohort_name       = character(),
  n_obs             = integer(),
  exposure          = character(),
  dropped_terms     = character(),
  dropped_reason    = character(),
  exposure_dropped  = logical(),
  model_skipped     = logical(),
  error_message     = character()
)

################################################################################
# Helper: count non-NA unique levels of a variable
################################################################################
n_levels_nonNA <- function(x) length(unique(x[!is.na(x)]))

################################################################################
# Helper: extract the variable name from a term like "factor(hospital_icu_id)"
# or "race_category" or "lang_cat_6 * race_category"
################################################################################
extract_var_names <- function(term) {
  # Handle factor() wrapper
  cleaned <- gsub("factor\\(([^)]+)\\)", "\\1", term)
  # Handle interactions (split on *, :, and trim)
  parts <- unlist(strsplit(cleaned, "\\s*[\\*:]\\s*"))
  trimws(parts)
}

################################################################################
# Main wrapper: fit a model, dropping zero-variance covariates as needed
################################################################################
fit_model_safe <- function(outcome,
                           exposure = NULL,
                           covariates,
                           data,
                           model_name,
                           cohort_name,
                           project    = NA_character_,   # NEW
                           family     = NULL,
                           weights    = NULL,
                           min_levels = 2) {
  
  # Step 1: identify which covariates have insufficient variance
  drop <- character(0)
  drop_reasons <- character(0)
  
  for (term in covariates) {
    vars <- extract_var_names(term)
    for (v in vars) {
      if (!v %in% names(data)) {
        drop <- c(drop, term)
        drop_reasons <- c(drop_reasons, paste0(term, ": variable not in data"))
        break
      }
      if (n_levels_nonNA(data[[v]]) < min_levels) {
        drop <- c(drop, term)
        drop_reasons <- c(drop_reasons,
                          paste0(term, ": <", min_levels, " unique non-NA levels"))
        break
      }
    }
  }
  drop <- unique(drop)
  
  # Step 2: check the exposure itself (if one was provided)
  exposure_dropped <- FALSE
  if (!is.null(exposure)) {
    exposure_vars <- extract_var_names(exposure)
    for (v in exposure_vars) {
      if (!v %in% names(data) || n_levels_nonNA(data[[v]]) < min_levels) {
        exposure_dropped <- TRUE
        break
      }
    }
  }
  
  # Step 3: if exposure was provided but is degenerate, log and skip
  if (exposure_dropped) {
    log_row <- data.table(
      project          = project,
      model_name       = model_name,
      outcome          = outcome,
      cohort_name      = cohort_name,
      n_obs            = nrow(data),
      exposure         = if (is.null(exposure)) NA_character_ else exposure,
      dropped_terms    = paste(drop, collapse = "; "),
      dropped_reason   = paste(drop_reasons, collapse = "; "),
      exposure_dropped = TRUE,
      model_skipped    = TRUE,
      error_message    = paste0("Exposure '", exposure, "' has <", min_levels, " levels")
    )
    model_dropped_terms_log <<- rbind(model_dropped_terms_log, log_row, fill = TRUE)
    message("Skipping ", model_name, " model: exposure '", exposure,
            "' has <", min_levels, " levels")
    return(NULL)
  }
  
  # Step 4: build formula. If no exposure was provided, use covariates only.
  rhs <- if (is.null(exposure)) {
    setdiff(covariates, drop)
  } else {
    c(exposure, setdiff(covariates, drop))
  }
  
  if (length(rhs) == 0) {
    message("Skipping ", model_name, " model: no surviving terms")
    return(NULL)
  }
  
  fml <- reformulate(rhs, response = outcome)
  
  if (length(drop) > 0) {
    message(model_name, ": dropping ", length(drop), " term(s): ",
            paste(drop, collapse = ", "))
  }
  
  # Step 5: fit
  fit_result <- tryCatch({
    if (is.null(family)) {
      if (is.null(weights)) {
        lm(fml, data = data)
      } else {
        lm(fml, data = data, weights = weights)
      }
    } else {
      # Use do.call so the family argument is evaluated immediately and stored
      # as a concrete object in the model. Without this, pscl::pR2() and other
      # downstream functions can't recover the family later.
      if (is.null(weights)) {
        do.call(glm, list(formula = fml, data = data, family = family))
      } else {
        do.call(glm, list(formula = fml, data = data, family = family, weights = weights))
      }
    }
  }, error = function(e) {
    log_row <- data.table(
      project          = project,                                          # NEW
      model_name       = model_name,
      outcome          = outcome,
      cohort_name      = cohort_name,
      n_obs            = nrow(data),
      exposure = if (is.null(exposure)) NA_character_ else exposure,
      dropped_terms    = paste(drop, collapse = "; "),
      dropped_reason   = paste(drop_reasons, collapse = "; "),
      exposure_dropped = FALSE,
      model_skipped    = TRUE,
      error_message    = conditionMessage(e)
    )
    model_dropped_terms_log <<- rbind(model_dropped_terms_log, log_row, fill = TRUE)
    message("Skipping ", model_name, " model: ", conditionMessage(e))
    return(NULL)
  })
  
  # Step 6: log success
  if (!is.null(fit_result)) {
    log_row <- data.table(
      project          = project,
      model_name       = model_name,
      outcome          = outcome,
      cohort_name      = cohort_name,
      n_obs            = nrow(data),
      exposure         = if (is.null(exposure)) NA_character_ else exposure,   # CHANGED
      dropped_terms    = if (length(drop) > 0) paste(drop, collapse = "; ") else NA_character_,
      dropped_reason   = if (length(drop_reasons) > 0) paste(drop_reasons, collapse = "; ") else NA_character_,
      exposure_dropped = FALSE,
      model_skipped    = FALSE,
      error_message    = NA_character_
    )
    model_dropped_terms_log <<- rbind(model_dropped_terms_log, log_row, fill = TRUE)
  }
  
  return(fit_result)
}

################################################################################
# Standard covariate sets for models
# Defined once so changes to the adjustment set propagate everywhere.
################################################################################
standard_covariates <- c(
  "race_category", "age_category", "sex_category", "bmi",
  "factor(hospital_icu_id)", "elixhauser_score", "SVM_SCORE",
  "laps2", "factor(observation_year)"
)

# For paralytic project models, which omit SVM_SCORE
paralytic_covariates <- c(
  "race_category", "ethnicity_category_final", "age_category", "bmi",
  "sex_category", "factor(hospital_icu_id)", "elixhauser_score",
  "laps2", "factor(observation_year)"
)

# For measurements project models, which use a different set
measurements_covariates <- c(
  "race_category", "age_category", "sex_category", "ethnicity_category_final",
  "bmi", "factor(hospital_icu_id)", "elixhauser_score", "laps2",
  "location_of_intubation", "time_from_begin_imv_to_icu_hours",
  "factor(observation_year)"
)

# For exploratory analyses in sedation project
exploratory_binary_covariates <- c(
  "race_category", "age_category", "sex_category", "bmi",
  "factor(icu_imv)", "elixhauser_score", "SVM_SCORE", "laps2",
  "factor(observation_year)"
)
```


# Build Language Categorizations
```{r}
# 1: English v Non-English
total_cohort[, lang_cat_1 := fifelse(
  language_category == "english", "english",
  fifelse(!is.na(language_category), "non-english", NA_character_)
)]

# 2: English v Spanish v all other non-English
total_cohort[, lang_cat_2 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  !is.na(language_category), "other non-english",
  default = NA_character_
)]

# 3: English v Spanish v third most frequent hospital system wide language - all others NA
# Define third_lang
third_lang <- language_data[
  !trimws(tolower(language_category)) %in% c("english", "spanish", "unknown or na"),
  language_category
][1]

# Standardize the level name for cross-site pooling
third_lang_label <- "third_lang_cat"

# Define the variable
total_cohort[, lang_cat_3 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category == third_lang, third_lang_label,    # <-- use label instead of third_lang
  default = NA_character_
)]

third_lang_crosswalk <- data.table(
  institution        = institution,
  third_lang_cat     = third_lang
)
fwrite(third_lang_crosswalk,
       file.path(cohort_characteristics, 
                 paste0(institution, "_third_lang_crosswalk.csv")))

# 4: English v Spanish v Chinese v Vietnamese (three non-English languages with the most evidence supporting the existence of health disparities) - all others NA
total_cohort[, lang_cat_4 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category == "chinese", "chinese",
  language_category == "vietnamese", "vietnamese",
  default = NA_character_
)]

# 5: English v Spanish v Broad Regional Groupings
# South Asian, East Asian, Southeast Asian, Sub-Saharan African, non-colonial European
# Others that do not fit are NA

# -----------------------------
# Define language groupings
# -----------------------------

south_asian_langs <- c(
  "gujarati",
  "hindi",
  "urdu",
  "punjabi",
  "bengali",
  "nepali, marathi, or other indic languages",
  "telugu",
  "tamil",
  "malayalam, kannada, or other dravidian languages"
)

east_asian_langs <- c(
  "chinese",
  "japanese",
  "korean"
)

southeast_asian_langs <- c(
  "vietnamese",
  "khmer",
  "thai, lao, or other tai-kadai languages"
)

other_asian_langs <- c(
  "other languages of asia",
  "tagalog",
  "ilocano, samoan, hawaiian, or other austronesian languages"
)

sub_saharan_african_langs <- c(
  "amharic, somali, or other afro-asiatic languages",
  "yoruba, twi, igbo, or other languages of western africa",
  "swahili or other languages of central, eastern, and southern africa"
)

continental_european_langs <- c(
  "german",
  "greek",
  "russian",
  "polish",
  "italian",
  "serbo-croatian",
  "ukrainian or other slavic languages",
  "yiddish, pennsylvania dutch, or other west germanic languages",
  "other european indo-european languages")
  
mena_langs <- c(
  "arabic",
  "persian",
  "hebrew",
  "armenian"
)

all_other_langs <- c(
  "french",
  "haitian creole",
  "portuguese",
  "other asian indo-european languages",
  "navajo",
  "other native languages of north america",
  "other and unspecified languages"
)

total_cohort[, lang_cat_5 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category %in% south_asian_langs, "south_asian",
  language_category %in% east_asian_langs, "east_asian",
  language_category %in% southeast_asian_langs, "southeast_asian",
  language_category %in% continental_european_langs, "continental_european_langs",
  language_category %in% mena_langs, "mena",
  default = NA_character_
)]

# 6: English v Spanish - all others NA
total_cohort[, lang_cat_6 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  default = NA_character_
)]

# Broad Regional without "other_asian" and "all_other_langs" groups?
total_cohort[, lang_cat_7 := fcase(
  language_category == "english", "english",
  language_category == "spanish", "spanish",
  language_category %in% south_asian_langs, "south_asian",
  language_category %in% east_asian_langs, "east_asian",
  language_category %in% southeast_asian_langs, "southeast_asian",
  language_category %in% other_asian_langs, "other_asian",
  language_category %in% sub_saharan_african_langs, "sub_saharan_african",
  language_category %in% continental_european_langs, "continental_european_langs",
  language_category %in% mena_langs, "mena",
  language_category %in% all_other_langs, "all_other_langs",
  default = NA_character_
)]

```

#IPW Calculation for 48-hours of observation
```{r}
################################################################################
# IPW Calculation for 48-hours of observation
################################################################################
ipw_outcome <- total_cohort[, .(
  reached_48h = as.integer(imv_duration_hours_final >= 48)
), by = hospitalization_id]

ipw_sensitivity_analysis <- merge(total_cohort, ipw_outcome, by = "hospitalization_id")

ipw_model <- glm(
  reached_48h ~ age_category + sex_category + race_category +
    ethnicity_category + laps2 + factor(hospital_icu_id) +
    elixhauser_score + factor(observation_year),
  data   = unique(ipw_sensitivity_analysis, by = "hospitalization_id"),
  family = binomial()
)

total_cohort[, prob_reach_48h := predict(ipw_model, newdata = total_cohort, type = "response")]
total_cohort[, ipw := 1 / prob_reach_48h]

# ---- IPW Model Diagnostics --------------------------------------------------
tryCatch({
  model_data <- unique(ipw_sensitivity_analysis, by = "hospitalization_id")
  model_data[, predicted_prob := predict(ipw_model, newdata = model_data, type = "response")]

  # C-statistic (AUC)
  roc_obj <- pROC::roc(model_data$reached_48h, model_data$predicted_prob, quiet = TRUE)
  c_stat  <- as.numeric(pROC::auc(roc_obj))
  roc_ci  <- as.numeric(pROC::ci.auc(roc_obj))

  # Hosmer-Lemeshow
  hl_test <- ResourceSelection::hoslem.test(
    model_data$reached_48h, model_data$predicted_prob, g = 10
  )

  # Event rate overall and by language group
  event_overall <- model_data[, .(
    n               = .N,
    n_reached_48h   = sum(reached_48h),
    pct_reached_48h = 100 * mean(reached_48h)
  )]

  event_by_lang <- model_data[!is.na(lang_cat_6), .(
    n               = .N,
    n_reached_48h   = sum(reached_48h),
    pct_reached_48h = 100 * mean(reached_48h)
  ), by = lang_cat_6]

  # Predicted probability distribution overall and by language group
  prob_overall <- model_data[, .(
    group         = "overall",
    mean_prob     = mean(predicted_prob),
    median_prob   = median(predicted_prob),
    sd_prob       = sd(predicted_prob),
    min_prob      = min(predicted_prob),
    max_prob      = max(predicted_prob),
    p1_prob       = quantile(predicted_prob, 0.01),
    p99_prob      = quantile(predicted_prob, 0.99),
    n_near_zero   = sum(predicted_prob < 0.05),
    pct_near_zero = 100 * mean(predicted_prob < 0.05)
  )]

  prob_by_lang <- model_data[!is.na(lang_cat_6), .(
    mean_prob     = mean(predicted_prob),
    median_prob   = median(predicted_prob),
    sd_prob       = sd(predicted_prob),
    min_prob      = min(predicted_prob),
    max_prob      = max(predicted_prob),
    p1_prob       = quantile(predicted_prob, 0.01),
    p99_prob      = quantile(predicted_prob, 0.99),
    n_near_zero   = sum(predicted_prob < 0.05),
    pct_near_zero = 100 * mean(predicted_prob < 0.05)
  ), by = lang_cat_6]
  setnames(prob_by_lang, "lang_cat_6", "group")

  prob_diagnostics <- rbind(prob_overall, prob_by_lang, fill = TRUE)
  prob_diagnostics[, site   := institution]
  prob_diagnostics[, cohort := "full_48hr_cohort"]

  # Compile model-level summary
  ipw_model_diagnostics <- data.table(
    site            = institution,
    cohort          = "full_48hr_cohort",
    n_total         = event_overall$n,
    n_reached_48h   = event_overall$n_reached_48h,
    pct_reached_48h = event_overall$pct_reached_48h,
    c_statistic     = c_stat,
    c_stat_ci_lb    = roc_ci[1],
    c_stat_ci_ub    = roc_ci[3],
    hl_chisq        = hl_test$statistic,
    hl_df           = hl_test$parameter,
    hl_pvalue       = hl_test$p.value,
    weight_type     = "unstabilized",
    weight_note     = "Stabilized weights not used due to instability at smaller federated sites"
  )

  event_by_lang[, site   := institution]
  event_by_lang[, cohort := "full_48hr_cohort"]

  fwrite(ipw_model_diagnostics,
         file.path(rerun_tables_dir,
                   paste0(institution, "_ipw_model_diagnostics_full_48hr.csv")))
  fwrite(prob_diagnostics,
         file.path(rerun_tables_dir,
                   paste0(institution, "_ipw_predicted_prob_diagnostics_full_48hr.csv")))
  fwrite(event_by_lang,
         file.path(rerun_tables_dir,
                   paste0(institution, "_ipw_event_rate_by_lang_full_48hr.csv")))

  message("IPW model diagnostics written for full_48hr_cohort")
  message(sprintf("  C-statistic: %.3f (%.3f - %.3f)", c_stat, roc_ci[1], roc_ci[3]))
  message(sprintf("  Hosmer-Lemeshow p-value: %.3f", hl_test$p.value))
  message(sprintf("  Predicted prob near zero (<0.05): %d (%.1f%%)",
                  prob_overall$n_near_zero, prob_overall$pct_near_zero))

}, error = function(e) {
  message("Skipping IPW model diagnostics: ", conditionMessage(e))
})
```


#Define Primary Cohort
```{r}
current_project <- "sedation"
primary_cohort_all_icus <- total_cohort[!is.na(lang_cat_6)]

################################################################################
# Define Primary Cohort
################################################################################
spanish_counts <- primary_cohort_all_icus[lang_cat_6 == "spanish", .N, by = hospital_icu_id]
all_hospitals <- unique(primary_cohort_all_icus$hospital_icu_id)
hospitals_with_spanish <- spanish_counts$hospital_icu_id
hospitals_to_remove <- setdiff(all_hospitals, hospitals_with_spanish)
primary_cohort <- primary_cohort_all_icus[!hospital_icu_id %in% hospitals_to_remove]

primary_cohort$language_category       <- relevel(as.factor(primary_cohort$language_category),       ref = "english")
primary_cohort$lang_cat_1              <- relevel(as.factor(primary_cohort$lang_cat_1),              ref = "english")
primary_cohort$lang_cat_2              <- relevel(as.factor(primary_cohort$lang_cat_2),              ref = "english")
primary_cohort$lang_cat_3              <- relevel(as.factor(primary_cohort$lang_cat_3),              ref = "english")
primary_cohort$lang_cat_4              <- relevel(as.factor(primary_cohort$lang_cat_4),              ref = "english")
primary_cohort$lang_cat_5              <- relevel(as.factor(primary_cohort$lang_cat_5),              ref = "english")
primary_cohort$lang_cat_6              <- relevel(as.factor(primary_cohort$lang_cat_6),              ref = "english")
primary_cohort$sex_category            <- relevel(as.factor(primary_cohort$sex_category),            ref = "male")
primary_cohort$race_category           <- relevel(as.factor(primary_cohort$race_category),           ref = "white")
primary_cohort$ethnicity_category      <- relevel(as.factor(primary_cohort$ethnicity_category),      ref = "non-hispanic")
#primary_cohort$ethnicity_category_final <- relevel(as.factor(primary_cohort$ethnicity_category_final), ref = "non-hispanic/unknown")
primary_cohort$continuous_nmba_48      <- relevel(as.factor(primary_cohort$continuous_nmba_48),      ref = "0")
primary_cohort$icu_imv <- relevel(as.factor(primary_cohort$icu_imv), ref = "medical_icu")

################################################################################
# Primary Cohort - Table 1s
################################################################################
make_table1_pair(
  dt = primary_cohort, by = "lang_cat_6", reference_level = "english",
  cat_vars = cat_vars, cont_vars = cont_vars, site_name = institution,
  tables_dir = rerun_tables_dir, institution_prefix = institution,
  cohort_name = "primary_cohort", file_stem = "rerun",
  project     = current_project
)
```

#IPW Diagnostics
```{r}
# Helper for the cohort-filtering subgroup pattern. Each subgroup needs:
#   1. Filter the primary cohort
#   2. Re-level factors (subset can change which levels are present)
#   3. Build descriptive + comparison table1s
#   4. Fit the standard sedation model
#   5. Write model output if it succeeded
relevel_subgroup <- function(dt) {
  if (is.null(dt) || nrow(dt) == 0) return(dt)
  dt$lang_cat_6   <- relevel(as.factor(dt$lang_cat_6),   ref = "english")
  dt$sex_category <- relevel(as.factor(dt$sex_category), ref = "male")
  dt$race_category <- relevel(as.factor(dt$race_category), ref = "white")
  dt$ethnicity_category <- relevel(as.factor(dt$ethnicity_category), ref = "non-hispanic")
  dt$ethnicity_category_final <- relevel(as.factor(dt$ethnicity_category_final), ref = "non-hispanic/unknown")
  dt
}

cat_vars <- c("sex_category", "race_category", "ethnicity_category", "age_category")

cont_vars <- c("laps2", "elixhauser_score")

################################################################################
# Full 48 Hours - IPW Weighted with Sandwich SEs
################################################################################
full_48hr_cohort <- primary_cohort[hours_mechanically_ventilated_72 >= 48]
full_48hr_cohort <- full_48hr_cohort[is.finite(ipw) & !is.na(ipw)]

# Track pre-winsorization N
n_pretrim <- nrow(full_48hr_cohort)

ipw_q <- quantile(full_48hr_cohort$ipw, c(0.01, 0.99), na.rm = TRUE)
full_48hr_cohort[, ipw := pmin(pmax(ipw, ipw_q[1]), ipw_q[2])]

full_48hr_cohort <- relevel_subgroup(full_48hr_cohort)

make_table1_pair(
  dt = full_48hr_cohort, by = "lang_cat_6", reference_level = "english",
  cat_vars = cat_vars, cont_vars = cont_vars, site_name = institution,
  tables_dir = sedation_tables_dir, institution_prefix = institution,
  cohort_name = "full_48hr_cohort", file_stem = "full_48hr_cohort",
  project     = current_project
)

sensitivity_model_full_48 <- fit_model_safe(
  outcome    = "percent_deep_sedation_48",
  exposure   = "lang_cat_6",
  covariates = standard_covariates,
  data       = full_48hr_cohort,
  weights    = full_48hr_cohort$ipw,
  model_name = "sensitivity_model_full_48",
  cohort_name = "full_48hr_cohort",
  project    = current_project
)

if (!is.null(sensitivity_model_full_48)) {
  tryCatch({
    robust_se <- tryCatch(
      vcovHC(sensitivity_model_full_48, type = "HC3"),
      error = function(e) NULL
    )
    if (is.null(robust_se) || anyNA(robust_se)) {
      warning("HC3 robust SE failed at this site; falling back to HC0.")
      robust_se <- vcovHC(sensitivity_model_full_48, type = "HC0")
    }

    sensitivity_model_full_48_sandwich <- coeftest(sensitivity_model_full_48, vcov = robust_se)

    sensitivity_model_full_48_results <- data.frame(
      term      = rownames(sensitivity_model_full_48_sandwich),
      estimate  = sensitivity_model_full_48_sandwich[, "Estimate"],
      std.error = sensitivity_model_full_48_sandwich[, "Std. Error"],
      variance  = sensitivity_model_full_48_sandwich[, "Std. Error"]^2,
      p.value   = sensitivity_model_full_48_sandwich[, "Pr(>|t|)"]
    )
    fwrite(sensitivity_model_full_48_results,
           file.path(sedation_sensitivity,
                     paste0(institution, "_sensitivity_analysis_full_48hr.csv")))
  }, error = function(e) {
    message("Skipping sensitivity_model_full_48 sandwich SE step: ", conditionMessage(e))
  })
}

# ---- IPW Weight Diagnostics (after model fit on final winsorized weights) ----
if (nrow(full_48hr_cohort) > 0 && !is.null(full_48hr_cohort$ipw)) {
  tryCatch({
    n_posttrim  <- nrow(full_48hr_cohort)
    n_winsorized <- full_48hr_cohort[
      ipw == ipw_q[1] | ipw == ipw_q[2], .N
    ]
    pct_winsorized <- 100 * n_winsorized / n_pretrim

    # Overall weight diagnostics
    ipw_overall <- full_48hr_cohort[, .(
      site           = institution,
      cohort         = "full_48hr_cohort",
      n_total        = n_pretrim,
      n_winsorized   = n_winsorized,
      pct_winsorized = pct_winsorized,
      trim_lower     = ipw_q[1],
      trim_upper     = ipw_q[2],
      mean_ipw       = mean(ipw, na.rm = TRUE),
      median_ipw     = median(ipw, na.rm = TRUE),
      sd_ipw         = sd(ipw, na.rm = TRUE),
      min_ipw        = min(ipw, na.rm = TRUE),
      max_ipw        = max(ipw, na.rm = TRUE),
      p25_ipw        = quantile(ipw, 0.25, na.rm = TRUE),
      p75_ipw        = quantile(ipw, 0.75, na.rm = TRUE),
      p99_ipw        = quantile(ipw, 0.99, na.rm = TRUE),
      ess            = (sum(ipw))^2 / sum(ipw^2),
      ess_pct        = 100 * ((sum(ipw))^2 / sum(ipw^2)) / .N
    )]

    # By-group weight diagnostics
    ipw_by_group <- full_48hr_cohort[!is.na(lang_cat_6), .(
      site           = institution,
      cohort         = "full_48hr_cohort",
      n              = .N,
      mean_ipw       = mean(ipw, na.rm = TRUE),
      median_ipw     = median(ipw, na.rm = TRUE),
      sd_ipw         = sd(ipw, na.rm = TRUE),
      min_ipw        = min(ipw, na.rm = TRUE),
      max_ipw        = max(ipw, na.rm = TRUE),
      p25_ipw        = quantile(ipw, 0.25, na.rm = TRUE),
      p75_ipw        = quantile(ipw, 0.75, na.rm = TRUE),
      p99_ipw        = quantile(ipw, 0.99, na.rm = TRUE),
      ess            = (sum(ipw))^2 / sum(ipw^2),
      ess_pct        = 100 * ((sum(ipw))^2 / sum(ipw^2)) / .N
    ), by = lang_cat_6]

    # ---- Covariate balance: continuous variables ----------------------------
    calc_smd_continuous <- function(x, group, w = NULL, g1 = "english", g2 = "spanish") {
      x1 <- x[group == g1]; x2 <- x[group == g2]
      w1 <- if (!is.null(w)) w[group == g1] else rep(1, length(x1))
      w2 <- if (!is.null(w)) w[group == g2] else rep(1, length(x2))
      x1 <- x1[!is.na(x1)]; x2 <- x2[!is.na(x2)]
      w1 <- w1[!is.na(x1)];  w2 <- w2[!is.na(x2)]
      if (length(x1) < 2 || length(x2) < 2) return(NA_real_)
      m1 <- weighted.mean(x1, w1); m2 <- weighted.mean(x2, w2)
      v1 <- sum(w1 * (x1 - m1)^2) / sum(w1)
      v2 <- sum(w2 * (x2 - m2)^2) / sum(w2)
      sd_pool <- sqrt((v1 + v2) / 2)
      if (sd_pool == 0) return(NA_real_)
      (m2 - m1) / sd_pool
    }

    # ---- Covariate balance: categorical variables ---------------------------
    calc_smd_categorical <- function(x, group, w = NULL, g1 = "english", g2 = "spanish") {
      # Multilevel categorical SMD using pooled covariance matrix (Yang & Dalton 2012)
      # For binary variables this reduces to the standard binary SMD
      dsub <- data.table(x = as.character(x), group = as.character(group),
                         w = if (!is.null(w)) w else rep(1, length(x)))
      dsub <- dsub[!is.na(group) & group %in% c(g1, g2)]
      dsub[is.na(x), x := "missing"]
      levs     <- sort(unique(dsub$x))
      if (length(levs) < 2) return(NA_real_)
      ref_lev  <- levs[1]
      keep_levs <- levs[levs != ref_lev]
      make_wt_props <- function(grp) {
        sub <- dsub[group == grp]
        if (nrow(sub) == 0) return(rep(NA_real_, length(keep_levs)))
        sapply(keep_levs, function(lv) weighted.mean(sub$x == lv, sub$w))
      }
      make_wt_cov <- function(grp) {
        sub <- dsub[group == grp]
        if (nrow(sub) < 2) return(matrix(NA_real_, length(keep_levs), length(keep_levs)))
        M <- sapply(keep_levs, function(lv) as.numeric(sub$x == lv))
        if (is.null(dim(M))) M <- matrix(M, ncol = 1)
        # Weighted covariance
        wts <- sub$w / sum(sub$w)
        wm  <- colMeans(M * wts * nrow(M))  # weighted means
        sweep_M <- sweep(M, 2, colMeans(sweep(M, 1, wts, "*") * nrow(M)))
        t(sweep_M * wts) %*% sweep_M * nrow(M)
      }
      p1 <- make_wt_props(g1); p2 <- make_wt_props(g2)
      if (any(is.na(p1)) || any(is.na(p2))) return(NA_real_)
      d <- p2 - p1
      S1 <- make_wt_cov(g1); S2 <- make_wt_cov(g2)
      S_pooled <- (S1 + S2) / 2
      S_inv <- tryCatch(solve(S_pooled), error = function(e) {
        tryCatch(MASS::ginv(S_pooled), error = function(e2) NULL)
      })
      if (is.null(S_inv)) return(NA_real_)
      as.numeric(sqrt(t(d) %*% S_inv %*% d))
    }

    balance_cont <- lapply(intersect(cont_vars, names(full_48hr_cohort)), function(v) {
  sub <- full_48hr_cohort[!is.na(lang_cat_6) & !is.na(get(v))]
  if (nrow(sub) == 0) return(NULL)
  smd_uw <- calc_smd_continuous(sub[[v]], sub$lang_cat_6, w = NULL)
  smd_w  <- calc_smd_continuous(sub[[v]], sub$lang_cat_6, w = sub$ipw)
  data.table(
    site             = institution,
    cohort           = "full_48hr_cohort",
    variable         = v,
    variable_type    = "continuous",
    smd_unweighted   = smd_uw,
    smd_weighted     = smd_w,
    smd_change       = smd_w - smd_uw,
    balance_improved = abs(smd_w) < abs(smd_uw)
  )
})

    balance_cat <- lapply(intersect(cat_vars, names(full_48hr_cohort)), function(v) {
  sub <- full_48hr_cohort[!is.na(lang_cat_6) & !is.na(get(v))]
  if (nrow(sub) == 0) return(NULL)
  smd_uw <- calc_smd_categorical(sub[[v]], sub$lang_cat_6, w = NULL)
  smd_w  <- calc_smd_categorical(sub[[v]], sub$lang_cat_6, w = sub$ipw)
  data.table(
    site             = institution,
    cohort           = "full_48hr_cohort",
    variable         = v,
    variable_type    = "categorical",
    smd_unweighted   = smd_uw,
    smd_weighted     = smd_w,
    smd_change       = smd_w - smd_uw,
    balance_improved = abs(smd_w) < abs(smd_uw)
  )
})

    balance_dt <- rbindlist(c(balance_cont, balance_cat), fill = TRUE, use.names = TRUE)

    # Summary: how many variables improved balance
    n_improved    <- sum(balance_dt$balance_improved, na.rm = TRUE)
    n_vars        <- sum(!is.na(balance_dt$balance_improved))
    n_below_thresh <- sum(abs(balance_dt$smd_weighted) < 0.1, na.rm = TRUE)

    balance_summary <- data.table(
      site                    = institution,
      cohort                  = "full_48hr_cohort",
      n_vars_assessed         = n_vars,
      n_vars_balance_improved = n_improved,
      n_vars_smd_below_0.1   = n_below_thresh,
      pct_vars_below_0.1      = 100 * n_below_thresh / n_vars,
      max_smd_unweighted      = max(abs(balance_dt$smd_unweighted), na.rm = TRUE),
      max_smd_weighted        = max(abs(balance_dt$smd_weighted),   na.rm = TRUE)
    )

    # ---- Weight density plot ------------------------------------------------
    tryCatch({
      plot_data <- full_48hr_cohort[!is.na(lang_cat_6), .(ipw, lang_cat_6)]
      p <- ggplot2::ggplot(plot_data, ggplot2::aes(x = ipw, fill = lang_cat_6)) +
        ggplot2::geom_density(alpha = 0.5) +
        ggplot2::geom_vline(xintercept = ipw_q, linetype = "dashed", color = "grey40") +
        ggplot2::scale_fill_manual(values = c("english" = "#4477AA", "spanish" = "#EE6677")) +
        ggplot2::labs(
          title = paste0(institution, ": IPW Weight Distribution (Full 48hr Cohort)"),
          x     = "Inverse Probability Weight",
          y     = "Density",
          fill  = "Language Group",
          caption = sprintf(
            "Dashed lines: winsorization bounds [%.3f, %.3f]. N winsorized: %d (%.1f%%)",
            ipw_q[1], ipw_q[2], n_winsorized, pct_winsorized
          )
        ) +
        ggplot2::theme_minimal()
      ggplot2::ggsave(
        file.path(sedation_sensitivity,
                  paste0(institution, "_ipw_weight_density_full_48hr.png")),
        plot = p, width = 8, height = 5, dpi = 300
      )
      message("  Weight density plot saved")
    }, error = function(e) {
      message("  Skipping weight density plot: ", conditionMessage(e))
    })

    # ---- Predicted probability overlap plot ---------------------------------
    tryCatch({
      model_data_plot <- unique(ipw_sensitivity_analysis, by = "hospitalization_id")
      model_data_plot[, predicted_prob := predict(
        ipw_model, newdata = model_data_plot, type = "response"
      )]
      plot_data_prob <- model_data_plot[!is.na(lang_cat_6),
                                        .(predicted_prob, lang_cat_6, reached_48h)]
      p2 <- ggplot2::ggplot(plot_data_prob,
                            ggplot2::aes(x = predicted_prob, fill = lang_cat_6)) +
        ggplot2::geom_density(alpha = 0.5) +
        ggplot2::geom_vline(xintercept = 0.05, linetype = "dotted", color = "red") +
        ggplot2::scale_fill_manual(values = c("english" = "#4477AA", "spanish" = "#EE6677")) +
        ggplot2::labs(
          title   = paste0(institution, ": Predicted P(IMV >= 48h) by Language Group"),
          x       = "Predicted Probability",
          y       = "Density",
          fill    = "Language Group",
          caption = "Red dotted line: positivity threshold (p < 0.05)"
        ) +
        ggplot2::theme_minimal()
      ggplot2::ggsave(
        file.path(sedation_sensitivity,
                  paste0(institution, "_ipw_predicted_prob_overlap_full_48hr.png")),
        plot = p2, width = 8, height = 5, dpi = 300
      )
      message("  Predicted probability overlap plot saved")
    }, error = function(e) {
      message("  Skipping predicted probability overlap plot: ", conditionMessage(e))
    })

    # ---- Write all weight diagnostics ---------------------------------------
    fwrite(ipw_overall,
           file.path(rerun_tables_dir,
                     paste0(institution, "_ipw_diagnostics_full_48hr_overall.csv")))
    fwrite(ipw_by_group,
           file.path(rerun_tables_dir,
                     paste0(institution, "_ipw_diagnostics_full_48hr_by_group.csv")))
    fwrite(balance_dt,
           file.path(rerun_tables_dir,
                     paste0(institution, "_ipw_balance_diagnostics_full_48hr.csv")))
    fwrite(balance_summary,
           file.path(rerun_tables_dir,
                     paste0(institution, "_ipw_balance_summary_full_48hr.csv")))

    message("IPW weight diagnostics written for full_48hr_cohort")
    message(sprintf("  Winsorized: %d (%.1f%%) at [%.3f, %.3f]",
                    n_winsorized, pct_winsorized, ipw_q[1], ipw_q[2]))
    message(sprintf("  Post-winsorization: mean=%.3f, median=%.3f, max=%.3f, ESS=%.0f (%.1f%%)",
                    ipw_overall$mean_ipw, ipw_overall$median_ipw,
                    ipw_overall$max_ipw, ipw_overall$ess, ipw_overall$ess_pct))
    message(sprintf("  Balance: %d/%d vars improved; %d/%d below SMD 0.1 threshold",
                    n_improved, n_vars, n_below_thresh, n_vars))

  }, error = function(e) {
    message("Skipping IPW weight diagnostics for full_48hr_cohort: ", conditionMessage(e))
  })
}
```


#ICU IMV and Hospital Type Tabulation
```{r}
lang_icu_counts <- primary_cohort[
  , .(count = .N),
  by = .(lang_cat_6, icu_imv)
][
  order(lang_cat_6, icu_imv)
]

lang_icu_counts

fwrite(lang_icu_counts,
       file.path(rerun_tables_dir, paste0(institution, "_icu_counts_table_1.csv")))


lang_hospital_type_counts <- primary_cohort[
  , .(count = .N),
  by = .(lang_cat_6, hospital_type)
][
  order(lang_cat_6, hospital_type)
]

lang_hospital_type_counts

fwrite(lang_hospital_type_counts,
       file.path(rerun_tables_dir, paste0(institution, "_hospital_type_counts_table_1.csv")))


```


#Risk Differences Re-Run for Binary Outcomes
```{r}
secondary_model_trach <- fit_model_safe(
  outcome     = "trached",
  exposure    = "lang_cat_6",
  covariates  = standard_covariates,
  data        = primary_cohort,
  model_name  = "secondary_model_trach",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(secondary_model_trach)) {
  fwrite(extract_glm_results(secondary_model_trach),
         file.path(rerun_final_data, paste0(institution, "_secondary_analysis_trach.csv")))
  
  fwrite(extract_glm_marginal(
    model        = secondary_model_trach,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "secondary_model_trach",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_secondary_analysis_trach_marginal.csv")))
}

secondary_model_28day_mortality <- fit_model_safe(
  outcome     = "mortality_28day",
  exposure    = "lang_cat_6",
  covariates  = standard_covariates,
  data        = primary_cohort,
  model_name  = "secondary_model_28day_mortality",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(secondary_model_28day_mortality)) {
  fwrite(extract_glm_results(secondary_model_28day_mortality),
         file.path(rerun_final_data, paste0(institution, "_secondary_analysis_28_day_mortality.csv")))
  fwrite(extract_glm_marginal(
    model        = secondary_model_28day_mortality,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "secondary_model_28day_mortality",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_secondary_analysis_28_day_mortality_marginal.csv")))
}

secondary_model_reintubation <- fit_model_safe(
  outcome     = "reintubation_final",
  exposure    = "lang_cat_6",
  covariates  = standard_covariates,
  data        = primary_cohort,
  model_name  = "secondary_model_reintubation",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(secondary_model_reintubation)) {
  fwrite(extract_glm_results(secondary_model_reintubation),
         file.path(rerun_final_data, paste0(institution, "_secondary_analysis_reintubation.csv")))
  fwrite(extract_glm_marginal(
    model        = secondary_model_reintubation,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "secondary_model_reintubation",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_secondary_analysis_reintubation_marginal.csv")))
}

secondary_model_inhospital_mortality <- fit_model_safe(
  outcome     = "death_during_admission",
  exposure    = "lang_cat_6",
  covariates  = standard_covariates,
  data        = primary_cohort,
  model_name  = "secondary_model_inhospital_mortality",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(secondary_model_inhospital_mortality)) {
  fwrite(extract_glm_results(secondary_model_inhospital_mortality),
         file.path(rerun_final_data, paste0(institution, "_secondary_analysis_in_hospital_mortality.csv")))
  fwrite(extract_glm_marginal(
    model        = secondary_model_inhospital_mortality,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "secondary_model_inhospital_mortality",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_secondary_analysis_in_hospital_mortality_marginal.csv")))
}


sensitivity_model_alt_outcome_definition_1 <- fit_model_safe(
  outcome     = "deep_sedation_weighted",
  exposure    = "lang_cat_6",
  covariates  = standard_covariates,
  data        = primary_cohort,
  model_name  = "sensitivity_model_weighted_deep_sedation",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(sensitivity_model_alt_outcome_definition_1)) {
  fwrite(extract_glm_results(sensitivity_model_alt_outcome_definition_1),
         file.path(rerun_final_data, paste0(institution, "_sensitivity_analysis_weighted_deep_sedation.csv")))
  fwrite(extract_glm_marginal(
    model        = sensitivity_model_alt_outcome_definition_1,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "sensitivity_model_weighted_deep_sedation",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_sensitivity_analysis_weighted_deep_sedation_marginal.csv")))
}

exploratory_model_benzo_odds <- fit_model_safe(
  outcome     = "cont_benzos_receipt",
  exposure    = "lang_cat_6",
  covariates  = exploratory_binary_covariates,
  data        = primary_cohort,
  model_name  = "exploratory_model_benzo_odds",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(exploratory_model_benzo_odds)) {
  fwrite(extract_glm_results(exploratory_model_benzo_odds),
         file.path(rerun_final_data, paste0(institution, "_exploratory_analysis_benzo_odds.csv")))
  fwrite(extract_glm_marginal(
    model        = exploratory_model_benzo_odds,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "exploratory_model_benzo_odds",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_exploratory_analysis_benzo_odds_marginal.csv")))
}

exploratory_model_opioid_only_analgesia <- fit_model_safe(
  outcome     = "cont_analgesia_only",
  exposure    = "lang_cat_6",
  covariates  = exploratory_binary_covariates,
  data        = primary_cohort,
  model_name  = "exploratory_model_opioid_only_analgesia",
  cohort_name = "primary_cohort",
  family      = binomial(link = "logit"),
  project     = current_project
)
if (!is.null(exploratory_model_opioid_only_analgesia)) {
  fwrite(extract_glm_results(exploratory_model_opioid_only_analgesia),
         file.path(rerun_final_data, paste0(institution, "_exploratory_analysis_opioid_only.csv")))
  fwrite(extract_glm_marginal(
    model        = exploratory_model_opioid_only_analgesia,
    data         = primary_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    site_name    = institution,
    model_name   = "exploratory_model_opioid_only_analgesia",
    cohort_name  = "primary_cohort"
  ), file.path(rerun_final_data,
               paste0(institution, "_exploratory_analysis_opioid_only_marginal.csv")))
}
```

#Rebuild Tables for Paralytics Re-Run
```{r}
relevel_paralytic_subgroup <- function(dt) {
  if (is.null(dt) || nrow(dt) == 0) return(dt)
  dt$sex_category       <- relevel(as.factor(dt$sex_category),       ref = "male")
  dt$race_category      <- relevel(as.factor(dt$race_category),      ref = "white")
  dt$ethnicity_category <- relevel(as.factor(dt$ethnicity_category), ref = "non-hispanic")
  dt$ethnicity_category_final <- relevel(as.factor(dt$ethnicity_category_final), ref = "non-hispanic/unknown")
  dt
}

lang_cat_3_cohort <- total_cohort[!is.na(lang_cat_3)]
lang_cat_3_cohort$language_category       <- relevel(as.factor(lang_cat_3_cohort$language_category),       ref = "english")
lang_cat_3_cohort$lang_cat_1              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_1),              ref = "english")
lang_cat_3_cohort$lang_cat_2              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_2),              ref = "english")
lang_cat_3_cohort$lang_cat_3              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_3),              ref = "english")
lang_cat_3_cohort$lang_cat_4              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_4),              ref = "english")
lang_cat_3_cohort$lang_cat_5              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_5),              ref = "english")
lang_cat_3_cohort$lang_cat_6              <- relevel(as.factor(lang_cat_3_cohort$lang_cat_6),              ref = "english")
lang_cat_3_cohort$sex_category            <- relevel(as.factor(lang_cat_3_cohort$sex_category),            ref = "male")
lang_cat_3_cohort$race_category           <- relevel(as.factor(lang_cat_3_cohort$race_category),           ref = "white")
lang_cat_3_cohort$ethnicity_category      <- relevel(as.factor(lang_cat_3_cohort$ethnicity_category),      ref = "non-hispanic")
lang_cat_3_cohort$ethnicity_category_final <- relevel(as.factor(lang_cat_3_cohort$ethnicity_category_final), ref = "non-hispanic/unknown")

lang_cat_4_cohort <- total_cohort[!is.na(lang_cat_4)]
lang_cat_4_cohort$language_category       <- relevel(as.factor(lang_cat_4_cohort$language_category),       ref = "english")
lang_cat_4_cohort$lang_cat_4              <- relevel(as.factor(lang_cat_4_cohort$lang_cat_4),              ref = "english")
lang_cat_4_cohort$sex_category            <- relevel(as.factor(lang_cat_4_cohort$sex_category),            ref = "male")
lang_cat_4_cohort$race_category           <- relevel(as.factor(lang_cat_4_cohort$race_category),           ref = "white")
lang_cat_4_cohort$ethnicity_category      <- relevel(as.factor(lang_cat_4_cohort$ethnicity_category),      ref = "non-hispanic")
lang_cat_4_cohort$ethnicity_category_final <- relevel(as.factor(lang_cat_4_cohort$ethnicity_category_final), ref = "non-hispanic/unknown")

lang_cat_5_cohort <- total_cohort[!is.na(lang_cat_5)]
lang_cat_5_cohort$language_category       <- relevel(as.factor(lang_cat_5_cohort$language_category),       ref = "english")
lang_cat_5_cohort$lang_cat_5              <- relevel(as.factor(lang_cat_5_cohort$lang_cat_5),              ref = "english")
lang_cat_5_cohort$sex_category            <- relevel(as.factor(lang_cat_5_cohort$sex_category),            ref = "male")
lang_cat_5_cohort$race_category           <- relevel(as.factor(lang_cat_5_cohort$race_category),           ref = "white")
lang_cat_5_cohort$ethnicity_category      <- relevel(as.factor(lang_cat_5_cohort$ethnicity_category),      ref = "non-hispanic")
lang_cat_5_cohort$ethnicity_category_final <- relevel(as.factor(lang_cat_5_cohort$ethnicity_category_final), ref = "non-hispanic/unknown")

primary_cohort_all_icus <- total_cohort[!is.na(lang_cat_6)]
primary_cohort_all_icus$language_category       <- relevel(as.factor(primary_cohort_all_icus$language_category),       ref = "english")
primary_cohort_all_icus$lang_cat_6              <- relevel(as.factor(primary_cohort_all_icus$lang_cat_6),              ref = "english")
primary_cohort_all_icus$sex_category            <- relevel(as.factor(primary_cohort_all_icus$sex_category),            ref = "male")
primary_cohort_all_icus$race_category           <- relevel(as.factor(primary_cohort_all_icus$race_category),           ref = "white")
primary_cohort_all_icus$ethnicity_category      <- relevel(as.factor(primary_cohort_all_icus$ethnicity_category),      ref = "non-hispanic")
primary_cohort_all_icus$ethnicity_category_final <- relevel(as.factor(primary_cohort_all_icus$ethnicity_category_final), ref = "non-hispanic/unknown")

lang_cat_1_2_arf_cohort <- relevel_paralytic_subgroup(total_cohort[arf_flag == TRUE])
lang_cat_1_2_arf_cohort$lang_cat_1 <- relevel(as.factor(lang_cat_1_2_arf_cohort$lang_cat_1), ref = "english")
lang_cat_1_2_arf_cohort$lang_cat_2 <- relevel(as.factor(lang_cat_1_2_arf_cohort$lang_cat_2), ref = "english")

lang_cat_3_arf_cohort <- relevel_paralytic_subgroup(lang_cat_3_cohort[arf_flag == TRUE])
lang_cat_3_arf_cohort$lang_cat_3 <- relevel(as.factor(lang_cat_3_arf_cohort$lang_cat_3), ref = "english")

lang_cat_4_arf_cohort <- relevel_paralytic_subgroup(lang_cat_4_cohort[arf_flag == TRUE])
lang_cat_4_arf_cohort$lang_cat_4 <- relevel(as.factor(lang_cat_4_arf_cohort$lang_cat_4), ref = "english")

lang_cat_5_arf_cohort <- relevel_paralytic_subgroup(lang_cat_5_cohort[arf_flag == TRUE])
lang_cat_5_arf_cohort$lang_cat_5 <- relevel(as.factor(lang_cat_5_arf_cohort$lang_cat_5), ref = "english")

lang_cat_6_arf_cohort <- relevel_paralytic_subgroup(primary_cohort_all_icus[arf_flag == TRUE])
lang_cat_6_arf_cohort$lang_cat_6 <- relevel(as.factor(lang_cat_6_arf_cohort$lang_cat_6), ref = "english")

lang_cat_7_arf_cohort <- relevel_paralytic_subgroup(total_cohort[arf_flag == TRUE])
lang_cat_7_arf_cohort$lang_cat_7 <- relevel(as.factor(lang_cat_7_arf_cohort$lang_cat_7), ref = "english")

lang_cat_1_2_non_arf_cohort <- relevel_paralytic_subgroup(total_cohort[arf_flag == FALSE])
lang_cat_1_2_non_arf_cohort$lang_cat_1 <- relevel(as.factor(lang_cat_1_2_non_arf_cohort$lang_cat_1), ref = "english")
lang_cat_1_2_non_arf_cohort$lang_cat_2 <- relevel(as.factor(lang_cat_1_2_non_arf_cohort$lang_cat_2), ref = "english")

lang_cat_3_non_arf_cohort <- relevel_paralytic_subgroup(lang_cat_3_cohort[arf_flag == FALSE])
lang_cat_3_non_arf_cohort$lang_cat_3 <- relevel(as.factor(lang_cat_3_non_arf_cohort$lang_cat_3), ref = "english")

lang_cat_4_non_arf_cohort <- relevel_paralytic_subgroup(lang_cat_4_cohort[arf_flag == FALSE])
lang_cat_4_non_arf_cohort$lang_cat_4 <- relevel(as.factor(lang_cat_4_non_arf_cohort$lang_cat_4), ref = "english")

lang_cat_5_non_arf_cohort <- relevel_paralytic_subgroup(lang_cat_5_cohort[arf_flag == FALSE])
lang_cat_5_non_arf_cohort$lang_cat_5 <- relevel(as.factor(lang_cat_5_non_arf_cohort$lang_cat_5), ref = "english")

lang_cat_6_non_arf_cohort <- relevel_paralytic_subgroup(primary_cohort_all_icus[arf_flag == FALSE])
lang_cat_6_non_arf_cohort$lang_cat_6 <- relevel(as.factor(lang_cat_6_non_arf_cohort$lang_cat_6), ref = "english")

lang_cat_7_non_arf_cohort <- relevel_paralytic_subgroup(total_cohort[arf_flag == FALSE])
lang_cat_7_non_arf_cohort$lang_cat_7 <- relevel(as.factor(lang_cat_7_non_arf_cohort$lang_cat_7), ref = "english")

paralytic_exploratory_analysis <- read_parquet(paste0(intermediate_dir, "analytic_file", institution, ".parquet"))
setDT(paralytic_exploratory_analysis)

paralytic_exploratory_analysis[, lang_cat_1 := fifelse(
  language_category == "english", "english",
  fifelse(!is.na(language_category), "non-english", NA_character_)
)]

paralytic_exploratory_analysis$lang_cat_1               <- relevel(as.factor(paralytic_exploratory_analysis$lang_cat_1),               ref = "english")
paralytic_exploratory_analysis$sex_category             <- relevel(as.factor(paralytic_exploratory_analysis$sex_category),             ref = "male")
paralytic_exploratory_analysis$race_category            <- relevel(as.factor(paralytic_exploratory_analysis$race_category),            ref = "white")
paralytic_exploratory_analysis$ethnicity_category       <- relevel(as.factor(paralytic_exploratory_analysis$ethnicity_category),       ref = "non-hispanic")
paralytic_exploratory_analysis$ethnicity_category_final <- relevel(as.factor(paralytic_exploratory_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")

arf_paralytic_exploratory_analysis <- paralytic_exploratory_analysis[arf_flag == TRUE]
arf_paralytic_exploratory_analysis$lang_cat_1               <- relevel(as.factor(arf_paralytic_exploratory_analysis$lang_cat_1),               ref = "english")
arf_paralytic_exploratory_analysis$sex_category             <- relevel(as.factor(arf_paralytic_exploratory_analysis$sex_category),             ref = "male")
arf_paralytic_exploratory_analysis$race_category            <- relevel(as.factor(arf_paralytic_exploratory_analysis$race_category),            ref = "white")
arf_paralytic_exploratory_analysis$ethnicity_category       <- relevel(as.factor(arf_paralytic_exploratory_analysis$ethnicity_category),       ref = "non-hispanic")
arf_paralytic_exploratory_analysis$ethnicity_category_final <- relevel(as.factor(arf_paralytic_exploratory_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")

non_arf_paralytic_exploratory_analysis <- paralytic_exploratory_analysis[arf_flag == FALSE]
non_arf_paralytic_exploratory_analysis$lang_cat_1               <- relevel(as.factor(non_arf_paralytic_exploratory_analysis$lang_cat_1),               ref = "english")
non_arf_paralytic_exploratory_analysis$sex_category             <- relevel(as.factor(non_arf_paralytic_exploratory_analysis$sex_category),             ref = "male")
non_arf_paralytic_exploratory_analysis$race_category            <- relevel(as.factor(non_arf_paralytic_exploratory_analysis$race_category),            ref = "white")
non_arf_paralytic_exploratory_analysis$ethnicity_category       <- relevel(as.factor(non_arf_paralytic_exploratory_analysis$ethnicity_category),       ref = "non-hispanic")
non_arf_paralytic_exploratory_analysis$ethnicity_category_final <- relevel(as.factor(non_arf_paralytic_exploratory_analysis$ethnicity_category_final), ref = "non-hispanic/unknown")

cont_vars_explore <- c("age_at_admission", "bmi", "elixhauser_score", "SVM_SCORE",
                       "laps2", "time_to_first_rass", "rass_frequency_hrs",
                       "observation_year")

# Covariate set used for all exploratory disparity models — note no SVM_SCORE,
# matches what the original code used.
exploratory_covariates_paralytic <- c(
  "race_category", "lang_cat_1", "age_category", "bmi", "sex_category",
  "factor(hospital_icu_id)", "elixhauser_score", "laps2",
  "factor(observation_year)"
)

```

#Risk Difference in Paralytic Use for Paralytic Project
```{r}
current_project <- "paralytic"
################################################################################
# PARALYTIC PROJECT - Model fitting + marginal RD extraction
# Table 1s omitted (no changes needed)
################################################################################

# ── STEP 1: All IMV patients ──────────────────────────────────────────────────

lang_cat_1_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_1",
  covariates  = paralytic_covariates,
  data        = total_cohort,
  model_name  = "lang_cat_1_nmbas",
  cohort_name = "total_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_1_nmbas)) {
  fwrite(extract_glm_results(lang_cat_1_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_1.csv")))
  fwrite(extract_glm_marginal(
    model        = lang_cat_1_nmbas,
    data         = total_cohort,
    exposure     = "lang_cat_1",
    ref_level    = "english",
    target_level = "non-english",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_1_nmbas",
    cohort_name  = "total_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_1_marginal.csv")))
}

lang_cat_2_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_2",
  covariates  = paralytic_covariates,
  data        = total_cohort,
  model_name  = "lang_cat_2_nmbas",
  cohort_name = "total_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_2_nmbas)) {
  fwrite(extract_glm_results(lang_cat_2_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_2.csv")))
  # lang_cat_2 has 3 levels (english/spanish/other non-english) — run marginal
  # separately for each non-english comparator
  fwrite(extract_glm_marginal(
    model        = lang_cat_2_nmbas,
    data         = total_cohort,
    exposure     = "lang_cat_2",
    ref_level    = "english",
    target_level = "spanish",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_2_nmbas_spanish",
    cohort_name  = "total_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_2_marginal_spanish.csv")))
  fwrite(extract_glm_marginal(
    model        = lang_cat_2_nmbas,
    data         = total_cohort,
    exposure     = "lang_cat_2",
    ref_level    = "english",
    target_level = "other non-english",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_2_nmbas_other",
    cohort_name  = "total_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_2_marginal_other_non_english.csv")))
}

lang_cat_3_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_3",
  covariates  = paralytic_covariates,
  data        = lang_cat_3_cohort,
  model_name  = "lang_cat_3_nmbas",
  cohort_name = "lang_cat_3_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_3_nmbas)) {
  fwrite(extract_glm_results(lang_cat_3_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_3.csv")))
  fwrite(extract_glm_marginal(
    model        = lang_cat_3_nmbas,
    data         = lang_cat_3_cohort,
    exposure     = "lang_cat_3",
    ref_level    = "english",
    target_level = "spanish",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_3_nmbas_spanish",
    cohort_name  = "lang_cat_3_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_3_marginal_spanish.csv")))
  fwrite(extract_glm_marginal(
    model        = lang_cat_3_nmbas,
    data         = lang_cat_3_cohort,
    exposure     = "lang_cat_3",
    ref_level    = "english",
    target_level = "third_lang_cat",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_3_nmbas_third_lang",
    cohort_name  = "lang_cat_3_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_3_marginal_third_lang.csv")))
}

lang_cat_4_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_4",
  covariates  = paralytic_covariates,
  data        = lang_cat_4_cohort,
  model_name  = "lang_cat_4_nmbas",
  cohort_name = "lang_cat_4_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_4_nmbas)) {
  fwrite(extract_glm_results(lang_cat_4_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_4.csv")))
  for (tgt in c("spanish", "chinese", "vietnamese")) {
    fwrite(extract_glm_marginal(
      model        = lang_cat_4_nmbas,
      data         = lang_cat_4_cohort,
      exposure     = "lang_cat_4",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("lang_cat_4_nmbas_", tgt),
      cohort_name  = "lang_cat_4_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_4_marginal_", tgt, ".csv")))
  }
}

lang_cat_5_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_5",
  covariates  = paralytic_covariates,
  data        = lang_cat_5_cohort,
  model_name  = "lang_cat_5_nmbas",
  cohort_name = "lang_cat_5_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_5_nmbas)) {
  fwrite(extract_glm_results(lang_cat_5_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_5.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "continental_european_langs", "mena")) {
    fwrite(extract_glm_marginal(
      model        = lang_cat_5_nmbas,
      data         = lang_cat_5_cohort,
      exposure     = "lang_cat_5",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("lang_cat_5_nmbas_", tgt),
      cohort_name  = "lang_cat_5_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_5_marginal_", tgt, ".csv")))
  }
}

lang_cat_6_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_6",
  covariates  = paralytic_covariates,
  data        = primary_cohort_all_icus,
  model_name  = "lang_cat_6_nmbas",
  cohort_name = "primary_cohort_all_icus",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_6_nmbas)) {
  fwrite(extract_glm_results(lang_cat_6_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_6.csv")))
  fwrite(extract_glm_marginal(
    model        = lang_cat_6_nmbas,
    data         = primary_cohort_all_icus,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "lang_cat_6_nmbas",
    cohort_name  = "primary_cohort_all_icus"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_6_marginal.csv")))
}

lang_cat_7_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_7",
  covariates  = paralytic_covariates,
  data        = total_cohort,
  model_name  = "lang_cat_7_nmbas",
  cohort_name = "total_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(lang_cat_7_nmbas)) {
  fwrite(extract_glm_results(lang_cat_7_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_7.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "other_asian", "sub_saharan_african", "continental_european_langs",
                "mena", "all_other_langs")) {
    fwrite(extract_glm_marginal(
      model        = lang_cat_7_nmbas,
      data         = total_cohort,
      exposure     = "lang_cat_7",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("lang_cat_7_nmbas_", tgt),
      cohort_name  = "total_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_7_marginal_", tgt, ".csv")))
  }
}

# ── STEP 2: ARF Subgroup ──────────────────────────────────────────────────────

arf_lang_cat_1_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_1",
  covariates  = paralytic_covariates,
  data        = lang_cat_1_2_arf_cohort,
  model_name  = "arf_lang_cat_1_nmbas",
  cohort_name = "lang_cat_1_2_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_1_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_1_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_1_arf.csv")))
  fwrite(extract_glm_marginal(
    model        = arf_lang_cat_1_nmbas,
    data         = lang_cat_1_2_arf_cohort,
    exposure     = "lang_cat_1",
    ref_level    = "english",
    target_level = "non-english",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "arf_lang_cat_1_nmbas",
    cohort_name  = "lang_cat_1_2_arf_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_1_arf_marginal.csv")))
}

arf_lang_cat_2_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_2",
  covariates  = paralytic_covariates,
  data        = lang_cat_1_2_arf_cohort,
  model_name  = "arf_lang_cat_2_nmbas",
  cohort_name = "lang_cat_1_2_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_2_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_2_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_2_arf.csv")))
  for (tgt in c("spanish", "other non-english")) {
    fwrite(extract_glm_marginal(
      model        = arf_lang_cat_2_nmbas,
      data         = lang_cat_1_2_arf_cohort,
      exposure     = "lang_cat_2",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("arf_lang_cat_2_nmbas_", gsub(" ", "_", tgt)),
      cohort_name  = "lang_cat_1_2_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_2_arf_marginal_",
                        gsub(" ", "_", tgt), ".csv")))
  }
}

arf_lang_cat_3_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_3",
  covariates  = paralytic_covariates,
  data        = lang_cat_3_arf_cohort,
  model_name  = "arf_lang_cat_3_nmbas",
  cohort_name = "lang_cat_3_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_3_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_3_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_3_arf.csv")))
  for (tgt in c("spanish", "third_lang_cat")) {
    fwrite(extract_glm_marginal(
      model        = arf_lang_cat_3_nmbas,
      data         = lang_cat_3_arf_cohort,
      exposure     = "lang_cat_3",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("arf_lang_cat_3_nmbas_", tgt),
      cohort_name  = "lang_cat_3_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_3_arf_marginal_", tgt, ".csv")))
  }
}

arf_lang_cat_4_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_4",
  covariates  = paralytic_covariates,
  data        = lang_cat_4_arf_cohort,
  model_name  = "arf_lang_cat_4_nmbas",
  cohort_name = "lang_cat_4_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_4_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_4_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_4_arf.csv")))
  for (tgt in c("spanish", "chinese", "vietnamese")) {
    fwrite(extract_glm_marginal(
      model        = arf_lang_cat_4_nmbas,
      data         = lang_cat_4_arf_cohort,
      exposure     = "lang_cat_4",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("arf_lang_cat_4_nmbas_", tgt),
      cohort_name  = "lang_cat_4_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_4_arf_marginal_", tgt, ".csv")))
  }
}

arf_lang_cat_5_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_5",
  covariates  = paralytic_covariates,
  data        = lang_cat_5_arf_cohort,
  model_name  = "arf_lang_cat_5_nmbas",
  cohort_name = "lang_cat_5_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_5_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_5_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_5_arf.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "continental_european_langs", "mena")) {
    fwrite(extract_glm_marginal(
      model        = arf_lang_cat_5_nmbas,
      data         = lang_cat_5_arf_cohort,
      exposure     = "lang_cat_5",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("arf_lang_cat_5_nmbas_", tgt),
      cohort_name  = "lang_cat_5_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_5_arf_marginal_", tgt, ".csv")))
  }
}

arf_lang_cat_6_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_6",
  covariates  = paralytic_covariates,
  data        = lang_cat_6_arf_cohort,
  model_name  = "arf_lang_cat_6_nmbas",
  cohort_name = "lang_cat_6_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_6_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_6_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_6_arf.csv")))
  fwrite(extract_glm_marginal(
    model        = arf_lang_cat_6_nmbas,
    data         = lang_cat_6_arf_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "arf_lang_cat_6_nmbas",
    cohort_name  = "lang_cat_6_arf_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_6_arf_marginal.csv")))
}

arf_lang_cat_7_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_7",
  covariates  = paralytic_covariates,
  data        = lang_cat_7_arf_cohort,
  model_name  = "arf_lang_cat_7_nmbas",
  cohort_name = "lang_cat_7_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(arf_lang_cat_7_nmbas)) {
  fwrite(extract_glm_results(arf_lang_cat_7_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_7_arf.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "other_asian", "sub_saharan_african", "continental_european_langs",
                "mena", "all_other_langs")) {
    fwrite(extract_glm_marginal(
      model        = arf_lang_cat_7_nmbas,
      data         = lang_cat_7_arf_cohort,
      exposure     = "lang_cat_7",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("arf_lang_cat_7_nmbas_", tgt),
      cohort_name  = "lang_cat_7_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_7_arf_marginal_", tgt, ".csv")))
  }
}

# ── STEP 3: Non-ARF Subgroup ──────────────────────────────────────────────────

non_arf_lang_cat_1_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_1",
  covariates  = paralytic_covariates,
  data        = lang_cat_1_2_non_arf_cohort,
  model_name  = "non_arf_lang_cat_1_nmbas",
  cohort_name = "lang_cat_1_2_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_1_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_1_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_1_non_arf.csv")))
  fwrite(extract_glm_marginal(
    model        = non_arf_lang_cat_1_nmbas,
    data         = lang_cat_1_2_non_arf_cohort,
    exposure     = "lang_cat_1",
    ref_level    = "english",
    target_level = "non-english",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "non_arf_lang_cat_1_nmbas",
    cohort_name  = "lang_cat_1_2_non_arf_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_1_non_arf_marginal.csv")))
}

non_arf_lang_cat_2_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_2",
  covariates  = paralytic_covariates,
  data        = lang_cat_1_2_non_arf_cohort,
  model_name  = "non_arf_lang_cat_2_nmbas",
  cohort_name = "lang_cat_1_2_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_2_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_2_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_2_non_arf.csv")))
  for (tgt in c("spanish", "other non-english")) {
    fwrite(extract_glm_marginal(
      model        = non_arf_lang_cat_2_nmbas,
      data         = lang_cat_1_2_non_arf_cohort,
      exposure     = "lang_cat_2",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("non_arf_lang_cat_2_nmbas_", gsub(" ", "_", tgt)),
      cohort_name  = "lang_cat_1_2_non_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_2_non_arf_marginal_",
                        gsub(" ", "_", tgt), ".csv")))
  }
}

non_arf_lang_cat_3_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_3",
  covariates  = paralytic_covariates,
  data        = lang_cat_3_non_arf_cohort,
  model_name  = "non_arf_lang_cat_3_nmbas",
  cohort_name = "lang_cat_3_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_3_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_3_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_3_non_arf.csv")))
  for (tgt in c("spanish", "third_lang_cat")) {
    fwrite(extract_glm_marginal(
      model        = non_arf_lang_cat_3_nmbas,
      data         = lang_cat_3_non_arf_cohort,
      exposure     = "lang_cat_3",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("non_arf_lang_cat_3_nmbas_", tgt),
      cohort_name  = "lang_cat_3_non_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_3_non_arf_marginal_", tgt, ".csv")))
  }
}

non_arf_lang_cat_4_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_4",
  covariates  = paralytic_covariates,
  data        = lang_cat_4_non_arf_cohort,
  model_name  = "non_arf_lang_cat_4_nmbas",
  cohort_name = "lang_cat_4_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_4_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_4_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_4_non_arf.csv")))
  for (tgt in c("spanish", "chinese", "vietnamese")) {
    fwrite(extract_glm_marginal(
      model        = non_arf_lang_cat_4_nmbas,
      data         = lang_cat_4_non_arf_cohort,
      exposure     = "lang_cat_4",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("non_arf_lang_cat_4_nmbas_", tgt),
      cohort_name  = "lang_cat_4_non_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_4_non_arf_marginal_", tgt, ".csv")))
  }
}

non_arf_lang_cat_5_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_5",
  covariates  = paralytic_covariates,
  data        = lang_cat_5_non_arf_cohort,
  model_name  = "non_arf_lang_cat_5_nmbas",
  cohort_name = "lang_cat_5_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_5_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_5_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_5_non_arf.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "continental_european_langs", "mena")) {
    fwrite(extract_glm_marginal(
      model        = non_arf_lang_cat_5_nmbas,
      data         = lang_cat_5_non_arf_cohort,
      exposure     = "lang_cat_5",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("non_arf_lang_cat_5_nmbas_", tgt),
      cohort_name  = "lang_cat_5_non_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_5_non_arf_marginal_", tgt, ".csv")))
  }
}

non_arf_lang_cat_6_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_6",
  covariates  = paralytic_covariates,
  data        = lang_cat_6_non_arf_cohort,
  model_name  = "non_arf_lang_cat_6_nmbas",
  cohort_name = "lang_cat_6_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_6_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_6_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_6_non_arf.csv")))
  fwrite(extract_glm_marginal(
    model        = non_arf_lang_cat_6_nmbas,
    data         = lang_cat_6_non_arf_cohort,
    exposure     = "lang_cat_6",
    ref_level    = "english",
    target_level = "spanish",
    n_boot       = 100,
    site_name    = institution,
    model_name   = "non_arf_lang_cat_6_nmbas",
    cohort_name  = "lang_cat_6_non_arf_cohort"
  ), file.path(rerun_paralytics,
               paste0(institution, "_paralysis_lang_cat_6_non_arf_marginal.csv")))
}

non_arf_lang_cat_7_nmbas <- fit_model_safe(
  outcome     = "continuous_nmba_48",
  exposure    = "lang_cat_7",
  covariates  = paralytic_covariates,
  data        = lang_cat_7_non_arf_cohort,
  model_name  = "non_arf_lang_cat_7_nmbas",
  cohort_name = "lang_cat_7_non_arf_cohort",
  project     = current_project,
  family      = binomial(link = "logit")
)
if (!is.null(non_arf_lang_cat_7_nmbas)) {
  fwrite(extract_glm_results(non_arf_lang_cat_7_nmbas),
         file.path(rerun_paralytics, paste0(institution, "_paralysis_lang_cat_7_non_arf.csv")))
  for (tgt in c("spanish", "south_asian", "east_asian", "southeast_asian",
                "other_asian", "sub_saharan_african", "continental_european_langs",
                "mena", "all_other_langs")) {
    fwrite(extract_glm_marginal(
      model        = non_arf_lang_cat_7_nmbas,
      data         = lang_cat_7_non_arf_cohort,
      exposure     = "lang_cat_7",
      ref_level    = "english",
      target_level = tgt,
      n_boot       = 100,
      site_name    = institution,
      model_name   = paste0("non_arf_lang_cat_7_nmbas_", tgt),
      cohort_name  = "lang_cat_7_non_arf_cohort"
    ), file.path(rerun_paralytics,
                 paste0(institution, "_paralysis_lang_cat_7_non_arf_marginal_", tgt, ".csv")))
  }
}

# ── STEP 4: Pandemic period loop ──────────────────────────────────────────────
# NOTE: marginal RD only added for lang_cat_6 (binary English/Spanish).
# Multi-level lang_cats loop over comparators the same way as Steps 1-3 above.

################################################################################
# Define pandemic-period cohorts for each language category × time period
# combination. Each base cohort gets filtered by admission_dttm.
################################################################################
pandemic_periods <- list(
  pre_pandemic = list(filter = function(dt) dt[admission_dttm < as.POSIXct("2020-03-01 00:00:00")]),
  pandemic     = list(filter = function(dt) dt[admission_dttm >= as.POSIXct("2020-03-01 00:00:00") &
                                               admission_dttm <= as.POSIXct("2021-12-31 23:59:59")]),
  endemic      = list(filter = function(dt) dt[admission_dttm > as.POSIXct("2021-12-31 23:59:59")])
)

# Map of which lang_cat to use and which base ARF cohort to filter from
pandemic_specs <- list(
  list(lang_var = "lang_cat_1", base_cohort = lang_cat_1_2_arf_cohort),
  list(lang_var = "lang_cat_2", base_cohort = lang_cat_1_2_arf_cohort),
  list(lang_var = "lang_cat_3", base_cohort = lang_cat_3_arf_cohort),
  list(lang_var = "lang_cat_4", base_cohort = lang_cat_4_arf_cohort),
  list(lang_var = "lang_cat_5", base_cohort = lang_cat_5_arf_cohort),
  list(lang_var = "lang_cat_6", base_cohort = lang_cat_6_arf_cohort),
  list(lang_var = "lang_cat_7", base_cohort = lang_cat_7_arf_cohort)
)

for (spec in pandemic_specs) {
  for (period_name in names(pandemic_periods)) {
    
    cohort_name <- paste0(spec$lang_var, "_arf_", period_name, "_cohort")
    file_stem   <- paste0(spec$lang_var, "_arf_", period_name, "_cohort")
    model_name  <- paste0(spec$lang_var, "_arf_", period_name, "_nmbas")
    out_file    <- paste0(institution, "_paralysis_", spec$lang_var, "_arf_", period_name, ".csv")
    
    # Build cohort: filter by period, then re-level
    period_cohort <- pandemic_periods[[period_name]]$filter(spec$base_cohort)
    period_cohort <- relevel_paralytic_subgroup(period_cohort)
    period_cohort[[spec$lang_var]] <- relevel(
      as.factor(period_cohort[[spec$lang_var]]), ref = "english"
    )
    
    make_table1_pair(
      dt = period_cohort, by = spec$lang_var, reference_level = "english",
      cat_vars = cat_vars, cont_vars = cont_vars, site_name = institution,
      tables_dir = paralytics_tables_dir, institution_prefix = institution,
      cohort_name = cohort_name, file_stem = file_stem,
      project = current_project
    )
    
    fit <- fit_model_safe(
      outcome     = "continuous_nmba_48",
      exposure    = spec$lang_var,
      covariates  = paralytic_covariates,
      data        = period_cohort,
      model_name  = model_name,
      cohort_name = cohort_name,
      project     = current_project,
      family      = binomial(link = "logit")
    )
    if (!is.null(fit)) {
      fwrite(extract_glm_results(fit),
             file.path(paralytics_final_data, out_file))
    }
    
    # Save the cohort to the global env so downstream sections (ARF project)
    # can reference it by the original variable name
    assign(cohort_name, period_cohort)
  }
}

for (spec in pandemic_specs) {
  for (period_name in names(pandemic_periods)) {

    cohort_name  <- paste0(spec$lang_var, "_arf_", period_name, "_cohort")
    file_stem    <- paste0(spec$lang_var, "_arf_", period_name, "_cohort")
    model_name   <- paste0(spec$lang_var, "_arf_", period_name, "_nmbas")
    out_file     <- paste0(institution, "_paralysis_", spec$lang_var, "_arf_", period_name, ".csv")

    period_cohort <- pandemic_periods[[period_name]]$filter(spec$base_cohort)
    period_cohort <- relevel_paralytic_subgroup(period_cohort)
    period_cohort[[spec$lang_var]] <- relevel(
      as.factor(period_cohort[[spec$lang_var]]), ref = "english"
    )

    fit <- fit_model_safe(
      outcome     = "continuous_nmba_48",
      exposure    = spec$lang_var,
      covariates  = paralytic_covariates,
      data        = period_cohort,
      model_name  = model_name,
      cohort_name = cohort_name,
      project     = current_project,
      family      = binomial(link = "logit")
    )
    if (!is.null(fit)) {
      fwrite(extract_glm_results(fit),
             file.path(rerun_paralytics, out_file))

      # Define comparators based on lang_cat
      comparators <- switch(spec$lang_var,
        "lang_cat_1" = "non-english",
        "lang_cat_2" = c("spanish", "other non-english"),
        "lang_cat_3" = c("spanish", "third_lang_cat"),
        "lang_cat_4" = c("spanish", "chinese", "vietnamese"),
        "lang_cat_5" = c("spanish", "south_asian", "east_asian", "southeast_asian",
                         "continental_european_langs", "mena"),
        "lang_cat_6" = "spanish",
        "lang_cat_7" = c("spanish", "south_asian", "east_asian", "southeast_asian",
                         "other_asian", "sub_saharan_african", "continental_european_langs",
                         "mena", "all_other_langs")
      )

      for (tgt in comparators) {
        marginal_file <- paste0(
          institution, "_paralysis_", spec$lang_var, "_arf_", period_name,
          "_marginal_", gsub(" ", "_", tgt), ".csv"
        )
        tryCatch(
          fwrite(extract_glm_marginal(
            model        = fit,
            data         = period_cohort,
            exposure     = spec$lang_var,
            ref_level    = "english",
            target_level = tgt,
            n_boot       = 100,
            site_name    = institution,
            model_name   = paste0(model_name, "_", gsub(" ", "_", tgt)),
            cohort_name  = cohort_name
          ), file.path(rerun_paralytics, marginal_file)),
          error = function(e) message(
            "Skipping marginal for ", spec$lang_var, " / ", period_name,
            " / ", tgt, ": ", conditionMessage(e)
          )
        )
      }
    }
    assign(cohort_name, period_cohort)
  }
}

# ── STEP 5: Exploratory analyses ──────────────────────────────────────────────
# NOTE: Interaction models cannot use extract_glm_marginal directly — marginal
# standardization with interaction terms requires specifying which variable is
# being standardized over, which is not supported by the current function.
# extract_glm_results only for those models.
exploratory_covariates_paralytic <- c(
  "race_category", "lang_cat_1", "age_category", "bmi", "sex_category",
  "factor(hospital_icu_id)", "elixhauser_score", "laps2",
  "factor(observation_year)"
)

# cat_vars used in this block excludes ethnicity_category since it's the `by`
cat_vars_explore_ethn <- c("sex_category", "race_category", "lang_cat_1",
                           "hospital_id", "icu_imv", "location_of_intubation",
                           "admission_type_category", "arf_flag", "aHrf_flag")

# Run the same pattern across the three subcohorts
ethnicity_specs <- list(
  list(name = "all_imv", data = paralytic_exploratory_analysis,         model_name = "ethnicity_all_imv_nmbas"),
  list(name = "arf",     data = arf_paralytic_exploratory_analysis,     model_name = "ethnicity_arf_nmbas"),
  list(name = "non_arf", data = non_arf_paralytic_exploratory_analysis, model_name = "ethnicity_non_arf_nmbas")
)

for (spec in ethnicity_specs) {
  fit <- fit_model_safe(
    outcome     = "continuous_nmba_48",
    exposure    = "ethnicity_category_final",
    covariates  = exploratory_covariates_paralytic,
    data        = spec$data,
    model_name  = spec$model_name,
    cohort_name = paste0("paralytic_ethnicity_", spec$name),
    project     = current_project,
    family      = binomial(link = "logit")
  )
  if (!is.null(fit)) {
    fwrite(extract_glm_results(fit),
           file.path(rerun_paralytics,
                     paste0(institution, "_paralysis_ethnicity_", spec$name, ".csv")))
    for (tgt in c("hispanic")) {
      tryCatch(
        fwrite(extract_glm_marginal(
          model        = fit,
          data         = spec$data,
          exposure     = "ethnicity_category_final",
          ref_level    = "non-hispanic/unknown",
          target_level = tgt,
          n_boot       = 100,
          site_name    = institution,
          model_name   = paste0(spec$model_name, "_", tgt),
          cohort_name  = paste0("paralytic_ethnicity_", spec$name)
        ), file.path(rerun_paralytics,
                     paste0(institution, "_paralysis_ethnicity_", spec$name,
                            "_marginal_", tgt, ".csv"))),
        error = function(e) message(
          "Skipping marginal ethnicity ", spec$name, " / ", tgt, ": ", conditionMessage(e)
        )
      )
    }
  }
}
cat_vars_explore_race <- c("sex_category", "ethnicity_category_final", "lang_cat_1",
                           "hospital_id", "icu_imv", "location_of_intubation",
                           "admission_type_category", "arf_flag", "aHrf_flag")

# Race covariates: race is exposure, so use ethnicity_category in covariates instead
race_covariates_paralytic <- c(
  "ethnicity_category_final", "lang_cat_1", "age_category", "bmi", "sex_category",
  "factor(hospital_icu_id)", "elixhauser_score", "laps2",
  "factor(observation_year)"
)

race_specs <- list(
  list(name = "all_imv", data = paralytic_exploratory_analysis,         model_name = "race_all_imv_nmbas"),
  list(name = "arf",     data = arf_paralytic_exploratory_analysis,     model_name = "race_arf_nmbas"),
  list(name = "non_arf", data = non_arf_paralytic_exploratory_analysis, model_name = "race_non_arf_nmbas")
)

for (spec in race_specs) {
  fit <- fit_model_safe(
    outcome     = "continuous_nmba_48",
    exposure    = "race_category",
    covariates  = race_covariates_paralytic,
    data        = spec$data,
    model_name  = spec$model_name,
    cohort_name = paste0("paralytic_race_", spec$name),
    project     = current_project,
    family      = binomial(link = "logit")
  )
  if (!is.null(fit)) {
    fwrite(extract_glm_results(fit),
           file.path(rerun_paralytics,
                     paste0(institution, "_paralysis_race_", spec$name, ".csv")))
    for (tgt in c("black", "hispanic", "asian", "other/unknown")) {
      tryCatch(
        fwrite(extract_glm_marginal(
          model        = fit,
          data         = spec$data,
          exposure     = "race_category",
          ref_level    = "white",
          target_level = tgt,
          n_boot       = 100,
          site_name    = institution,
          model_name   = paste0(spec$model_name, "_", gsub("/| ", "_", tgt)),
          cohort_name  = paste0("paralytic_race_", spec$name)
        ), file.path(rerun_paralytics,
                     paste0(institution, "_paralysis_race_", spec$name,
                            "_marginal_", gsub("/| ", "_", tgt), ".csv"))),
        error = function(e) message(
          "Skipping marginal race ", spec$name, " / ", tgt, ": ", conditionMessage(e)
        )
      )
    }
  }
}

cat_vars_explore_sex <- c("ethnicity_category_final", "race_category", "lang_cat_1",
                          "hospital_id", "icu_imv", "location_of_intubation",
                          "admission_type_category", "arf_flag", "aHrf_flag")

# Sex covariates: sex is exposure, race + ethnicity stay as covariates
sex_covariates_paralytic <- c(
  "race_category", "ethnicity_category_final", "lang_cat_1", "age_category", "bmi",
  "factor(hospital_icu_id)", "elixhauser_score", "laps2",
  "factor(observation_year)"
)

sex_specs <- list(
  list(name = "all_imv", data = paralytic_exploratory_analysis,         model_name = "sex_all_imv_nmbas"),
  list(name = "arf",     data = arf_paralytic_exploratory_analysis,     model_name = "sex_arf_nmbas"),
  list(name = "non_arf", data = non_arf_paralytic_exploratory_analysis, model_name = "sex_non_arf_nmbas")
)

for (spec in sex_specs) {
  fit <- fit_model_safe(
    outcome     = "continuous_nmba_48",
    exposure    = "sex_category",
    covariates  = sex_covariates_paralytic,
    data        = spec$data,
    model_name  = spec$model_name,
    cohort_name = paste0("paralytic_sex_", spec$name),
    project     = current_project,
    family      = binomial(link = "logit")
  )
  if (!is.null(fit)) {
    fwrite(extract_glm_results(fit),
           file.path(rerun_paralytics,
                     paste0(institution, "_paralysis_sex_", spec$name, ".csv")))
    tryCatch(
      fwrite(extract_glm_marginal(
        model        = fit,
        data         = spec$data,
        exposure     = "sex_category",
        ref_level    = "male",
        target_level = "female",
        n_boot       = 100,
        site_name    = institution,
        model_name   = paste0(spec$model_name, "_female"),
        cohort_name  = paste0("paralytic_sex_", spec$name)
      ), file.path(rerun_paralytics,
                   paste0(institution, "_paralysis_sex_", spec$name, "_marginal_female.csv"))),
      error = function(e) message(
        "Skipping marginal sex ", spec$name, ": ", conditionMessage(e)
      )
    )
  }
}

# Interaction models — extract_glm_results only, marginal not applicable
#if (!is.null(paralytic_model_total_cohort_race_language_interaction)) {
#  fwrite(extract_glm_results(paralytic_model_total_cohort_race_language_interaction),
#         file.path(rerun_paralytics,
#                   paste0(institution, "_paralytic_model_total_cohort_race_language_interaction.csv")))
#}
#if (!is.null(paralytic_model_total_cohort_ethnicity_language_interaction)) {
#  fwrite(extract_glm_results(paralytic_model_total_cohort_ethnicity_language_interaction),
#         file.path(rerun_paralytics,
#                   paste0(institution, "_paralytic_model_total_cohort_ethnicity_language_interaction.csv")))
#}
#if (!is.null(paralytic_model_arf_cohort_race_language_interaction)) {
#  fwrite(extract_glm_results(paralytic_model_arf_cohort_race_language_interaction),
#         file.path(rerun_paralytics,
#                   paste0(institution, "_paralytic_model_arf_cohort_race_language_interaction.csv")))
#}
##if (!is.null(paralytic_model_arf_cohort_ethnicity_language_interaction)) {
#  fwrite(extract_glm_results(paralytic_model_arf_cohort_ethnicity_language_interaction),
#         file.path(rerun_paralytics,
#                   paste0(institution, "_paralytic_model_arf_cohort_ethnicity_language_interaction.csv")))
#}

```